---
title: "Fish Analysis OCNMS"
author: "Ole Shelton"
date: "4/23/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
source("/Users/ole.shelton/GitHub/OCNMS/R scripts/Fish Analysis/Swath Fish.R")
```

# Exploratory analysis 
These are some analyses based on the 2015-17 survey data for fish. I've done a bunch of processing in the Git repo (file "/GitHub/OCNMS/R scripts/Fish Analysis/Swath Fish.R").

I first plotted a bunch of species together to look at the fish community.  Then I plot individual species by site and time. 

During the sampling, the sampling they observed a total of `r length(unique(dat.fish$species))` species and species groups.  A few of these are redundant, though.  Here is how the sampling was distributed across sites and depths among years (underscores indicate sampling depth).

```{r echo=FALSE, message=FALSE, warning=FALSE}
A<- dat.large.fish.summary %>% group_by(year,site,zone) %>% dplyr::summarise(N=max(N))
A$site <- factor(A$site,levels=nom)
#A <- left_join(A,nom.merge,by=c("site"="nom"))
B <- dcast(A, site~year+zone,value.var=c("N")) %>% left_join(.,nom.merge,by=c("site"="nom")) %>% arrange(id) %>% dplyr::select(-id)
B[is.na(B)==T] <- 0
print(B)
```

\newpage
Here is a cheat sheet for species abbreviations and associated common names (ordered by abbreviation): 
```{r echo=FALSE} 
print(SP.all.common.names %>% arrange(species))
```

\newpage
## Small Rockfish (<10cm)
Here are the rockfish plotted in a couple different ways.  This is for all small rockfish combined. I have included all transects here, not just those with visibility > 2m.  See the end for information about the transects and visibility.  In the bottom panel, the black boxes and error bars are simple the among site means and SE using the site means. For all panels, the 5 vs. 10 colums indicate water depth category.

```{r echo=FALSE, warning=FALSE, fig.rock.yoy1, fig.height = 4.25, fig.width = 7,fig.align='center'}
print(P.complex.small)
```
```{r echo=FALSE, warning=FALSE, fig.rock.yoy2, fig.height = 3.75, fig.width = 7,fig.align='center'}
print(Q.complex.small)
```

Here is why I didn't make plots of species by sub-types.  Different levels of taxonomic specificity are available for different observers and years. 
```{r echo=FALSE, warning=FALSE, fig.multispecies.small, fig.height = 4.5, fig.width = 7,fig.align='center'}
print(small.count.2015)
print(small.count.2016.5m)
print(small.count.2017.5m)
print(small.count.2016.10m)
print(small.count.2017.10m)
```

\newpage
## All Other Fish Except small rockfish
Here, here are the fish > 10 cm.  In general, you will notice a lot of grey in the figures.  This means most species were not observed in any transect at that site-year combination.  

```{r echo=FALSE, warning=FALSE, fig.multispecies.large, fig.height = 4.5, fig.width = 7,fig.align='center'}
print(all.sp.count.2015)
print(all.sp.count.2016.5m)
print(all.sp.count.2017.5m)
print(all.sp.count.2016.10m)
print(all.sp.count.2017.10m)
```
\newpage
# Example time series and spatial variation for the top 8 most common species.

```{r echo=FALSE,fig.height = 4.25, fig.width = 7,fig.align='center'}
for(i in 1:8){
    print(P.large[[i]])
    print(Q.large[[i]])  
}

```

\newpage
#### Some plots on visibility by site and depth zone

These suggest to me that the 2m cut off that PISCO uses is going to be problematic... there are a lot of surveys that are right on the cusp of 2m visibility.
```{r echo=FALSE,warning=FALSE,fig.height = 4.25, fig.width = 7,fig.align='center'}
print(vis.by.site)
print(vis.hist)
```