---
title: "Invert and KelpAnalysis OCNMS"
author: "Ole Shelton"
date: "01/20/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(xtable)
source("/Users/ole.shelton/GitHub/OCNMS/R scripts/Fish, Invert, Kelp Analysis/Swath Inverts.R")
```

# Exploratory analysis 
These are some analyses based on the 2015- survey data for invertebrates. I've done a bunch of processing in the Git repo (file "/GitHub/OCNMS/R scripts/Fish, Invert, Kelp Analysis/Swath Inverts.R"").

During the sampling, a total of `r A<-dat.swath %>% filter(group=="Invert") %>% dplyr::select(species) %>% as.data.frame(); length(unique(A$species))` invertebrate species and species groups were observed. Here is how the sampling was distributed across sites and 5m and 10m depths among years (2015-present). Surveys from pre-2015 were not collected by NWFSC and have a different design and methods.  They also are only available for some species

```{r  results="asis",echo=F,warning=F,message=F}
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

library(knitr)
invert.transects <- dat.invert %>% group_by(year,site,transect,observer,area,zone) %>% 
        summarise(n.sp=length(transect)) %>% group_by(year,site,zone) %>% 
        summarise(N=length(n.sp)) %>% dcast(year+site~zone)
invert.transects[is.na(invert.transects) ==T] <-0 

invert.transects$site <- factor(invert.transects$site,levels=
                                c("Destruction Island",
                                "Teahwhit Head",  
                                "Rock 305",
                                "Cape Johnson",
                                "Anderson Point",      
                                "Cape Alava",                  
                                "Point of the Arches", 
                                "Tatoosh Island",
                                "Chibadehl Rocks",
                                "Neah Bay"))
invert.transects <- invert.transects %>% arrange(year,site)
kable(invert.transects)
```

Let's go through the main species groups individually to look at changes through time.

## Sea urchins 
Urchins include these species

```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "urchin"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% urchin) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)

```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle("Sea Urchins, all depths") +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
        geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle("Sea Urchins by depth") +
    facet_wrap(~zone,nrow=2) +
    theme_bw()


A<- species_names %>% filter(species %in% urchin) %>% rename(code=species) %>% dplyr::select(-group)
P.sp <- list()
for(i in 1:nrow(A)){
P.sp[[i]] <- ggplot(dat.invert.species.zone %>% filter(species==A$code[i],site %in% SITES )) +
        geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    #geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(A$code[i],A$common.name[i])) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()
}

print(p1)
print(p2)
for(i in 1:nrow(A)){
    print(P.sp[[i]])
}
```
<!-- 
###################################################
SEASTAR SECTION
###################################################
-->

## Seastars 
Seastar plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "seastar"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% seastar) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)

```

We also see these species, but they are often below the size threshold (2.5cm radius) so we do not include them in the summaries below
```{r  results="asis",echo=F,warning=F,message=F}
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% seastar.2) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)

```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle("Seastars, all depths") +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle("Seastars by depth") +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
A<- species_names %>% filter(species %in% seastar) %>% rename(code=species) %>% dplyr::select(-group)
P.sp <- list()
for(i in 1:nrow(A)){
    if(nrow(dat.invert.species.zone %>% filter(species==A$code[i],site %in% SITES ))>0){
    upper.lim = dat.invert.species.zone %>% filter(species==A$code[i],site %in% SITES )
    upper.lim <- max(upper.lim$Mean+upper.lim$SE,na.rm=T)

    P.sp[[i]] <- ggplot(dat.invert.species.zone %>% filter(species==A$code[i],site %in% SITES )) +
        geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    #geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    scale_y_continuous(limits = c(0,upper.lim)) +
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(A$code[i],A$common.name[i])) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()
    }
}
for(i in 1:nrow(A)){
    quartz(file=paste("seastar",i,".jpeg"),type="jpeg",dpi=600,height=5,width=7); print(P.sp[[i]]); dev.off()
}

```
<!-- 
###################################################
CRAB SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")

-->

## Crabs 
Crabs plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "crab"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% crab) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)
```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```
<!-- 
###################################################
BIVALVE SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")

-->

## Bivalves 
Bivalve plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "bivalve"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% bivalve) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)


```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```
<!-- 
###################################################
CUCUMBER SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")

-->

## Sea Cucumber 
Sea cucumber plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "cucumber"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% cucumber) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)


```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```

<!-- 
###################################################
GASTROPOD SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")

-->

## Gastropods
Gastropod plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "gastropod"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% gastropod) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)


```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```

<!-- 
###################################################
TUNICATE SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")

-->

## Tunicates
Tunicates plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "tunicate"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% tunicate) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)


```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```

<!-- 
###################################################
TUNICATE SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")

-->

## Anenomes
Anenomes plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "anenome"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% anenome) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)


```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```

<!-- 
###################################################
NUDIBRANCH SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")
-->
## Nudibranch
Nudibranch plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "nudibranch"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% nudibranch) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)


```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```

<!-- 
###################################################
CHITON SECTION
###################################################
GROUPS <- c("urchin","seastar","crab","bivalve","chiton","cucumber","gastropod","tunicate","anenome")
-->
## Chiton
Chiton plots include these species
```{r  results="asis",echo=F,warning=F,message=F}
GROUP <- "chiton"
### The relevant dataframes are dat.invert and dat.algae for 2015 and later.
### dat.invert.density for 1987 to present.

# you can find out what the GROUPS are option 

library(knitr)
A<- species_names %>% filter(species %in% chiton) %>% rename(code=species) %>% dplyr::select(-group)
kable(A)


```

```{r ,echo=F,warning=F,message=F}
ALP <- 0.5
p1 <- ggplot(dat.invert.group %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"all depths")) +
    theme_bw()

p2 <- ggplot(dat.invert.group.zone %>% filter(group.name == GROUP,site %in% SITES )) +
    geom_point(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=Mean,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=Mean-SE,ymax=Mean+SE,color=site),alpha=ALP)+
    ylab(expression("Mean density (m"^-2*")"))+
    xlab("Year") +
    scale_color_discrete("Site") +
    ggtitle(paste(GROUP,"by depth")) +
    facet_wrap(~zone,nrow=2) +
    theme_bw()

print(p1)
print(p2)
```

# Algae

Ok. This section makes the same set of plots but does not group the species, but rather orders them from important to rare.


```{r ,echo=F,warning=F,message=F}
# drop No_ALG
dat.algae.density <- dat.algae.density %>% filter(!species == "NO_ALG")
dat.algae.density <- dat.algae.density %>% rename(Mean = MEAN)
dat.algae.density$species <- factor(dat.algae.density$species,
                                    levels= c(
                                      "MACPYR","NERLUE","PTECAL",
                                      "AGAFIM","ALAMAR","COSCOS",
                                      "CYMTRI","DESSPP","EGRMEN","LAMSET",
                                      "PLEGAR","SACGRO","SACLAT"))



ALP <- 0.5
p1 <- ggplot(dat.algae.density %>% 
               filter(site %in% SITES, species %in% c("MACPYR","NERLUE","PTECAL" ))) +
    geom_point(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=MEAN.site.zone-SE.site.zone,
                      ymax=MEAN.site.zone+SE.site.zone,color=site),alpha=ALP)+
    scale_y_continuous(expression("Mean density (m"^-2*")"),
                       trans="sqrt",expand=c(NA,0))+
   
    xlab("Year") +
    scale_color_discrete("Site") +
    #ggtitle(paste(GROUP,"all depths")) +
    facet_grid(species~zone,scales="free") +
    theme_bw()

p2 <-  ggplot(dat.algae.density %>% 
               filter(site %in% SITES, species %in% c("AGAFIM","ALAMAR","COSCOS" ))) +
    geom_point(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=MEAN.site.zone-SE.site.zone,
                      ymax=MEAN.site.zone+SE.site.zone,color=site),alpha=ALP)+
    scale_y_continuous(expression("Mean density (m"^-2*")"),
                       trans="sqrt",expand=c(NA,0))+
   
    xlab("Year") +
    scale_color_discrete("Site") +
    #ggtitle(paste(GROUP,"all depths")) +
    facet_grid(species~zone,scales="free") +
    theme_bw()

p3 <-  ggplot(dat.algae.density %>% 
             filter(site %in% SITES, species %in% c("CYMTRI","DESSPP","EGRMEN","LAMSET"))) +
    geom_point(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=MEAN.site.zone-SE.site.zone,
                      ymax=MEAN.site.zone+SE.site.zone,color=site),alpha=ALP)+
    scale_y_continuous(expression("Mean density (m"^-2*")"),
                       trans="sqrt",expand=c(NA,0))+
   
    xlab("Year") +
    scale_color_discrete("Site") +
    #ggtitle(paste(GROUP,"all depths")) +
    facet_grid(species~zone,scales="free") +
    theme_bw()

p4 <-  ggplot(dat.algae.density %>% 
             filter(site %in% SITES, species %in% c("PLEGAR","SACGRO","SACLAT"))) +
    geom_point(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_line(aes(x=year,y=MEAN.site.zone,color=site),alpha=ALP) +
    geom_errorbar(aes(x=year,ymin=MEAN.site.zone-SE.site.zone,
                      ymax=MEAN.site.zone+SE.site.zone,color=site),alpha=ALP)+
    scale_y_continuous(expression("Mean density (m"^-2*")"),
                       trans="sqrt",expand=c(NA,0))+
   
    xlab("Year") +
    scale_color_discrete("Site") +
    #ggtitle(paste(GROUP,"all depths")) +
    facet_grid(species~zone,scales="free") +
    theme_bw()

```

```{r ,echo=F,warning=F,message=F,fig.height=7}
print(p1)
```

```{r ,echo=F,warning=F,message=F,fig.height=7}
print(p2)

```

```{r ,echo=F,warning=F,message=F,fig.height=7}
print(p3)

```

```{r ,echo=F,warning=F,message=F,fig.height=7}
print(p4)
```

