---
title: "Kelp vs Urchins"
author: Nick Toimieri
date: sys.date()
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
fontsize: 11pt

---

```{r SETUP, include=FALSE}
# A bunch of initial set up junk

rm(list = ls())

##### knitr stuff and libraries ####
source("R-functions-ocnms.r")

# Paths for files
setwd('..')
HomeFile = getwd()
# HomeFile = "C:/Users/nick.tolimieri/Documents/GitHub/OCNMS/"

Fig_Loc = paste0(HomeFile,"/Plots/")
Data_Loc = paste0(HomeFile,"/Data/")
Results_Loc = paste0(HomeFile,"/Results/")
Other_Files = paste0(HomeFile,"/Other Files/")

# tools in use 
library(knitr)
# library(tidyr)
# library(dplyr)
library(tidyverse)
library(stringr)
library(tinytex) # for pdfs
library(RColorBrewer)
# display.brewer.all(colorblindFriendly = TRUE)
library(readxl)
# library(reprex)
library(ggpubr)
# library(ggtext)
# stats packages etc
library(vegan)
library(BiodiversityR)
# library(pracma)
# library(factoextra)

# install.packages("remotes")
# for Bray curtis with no spp in common.
# remotes::install_github("phytomosaic/ecole")
library(ecole)

# options
knitr::opts_chunk$set(
  fig.path = Fig_Loc,
  cache.path = 'cache/graphics-', 
  echo=FALSE,
  error=FALSE,
  include = TRUE,
  dev='png',
  dpi=300,
  warning=FALSE,
  #out.width = '100%',
  fig.align='center'
  
  )

# just to keep track of when I ran things.

Last_Run0 = Sys.time()
Last_Run = str_replace(Last_Run0, ":", ".")
Last_Run = str_replace(Last_Run, ":", ".")
capture.output(Last_Run, file = 'Last_Run.txt') # over-writes previous file

par(ps=10, cex=1)

# load some external functions
# source("R-functions-ocnms.r")

```

# Triggers and settings

```{r Triggers-and-Settings}

# bring in settings from previous univariate figure or set
# maintain consistency between files

settings = readRDS(paste0(Data_Loc,'settings.rds') )

min.vis = as.numeric(settings['min.vis'])
years = as.numeric(unlist(settings['years']))
site.col = data.frame(settings['site.col'])
colnames(site.col) = c('site','col')
sites = site.col$site
year.pch = data.frame(settings['year.pch'])
colnames(year.pch) = c('year','pch','col')

# Set common stuff for ordinations here ####
# Options are: NA, "sqrt", "4th-root","log"
# for 'log' the transform is actually log(x+1), need the log+1 because of zeros.

kelp.transform = 'log'
fish.transform = 'log'
invert.transform = 'log'
nperm  = 999

# Some csv files for naming later

fish_codes = data.frame(read.csv( paste0(Data_Loc,"spp_codes_fish.csv") ))
swath_codes = data.frame(read.csv( paste0(Data_Loc,"spp_codes_swath.csv") ))
kelp_codes = data.frame(read.csv( paste0(Data_Loc,"spp_codes_kelp.csv") ))

theme_nt = readRDS(paste0(Data_Loc, 'theme_nt.rds') )

```

# Import Data

Note: for Neah Bay in 2016

1. There are not algae data for the South Area only, the North.
2. There are are only two transects of invert data for the North. There are 5 transects in the South area. 

```{r import-kelp-urchins}

df_transect1 = readRDS(paste0(Data_Loc,'Data_Inverts_Kelp_transect_wide.rds')) %>% 
               filter(site %in% sites) 
              
df_transect2 = df_transect1[,c('site','year','transect','area','zone', 
                             'NERLUE','MACPYR', 'PTECAL',
                             'purple urchin','red urchin','green urchin')]

df_transect =  df_transect2 %>%
                  pivot_longer(NERLUE:PTECAL , names_to = 'kelp_spp', values_to = 'kelp_density') %>%
                    mutate(kelp_spp = factor(kelp_spp, levels=c('MACPYR','NERLUE','PTECAL'))) %>%
                  pivot_longer('purple urchin':'green urchin' , 
                               names_to = 'urchin_spp', values_to = 'urchin_density')
df_transect$kelp_spp <- as.character(df_transect$kelp_spp)

df_transect$kelp_spp[ df_transect$kelp_spp == 'NERLUE'] <- 'Nereocystis'
df_transect$kelp_spp[ df_transect$kelp_spp == 'MACPYR'] <- 'Macrocystis'
df_transect$kelp_spp[ df_transect$kelp_spp == 'PTECAL'] <- 'Pterygophora'

df_transect$kelp_spp = factor(df_transect$kelp_spp , levels = c('Macrocystis','Nereocystis','Pterygophora'))

df_transect$urchin_spp <- as.character(df_transect$urchin_spp)

df_transect$urchin_spp[df_transect$urchin_spp=='purple urchin'] <- 'Purple urchins'
df_transect$urchin_spp[df_transect$urchin_spp=='red urchin'] <- 'Red urchins'
df_transect$urchin_spp[df_transect$urchin_spp=='green urchin'] <- 'Green urchins'

df_transect$urchin_spp = factor(df_transect$urchin_spp , levels=c('Purple urchins','Red urchins','Green urchins'))

```

# BY SPECIES

## Transect level, all depths

```{r Supplement-kelp-vs-urchins-transect}
library(cowplot)
library(viridis)

df_transect$pch = substring(df_transect$year,4,4)

pl <- df_transect %>%
     #filter(zone == 5) %>%, 
  ggplot(aes(x=urchin_density,y=kelp_density ,col=site))+
  geom_point(size=2.5, pch=df_transect$pch)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste( 'Kelp stipes per 60 ', m^2)) )+
  # geom_point(size=1)+
  theme_minimal()+
  # scale_color_manual(values=color5 , pch=kelp_urchins_long$yr)+
  scale_color_manual(values=site.col$col )+
  facet_grid(urchin_spp ~ kelp_spp)+
  theme_bw()+ theme_nt + 
  theme( legend.text = element_text(size = 8),
         strip.text.x = element_text(size = 8, face='italic')) 
  # labs(x="Urchin density", y="Kelp density",col="Site", size=10)
pl



```
## By Site and Depth level 

```{r kelp-urchins-site-5&10m}

kelp_site   <- df_transect %>% group_by(year, site, zone, kelp_spp) %>% 
               summarise(kelp_mean = mean(kelp_density , na.rm = TRUE )   ) 
                # rename(mean = kelp_mean) %>%
                # rename(spp = kelp_spp)

            # 
urchin_site <- df_transect %>% group_by(year, site, zone,  urchin_spp) %>% 
               summarise(urchin_mean = mean(urchin_density, na.rm = TRUE) ) 
                # rename(mean = urchin_mean) %>%
                # rename(spp = urchin_spp)


### this file triples up the data but that is ok. 
### it is joinging all three urchins to each kelp.  
### do NOT use for calculation of means etc down line.

df_site <-  full_join(kelp_site, urchin_site, by=c('year', 'site','zone'))

dim(df_site)
dim(kelp_site)
dim(urchin_site)

pl <-  ggplot(df_site, aes(x=urchin_mean, y=kelp_mean, col=site) )+
  # geom_point(size=2.5, pch=kelp_urchins_long$yr)+
  geom_point(size=1)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
  theme_minimal()+
  # scale_color_manual(values=color5 , pch=kelp_urchins_long$yr)+
  scale_color_manual(values=site.col$col )+
  facet_grid(urchin_spp ~ kelp_spp)+
  # labs(x="Urchin density",y=("Kelp density"), col="Site")+
  theme_bw()+ theme_nt +
  theme( legend.key.size =  unit(1,"point"),
         axis.text = element_text( size = 8 ),
         panel.spacing = unit(0.5, "lines"),
         strip.text.x = element_text(face='italic')
         )
   
pl


dfx = df_site %>% filter(., kelp_spp == "Nereocystis" & urchin_spp == "Purple urchins")

purple_nero_site <- cor.test(dfx$kelp_mean,dfx$urchin_mean)

```


correlation purple vs nereo at Tatoosh r = `r purple_nero_site$estimate`, p = `r purple_nero_site$p.value`


## By Site and 5 m only

```{r kelp-urchins-site-5}

kelp_site   <- df_transect %>% group_by(year, site, zone, kelp_spp) %>% 
               summarise(kelp_mean = mean(kelp_density, na.rm = TRUE) ) 
                # rename(mean = kelp_mean) %>%
                # rename(spp = kelp_spp)
                # 
urchin_site <- df_transect %>% group_by(year, site, zone,  urchin_spp) %>% 
               summarise(urchin_mean = mean(urchin_density, na.rm = TRUE) ) 
                # rename(mean = urchin_mean) %>%
                # rename(spp = urchin_spp)


### this file triples up the data but that is ok. 
### it is joinging all three urchins to each kelp.  
### do NOT use for calculation of means etc down line.

df_site <-  left_join(kelp_site, urchin_site, by=c('year', 'site','zone'))

dim(df_site)
dim(kelp_site)
dim(urchin_site)

pl <- df_site %>% filter(zone == 5) %>%
  ggplot(., aes(x=urchin_mean, y=kelp_mean, col=site) )+
  # geom_point(size=2.5, pch=kelp_urchins_long$yr)+
  geom_point(size=1)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
  theme_minimal()+
  # labs(x="Urchin density",y=("Kelp density"),col="Site")+
  scale_color_manual(values=site.col$col )+
  facet_grid(urchin_spp ~ kelp_spp)+
  theme_bw()+ theme_nt + 
  theme( legend.key.size =  unit(0.25,"point"),
         axis.text = element_text( size = 8 ),
         axis.text.x = element_text( size = 8 ),
         strip.text.x = element_text(size = 8, face='italic'),
         panel.spacing = unit(0.5, "lines"),
         )
  labs(x="Urchin density",y=("Kelp density"),col="Site")
pl

```

## By site, averaged across depths, data = years



```{r kelp-urchins-site-mean-across-depths}

kelp_site   <- df_transect %>% group_by(year, site,  kelp_spp) %>% 
               summarise(kelp_mean = mean(kelp_density, na.rm = TRUE) ) 
                # rename(mean = kelp_mean) %>%
                # rename(spp = kelp_spp)
                # 
urchin_site <- df_transect %>% group_by(year, site,  urchin_spp) %>% 
               summarise(urchin_mean = mean(urchin_density, kelp_site = TRUE) ) 
                # rename(mean = urchin_mean) %>%
                # rename(spp = urchin_spp)


### this file triples up the data but that is ok. 
### it is joinging all three urchins to each kelp.  
### do NOT use for calculation of means etc down line.

df_site <-  full_join(kelp_site, urchin_site, by=c('year', 'site'))

df_site$pch = substring(df_site$year,4,4)

dim(df_site)
dim(kelp_site)
dim(urchin_site)

pl <- df_site %>%
  ggplot(., aes(x=urchin_mean, y=kelp_mean, col=site) )+
  geom_point(size=2.5, pch=df_site$pch)+
  # geom_point(size=1)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
  theme_minimal()+
  scale_color_manual(values=site.col$col )+
  # labs(x="Urchin density",y=("Kelp density"),col="Site")+
  facet_grid(urchin_spp ~ kelp_spp)+
  theme_bw()+ theme_nt + 
  theme( legend.key.size =  unit(0.25,"point"),
          strip.text.x = element_text(size = 8, face='italic'),
       panel.spacing = unit(0.5, "lines"),
         )
 
pl


cor.nero.data = df_site[ df_site$site == "Tatoosh Island" & 
                           df_site$kelp_spp == "Nereocystis" & 
                           df_site$urchin_spp == "purple urchin",]
cor.nero.data = na.omit(cor.nero.data)

cor.nero = cor(cor.nero.data$urchin_mean, cor.nero.data$kelp_mean )
cor.nero


cor.ptero.data = df_site[ df_site$site == "Tatoosh Island" & 
                           df_site$kelp_spp == "PterygophoraL" & 
                           df_site$urchin_spp == "purple urchin",]
cor.ptero.data = na.omit(cor.ptero.data)

cor.ptero = cor(cor.ptero.data$urchin_mean, cor.ptero.data$kelp_mean )
cor.ptero

```


This plot compared to the previous is interesting. 

1. At the transect level, there is a negative correlation between urchin density and kelp neroycystis density at Tatoosh

2. At the site level, there is a  positive correlation for *Nerocystis* (r = `r cor.nero`) and for *Pterogophora* (r = `r cor.ptero`)at Tatoosh across years. 


# TATOOSH

## By transect  

```{r tatoosh-kelp-urchins-transect, fig.width=3.5, fig.height=4}

df_tat = df_transect %>% filter( site == 'Tatoosh Island' ) %>%
                         filter( kelp_spp %in% c("Nereocystis",'Pterygophora') ) %>%
                         filter( urchin_spp == "Purple urchins")

dfx = df_tat %>% filter(kelp_spp == "Nereocystis")
dfx = na.omit(dfx)
X = dfx$urchin_density
Y = dfx$kelp_density


m1 <- nls( Y ~ a * exp(k * X), 
           start=list(a=1,k=0), 
           control=nls.control(maxiter=100,  
                               minFactor=1e-7, 
                               tol=1e-5, 
                               printEval=F, 
                               warnOnly=F) )


summary(m1)
coef(m1)
AIC(m1)


plot_tatoosh_transect <- ggplot( df_tat , aes(x=urchin_density , y=kelp_density  )) + 
  geom_point( size = 3, pch = df_tat$pch , show.legend = FALSE ) +
  facet_wrap(facets = df_tat$kelp_spp, ncol=1)+
  xlim(0,NA)+
  ylim(0,NA)+
  xlab( expression(paste('Purple urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
  geom_smooth( # subset(df_tat, df_tat$kelp_spp == "Nereocystis"),
               # aes(x = kelp_density, y = urchin_density),
               method = "nls",
               formula=y ~ a * exp(k * x),
               method.args = list(start=c(a=1,k=0) ),
               se=FALSE) + 
  theme_bw() + theme_nt + 
  theme( legend.key.size =  unit(0.25,"point"),
         legend.position = c(0.7, 0.1),
          strip.text.x = element_text(size = 8, face='italic'),
          panel.spacing = unit(0.5, "lines"),
         )

plot_tatoosh_transect

## 


```

## Across years

```{r tatoosh-kelp-urchins-site, fig.width=3.5, fig.height=4}

df_tat = df_site %>% filter( site == 'Tatoosh Island' ) %>%
                     filter( kelp_spp %in% c("Nereocystis",'Pterygophora') ) %>%
                     filter( urchin_spp == "Purple urchins")

df_tat$pch2 = as.character(df_tat$year)

dfn = df_tat %>% filter(kelp_spp == 'Nereocystis')
nereo.cor = cor.test( dfn$kelp_mean , dfn$urchin_mean )
dfp = df_tat %>% filter(kelp_spp == 'Pterygophora')
ptero.cor = cor.test( dfp$kelp_mean , dfp$urchin_mean )

x = c(5,5)
y = c(1.5,1,5)
y2 = c(1,1)
kelp_spp = c("Nereocystis", "Pterygophora")
r = c(nereo.cor$estimate, ptero.cor$estimate)
p = c(nereo.cor$p.value, ptero.cor$p.value)

rstatement = paste0('r = ', round(as.numeric(r),2))
pstatement = paste0('p = ', round(as.numeric(p),3))

z = data.frame(kelp_spp, rstatement, pstatement)


plot_tatoosh_year <- ggplot( df_tat , aes(x=urchin_mean  ,y=kelp_mean)) + 
  #geom_point( size = 4, pch = df_tat$pch ) +
  geom_text(data=df_tat , aes(x=urchin_mean, y=kelp_mean,label=pch2), size = 3)+
  # facet_wrap(facets = df_tat$kelp_spp, ncol=1)+
  facet_grid(rows=vars(kelp_spp) )+
  # annotate(data=z , geom="text", x=5, y=1.5, label= rstatement, hjust = 0,color="black", size = 3) +
  # annotate(data=z , geom="text", x=5, y=1.0, label= pstatement, hjust = 0,color="black", size = 3) +
  geom_text(data = z, x = 5, y = 1.0, size = 3, aes(label = rstatement), hjust = 0 )+
  geom_text(data = z, x = 5, y = 0.5, size = 3, aes(label = pstatement), hjust = 0 )+
  xlim(0,NA)+
  ylim(0,NA)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
  theme_bw() + theme_nt + 
  theme(  legend.title = element_blank(),
          strip.text.y = element_text(size = 8, face = 'italic'),
          panel.spacing = unit(0.5, "lines"),
         )

plot_tatoosh_year 
     
  


```



## Total kelp



```{r total-kelp-urchins-site-year, fig.width=4.5, fig.height=2.5}

df_tot <- df_transect %>% 
              group_by(year,site, area, zone, transect) %>%
              summarise( kelp_sum = sum(kelp_density, na.rm = TRUE),
                         urchin_sum = sum(urchin_density, na.rm = TRUE) ) %>%
              group_by(year,site) %>%
              summarise( kelp = mean(kelp_sum, na.rm=TRUE) , urchins = mean(urchin_sum, na.rm=TRUE)) %>%
              mutate( pch = substring(year,4,4))
              


plot_site_tot <- ggplot(df_tot , aes(x = urchins, y = kelp ,  color = site)) +
  geom_point(size = 3 , pch = df_tot$pch) +
  scale_color_manual(values=site.col$col)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
  xlim(0,NA) + 
  ylim(0,NA)+
  theme_bw() + theme_nt + 
  theme(  legend.title = element_blank(),
          legend.key.size =  unit(0.25,"point"),
          #legend.position = c(0.25, 0.8),
          strip.text.x = element_text(size = 8, face = 'italic'),
          panel.spacing = unit(0.5, "lines"),
         )



plot_site_tot
```


# Coastwide urchin -  kelp patterns 

## All kelp: Nereo + Macro + Ptero

```{r total-kelp-urchins, fig.height=2.5 , fig.width=3.5}

df_tot <- df_transect %>% 
              group_by(year,site, area, zone, transect) %>%
              summarise( kelp_sum = sum(kelp_density, na.rm = TRUE),
                         urchin_sum = sum(urchin_density, na.rm = TRUE) ) %>%
              group_by(year,site) %>%
              summarise( kelp = mean(kelp_sum, na.rm=TRUE) , urchins = mean(urchin_sum, na.rm=TRUE)) %>%
              mutate( pch = substring(year,4,4)) %>%
              group_by(year) %>%
              summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))
              
KU_year = cor.test(df_tot$Kelp , df_tot$Urchins)
rstatement  = paste0('r = ' , round(KU_year$estimate,2))
pstatement  = paste0('p = ' , round(KU_year$p.value,2))

plot_tot <- ggplot(df_tot , aes(x = Urchins , y = Kelp )) +
  # geom_point(size = 5 , show.legend = FALSE) +
  geom_text(aes(x=Urchins,y=Kelp, label=year), size = 3)+
  # scale_color_manual(values=site.col$col)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
  xlim(0,NA) + 
  ylim(0,NA)+
  annotate(geom="text", x=4, y=2.5, label= rstatement, position=4, hjust = 0,color="black", size = 3) +
  annotate(geom="text", x=4, y=1.5, label= pstatement, position=4, hjust = 0,color="black", size = 3) +
  theme_bw() + theme_nt + 
  theme(  legend.position = c(0.1, 0.8),
          strip.text.x = element_text(size = 8, face = 'italic'),
          panel.spacing = unit(0.5, "lines"),
         )






plot_tot
```

cor:  r = `r KU_year$estimate`; p = `r KU_year$p.value`

## Canopy only 



```{r total-canopy-kelp-urchins, fig.height=2.5 , fig.width=3.5}

df_canopy <- df_transect %>% filter(., kelp_spp != 'Pterygophora') %>%
              group_by(year,site, area, zone, transect) %>%
              summarise( kelp_sum = sum(kelp_density, na.rm = TRUE),
                         urchin_sum = sum(urchin_density, na.rm = TRUE) ) %>%
              group_by(year,site) %>%
              summarise( kelp = mean(kelp_sum, na.rm=TRUE) , urchins = mean(urchin_sum, na.rm=TRUE)) %>%
              mutate( pch = substring(year,4,4)) %>%
              group_by(year) %>%
              summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))
              
canopy_year = cor.test(df_canopy$Kelp , df_canopy$Urchins)
rstatement  = paste0('r = ' , round(canopy_year$estimate,2))
pstatement  = paste0('p = ' , round(canopy_year$p.value,2))

plot_canopy <- ggplot(df_canopy , aes( x = Urchins, y = Kelp )) +
  # geom_point(size = 5 , show.legend = FALSE) +
  geom_text(aes(x=Urchins,y=Kelp,label=year), size = 3)+
  # scale_color_manual(values=site.col$col)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste('Canopy kelp stipes per 60 ', m^2)) )+
  xlim(0,NA) + 
  ylim(0,NA)+
  annotate(geom="text", x=3, y=2.25, label= rstatement, hjust = 0, color="black", size = 3) +
  annotate(geom="text", x=3, y=1.5, label= pstatement, hjust = 0, color="black", size = 3) +
  theme_bw() + theme_nt + 
  theme(  legend.position = c(0.1, 0.8),
          strip.text.x = element_text(size = 8, face = 'italic'),
          panel.spacing = unit(0.5, "lines"),
         )





plot_canopy
```

cor r = `r canopy_year$estimate`


## Nero only

```{r total-nero-kelp-urchins, fig.height=2.5 , fig.width=3.5}

df_nereo <- df_transect %>% filter( kelp_spp == 'Nereocystis') %>%
              group_by(year,site, area, zone, transect) %>%
              summarise( kelp_sum = sum(kelp_density, na.rm = TRUE),
                         urchin_sum = sum(urchin_density, na.rm = TRUE) ) %>%
              group_by(year,site) %>%
              summarise( kelp = mean(kelp_sum, na.rm=TRUE) , urchins = mean(urchin_sum, na.rm=TRUE)) %>%
              mutate( pch = substring(year,4,4)) %>%
              group_by(year) %>%
              summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))
              
nereo_year = cor.test(df_nereo$Kelp ,df_nereo$Urchins)
rstatement  = paste0('r = ' , round(nereo_year$estimate,2))
pstatement  = paste0('p = ' , round(nereo_year$p.value,2))

plot_nereo <- ggplot(df_nereo , aes(x = Urchins, y = Kelp )) +
  # geom_point(size = 5 , show.legend = FALSE) +
  geom_text(aes(x=Urchins,y=Kelp, label=year), size = 3)+
  # scale_color_manual(values=site.col$col)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste(italic('Nereocystis'), ' stipes per 60 ', m^2)) )+
  xlim(0,NA) + 
  ylim(0,NA)+
  annotate(geom="text", x=1.25, y=1.0, label= rstatement, hjust = 0, color="black", size = 3) +
  annotate(geom="text", x=1.25, y=0.5, label= pstatement, hjust = 0, color="black", size = 3) +
  theme_bw() + theme_nt + 
  theme(  legend.position = c(0.1, 0.8),
          strip.text.x = element_text(size = 8, face = 'italic'),
          panel.spacing = unit(0.5, "lines"),
         )






plot_nereo
```

## Macro only 

```{r total-macro-kelp-urchins, fig.height=2.5 , fig.width=3.5}

df_macro <- df_transect %>% filter( kelp_spp == 'Macrocystis') %>%
              group_by(year,site, area, zone, transect) %>%
              summarise( kelp_sum = sum(kelp_density, na.rm = TRUE),
                         urchin_sum = sum(urchin_density, na.rm = TRUE) ) %>%
              group_by(year,site) %>%
              summarise( kelp = mean(kelp_sum, na.rm=TRUE) , urchins = mean(urchin_sum, na.rm=TRUE)) %>%
              mutate( pch = substring(year,4,4)) %>%
              group_by(year) %>%
              summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))
              
macro_year = cor.test(df_macro$Kelp ,df_macro$Urchins)
rstatement  = paste0('r = ' , round(macro_year$estimate,2))
pstatement  = paste0('p = ' , round(macro_year$p.value,2))

plot_macro <- ggplot(df_macro , aes(x = Urchins , y = Kelp)) +
  # geom_point(size = 5 , show.legend = FALSE) +
  geom_text(aes(x=Urchins,y=Kelp, label=year), size = 3)+
  # scale_color_manual(values=site.col$col)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste(italic('Macrocystis'), ' stipes per 60 ', m^2)) )+
  xlim(0,NA) + 
  ylim(0,NA)+
  annotate(geom="text", x=1.5, y=0.5, label= rstatement, hjust = 0, color="black", size = 3) +
  annotate(geom="text", x=1.5, y=0.25, label= pstatement, hjust = 0, color="black", size = 3) +
  theme_bw() + theme_nt + 
  theme(  legend.position = c(0.1, 0.8),
          strip.text.x = element_text(size = 8, face = 'italic'),
          panel.spacing = unit(0.5, "lines"),
         )






plot_macro
```
## Ptero only

```{r total-ptero-kelp-urchins, fig.height=2.5 , fig.width=3.5}

df_ptero <- df_transect %>% filter( kelp_spp == 'Pterygophora') %>%
              group_by(year,site, area, zone, transect) %>%
              summarise( kelp_sum = sum(kelp_density, na.rm = TRUE),
                         urchin_sum = sum(urchin_density, na.rm = TRUE) ) %>%
              group_by(year,site) %>%
              summarise( kelp = mean(kelp_sum, na.rm=TRUE) , urchins = mean(urchin_sum, na.rm=TRUE)) %>%
              mutate( pch = substring(year,4,4)) %>%
              group_by(year) %>%
              summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))

ptero_year = cor.test(df_ptero$Kelp ,df_ptero$Urchins)              
rstatement  = paste0('r = ' , round(ptero_year$estimate,2))
pstatement  = paste0('p = ' , round(ptero_year$p.value,2))

plot_ptero <- ggplot(df_ptero , aes(x = Urchins, y = Kelp )) +
  # geom_point(size = 5 , show.legend = FALSE) +
  geom_text(aes(x=Urchins,y=Kelp, label=year), size = 3)+
  # scale_color_manual(values=site.col$col)+
  xlab( expression(paste('Urchins per 60 ', m^2) ) )+
  ylab( expression(paste(italic('Pterygophora'), ' stipes per 60 ', m^2)) )+
  xlim(0,NA) + 
  ylim(0,NA)+
  annotate(geom="text", x=1.5, y=1.50, label= rstatement, hjust = 0, color="black", size = 3) +
  annotate(geom="text", x=1.5, y=1.15, label= pstatement, hjust = 0, color="black", size = 3) +
  theme_bw() + theme_nt + 
  theme(  legend.position = c(0.1, 0.8),
          strip.text.x = element_text(size = 8, face = 'italic'),
          panel.spacing = unit(0.5, "lines"),
         )




ptero_year = cor.test(df_ptero$Kelp ,df_ptero$Urchins)

plot_ptero
```

cor r = `r ptero_year$estimate`;  = `r ptero_year$p.value`

# Tatoosh only

```{r tatoosh-transect-level , fig.height=2.5 , fig.width=3.5}

dfx <-df_transect %>% filter(., site == 'Tatoosh Island') %>% 
                      group_by(site,year,transect,area,zone) %>%
                      summarize( kelp = sum(kelp_density) , urchins = sum(urchin_density) )
dfx$pch = as.character(substring(dfx$year,4,4))

dfy = na.omit(dfx)
Y = dfy$kelp
X = dfy$urchins

m1 <- nls( Y ~ a * exp(k * X),
           start=list(a=1,k=0),
           control=nls.control(maxiter=100,
                               minFactor=1e-7,
                               tol=1e-5,
                               printEval=F,
                               warnOnly=F))
summary(m1)
coef(m1)
AIC(m1)

                  
p.tat = ggplot(dfx , aes(x = urchins, y = kelp)) + 
        # geom_point() +
        geom_text(data=dfx , aes(x=urchins,y=kelp, label=pch), size = 4)+ 
        # scale_y_continuous(trans="log") +
        geom_smooth( method = "nls",
                      formula=y ~ a * exp(k * x),
                      method.args = list(start=c(a=1,k=0) ),
                      se=FALSE) + 
        xlab( expression(paste('Urchins per 60 ', m^2) ) )+
        ylab( expression(paste('Kelp stipes per 60 ', m^2)) )+
        theme_bw() + theme_nt

p.tat
```


```{r tatoosh-transect-level-nereo , fig.height=2.5 , fig.width=3.5}

dfx <-df_transect %>% filter(., site == 'Tatoosh Island') %>% filter(kelp_spp == 'Nereocystis') %>%
                      group_by(site,year,transect,area,zone) %>%
                      summarize( kelp = sum(kelp_density) , urchins = sum(urchin_density) )
dfx$pch = as.character(substring(dfx$year,4,4))

dfy = na.omit(dfx)

X = dfy$urchins
Y = dfy$kelp


m1 <- nls( Y ~ a * exp(k * X),
           start=list(a=1,k=0),
           control=nls.control(maxiter=100,
                               minFactor=1e-7,
                               tol=1e-5,
                               printEval=F,
                               warnOnly=F))
#summary(m1)
#coef(m1)
#AIC(m1)
m2 = lm( log(Y+0.01) ~ X ) 
s2 = summary(m2)

p2 = exp(predict(m2))-0.01
pr = data.frame(cbind(X,p2))    

rstatement = bquote(italic(r)^2 == .(format(s2$r.squared, digits = 2)))
pstatement = paste0('p = ' , round(s2$coefficients[2,4],3) )

# expression(paste("X comes from ",italic("normal distribution")))
# epression(paste(italic('Pterygophra'), 'stipe density'))

nereo.tat = ggplot(dfx , aes(x = urchins, y = kelp)) + 
        geom_line(data=pr, aes(x=X,y=p2), color='red') +
        geom_text(data=dfx , aes(x=urchins, y=kelp, label=pch), size = 4)+ 
        xlab( expression(paste('Urchins per 60 ', m^2) ) )+
        ylab( expression(paste(italic('Nereocystis'), ' stipes per 60 ', m^2)) )+
        annotate(geom="text", x=15, y=30, label = rstatement, hjust = 0, color="black", size = 3) +
        annotate(geom="text", x=15, y=25, label= pstatement, hjust = 0, color="black", size = 3) +
        theme_bw() + theme_nt

nereo.tat

```
```{r tatoosh-transect-level-ptery , fig.height=2.5 , fig.width=3.5}

dfx <-df_transect %>% filter(., site == 'Tatoosh Island') %>% filter(kelp_spp == 'Pterygophora') %>%
                      group_by(site,year,transect,area,zone) %>%
                      summarize( kelp = sum(kelp_density) , urchins = sum(urchin_density) )
dfx$pch = as.character(substring(dfx$year,4,4))

dfy = na.omit(dfx)

X = dfy$urchins
Y = dfy$kelp


m1 <- nls( Y ~ a * exp(k * X),
           start=list(a=1,k=0),
           control=nls.control(maxiter=100,
                               minFactor=1e-7,
                               tol=1e-5,
                               printEval=F,
                               warnOnly=F))
summary(m1)
coef(m1)
AIC(m1)
m2 = lm( log(Y+0.01) ~ X ) 
s2 = summary(m2)

p2 = exp(predict(m2))-0.01
pr = data.frame(cbind(X,p2))    

rstatement = bquote(italic(r)^2 == .(format(s2$r.squared, digits = 2)))
pstatement = paste0('p = ' , round(s2$coefficients[2,4],3) )


ptery.tat = ggplot(dfx , aes(x = urchins, y = kelp)) + 
        geom_line(data=pr, aes(x=X,y=p2), color='red') +
        geom_text(data=dfx , aes(x=urchins, y=kelp, label=pch), size = 4)+ 
        xlab( expression(paste('Urchins per 60 ', m^2) ) )+
        ylab( expression(paste(italic('Pterygophora'), ' stipes per 60 ', m^2)) )+
        annotate(geom="text", x=15, y=30, label = rstatement, hjust = 0, color="black", size = 3) +
        annotate(geom="text", x=15, y=27, label= pstatement, hjust = 0, color="black", size = 3) +
        theme_bw() + theme_nt

ptery.tat

```

# Additional supplement figure


```{r Supplement-kelp-vs-urchins-year, fig.height=6, fig.width=6}


library(ggpubr)

ggarrange( plot_tot, plot_canopy, plot_macro, plot_nereo,  plot_ptero,
           labels = c('a)','b)','c)','d)', 'e)'), 
           font.label = list(face='plain', size = 10) ,
           hjust = -5.5,
           vjust = 2,
           align = 'v',
           nrow = 3, ncol=2)




```

# Figure 6 output for main MS


```{r Figure-6, fig.height=4, fig.width=3.5}




ggarrange(plot_ptero, nereo.tat,
           labels = c('a)','b)'), 
           font.label = list(face='plain', size = 10),
           hjust = -5.5,
           vjust = 2,
           align = 'v',
           nrow = 2, ncol=1)




```

# Combined Figure 6


```{r Figure-6-Combined, fig.height=5, fig.width=3.5}

df_tot$group = 'Total'
df_canopy$group = 'Canopy'
df_macro$group = "Macrocystis"
df_nereo$group = "Nereocystis"
df_ptero$group = "Pterygophora"

df = rbind(df_tot,df_canopy,df_macro,df_nereo,df_ptero)
df$group = factor(df$group , levels = c('Total','Canopy','Macrocystis',"Nereocystis","Pterygophora"))

# Total kelp stipes

dfx = na.omit(df %>% filter(., group == "Total"))
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement1 = paste0('r = ' , round(R$estimate,2) )
pstatement1 = paste0('p = ' , round(R$p.value,2) )



# Pterygophora
dfx = na.omit(df %>% filter(., group == "Pterygophora"))
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement2 = paste0('r = ' , round(R$estimate,2) )
pstatement2 = paste0('p = ' , round(R$p.value,2) )

stats.x = 4.5



plot1 <- ggplot( df , aes(x = Urchins, y = Kelp, color=group)) +
          geom_point(size = 2)+
          scale_color_manual(values = site.col$col[c(4,5,1,2,3)],
                             labels = c('Total','Canopy',
                                        bquote(italic('Macrocystis')),
                                        bquote(italic('Nereocystis')),
                                        bquote(italic('Pterygophora'))),
                             )+
          xlab( expression(paste('Urchins per 60 ', m^2) ) )+
          ylab( expression(paste('Stipes per 60 ', m^2)) )+
          annotate(geom="text", x=stats.x, y=5.0, 
                   label = "Total stipes",hjust = 0, color=site.col$col[4], size = 3) +
          annotate(geom="text", x=stats.x, y=4.3, 
                   label = rstatement1, hjust = 0, color=site.col$col[4], size = 3) +
          annotate(geom="text", x=stats.x, y=3.8, 
                   label = pstatement1, hjust = 0, color=site.col$col[4], size = 3) +
          
          annotate(geom="text", x=stats.x, y=3.0, 
                   label = expression(italic("Pterygophora")), 
                   hjust = 0, color=site.col$col[3], size = 3) +
          annotate(geom="text", x=stats.x, y=2.35, 
                   label = rstatement2, hjust = 0, color=site.col$col[3], size = 3) +
          annotate(geom="text", x=stats.x, y=1.8, 
                   label = pstatement2, hjust = 0, color=site.col$col[3], size = 3) +
          theme_bw() + theme_nt + theme( legend.position = c(0.25, 0.85),
                                         legend.key.size = unit(0.5,'lines'))
  


ggarrange(plot1, nereo.tat,
           labels = c('a)','b)'), 
           font.label = list(face='plain', size = 10),
           hjust = -5.5,
           vjust = 2,
           align = 'v',
           nrow = 2, ncol=1)



```






























