---
title: "Kelp vs Urchins - the differenced version"
author: Nick Toimieri
date: sys.date()
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
fontsize: 11pt

---
# Set up

```{r SETUP, include=FALSE}
# A bunch of initial set up junk

rm(list = ls())

##### knitr stuff and libraries ####
source("R-functions-ocnms.r")

# Paths for files

HomeFile = "C:/Users/nick.tolimieri/Documents/GitHub/OCNMS/Flagstone paper/"

Fig_Loc = paste0(HomeFile,"/Plots/")
Data_Loc = paste0(HomeFile,"/Data/")
Results_Loc = paste0(HomeFile,"/Results/")
Other_Files = paste0(HomeFile,"/Other Files/")

# tools in use 
library(knitr)
library(tidyverse)
library(stringr)
library(tinytex) # for pdfs
library(RColorBrewer)
# display.brewer.all(colorblindFriendly = TRUE)
library(readxl)
# library(reprex)
library(ggplot2)
library(ggpubr)
# library(ggtext)
# stats packages etc
# library(vegan)
# library(BiodiversityR)
# library(pracma)
# library(factoextra)

# install.packages("remotes")
# for Bray curtis with no spp in common.
# remotes::install_github("phytomosaic/ecole")


# options
knitr::opts_chunk$set(
  fig.path = Fig_Loc,
  cache.path = 'cache/graphics-', 
  echo=FALSE,
  error=FALSE,
  include = TRUE,
  dev='png',
  dpi=300,
  warning=FALSE,
  #out.width = '100%',
  fig.align='center'
  
  )

# just to keep track of when I ran things.


par(ps=10, cex=1)

# load some external functions
# source("R-functions-ocnms.r")

```

# Triggers and settings

```{r Triggers-and-Settings}

# bring in settings from previous univariate figure or set
# maintain consistency between files

settings = readRDS(paste0(Data_Loc,'settings.rds') )

min.vis = as.numeric(settings['min.vis'])
years = as.numeric(unlist(settings['years']))
site.col = data.frame(settings['site.col'])
colnames(site.col) = c('site','col')
sites = site.col$site
year.pch = data.frame(settings['year.pch'])
colnames(year.pch) = c('year','pch','col')

# Set common stuff for ordinations here ####
# Options are: NA, "sqrt", "4th-root","log"
# for 'log' the transform is actually log(x+1), need the log+1 because of zeros.

kelp.transform = 'log'
fish.transform = 'log'
invert.transform = 'log'
nperm  = 999

# Some csv files for naming later

fish_codes = data.frame(read.csv( paste0(Data_Loc,"spp_codes_fish.csv") ))
swath_codes = data.frame(read.csv( paste0(Data_Loc,"spp_codes_swath.csv") ))
kelp_codes = data.frame(read.csv( paste0(Data_Loc,"spp_codes_kelp.csv") ))

theme_nt = readRDS(paste0(Data_Loc, 'theme_nt.rds') )

```

# Bring in data

Note: for Neah Bay in 2016

1. There are not algae data for the South Area only, the North.
2. There are are only two transects of invert data for the North. There are 5 transects in the South area. 

```{r import-kelp-urchins}

df_transect1 = readRDS(paste0(Data_Loc,'Data_Inverts_Kelp_transect_wide.rds')) %>% 
               mutate(Other = c(AGAFIM+ALAMAR+COSCOS+CYMTRI+DESSPP+LAMSET+PLEGAR+SACGRO+SACLAT)) %>%
               filter(site %in% sites) 

df_transect2 = df_transect1[,c('site','year','transect','area','zone', 
                             'NERLUE','MACPYR', 'PTECAL','Other',
                             'purple urchin','red urchin','green urchin')]

df_transect =  df_transect2 %>%
                  pivot_longer(NERLUE:Other , names_to = 'kelp_spp', values_to = 'kelp_density') %>%
                    mutate(kelp_spp = factor(kelp_spp, levels=c('MACPYR','NERLUE','PTECAL','Other'))) %>%
                  pivot_longer('purple urchin':'green urchin' , 
                               names_to = 'urchin_spp', values_to = 'urchin_density')
df_transect$kelp_spp <- as.character(df_transect$kelp_spp)

df_transect$kelp_spp[ df_transect$kelp_spp == 'NERLUE'] <- 'Nereocystis'
df_transect$kelp_spp[ df_transect$kelp_spp == 'MACPYR'] <- 'Macrocystis'
df_transect$kelp_spp[ df_transect$kelp_spp == 'PTECAL'] <- 'Pterygophora'

df_transect$kelp_spp = factor(df_transect$kelp_spp , levels = 
                                c('Macrocystis','Nereocystis','Pterygophora','Other'))

df_transect$urchin_spp <- as.character(df_transect$urchin_spp)

df_transect$urchin_spp[df_transect$urchin_spp=='purple urchin'] <- 'Purple urchins'
df_transect$urchin_spp[df_transect$urchin_spp=='red urchin'] <- 'Red urchins'
df_transect$urchin_spp[df_transect$urchin_spp=='green urchin'] <- 'Green urchins'

df_transect$urchin_spp = factor(df_transect$urchin_spp , levels=c('Purple urchins','Red urchins','Green urchins'))

```

# Year x Kelp - average across sites

```{r Figure-5-prep-diff, fig.height=6, fig.width=6}
df_tot <- df_transect %>% 
  group_by(year,site, area, zone, transect) %>%
  summarise( kelp_mean = mean(kelp_density, na.rm = TRUE),
             urchin_mean = mean(urchin_density, na.rm = TRUE) ) %>%
  group_by(year,site) %>%
  summarise( kelp = mean(kelp_mean, na.rm=TRUE) , 
             urchins = mean(urchin_mean, na.rm=TRUE)) %>%
  mutate( pch = substring(year,4,4)) %>%
  group_by(year) %>%
  summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))

df_canopy <- df_transect %>% filter(., kelp_spp != 'Pterygophora') %>%
  group_by(year,site, area, zone, transect) %>%
  summarise( kelp_mean = mean(kelp_density, na.rm = TRUE),
             urchin_mean = mean(urchin_density, na.rm = TRUE) ) %>%
  group_by(year,site) %>%
  summarise( kelp = mean(kelp_mean, na.rm=TRUE) , 
             urchins = mean(urchin_mean, na.rm=TRUE)) %>%
  mutate( pch = substring(year,4,4)) %>%
  group_by(year) %>%
  summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))

df_nereo <- df_transect %>% filter( kelp_spp == 'Nereocystis') %>%
  group_by(year,site, area, zone, transect) %>%
  summarise( kelp_mn = mean(kelp_density, na.rm = TRUE),
             urchin_mn = mean(urchin_density, na.rm = TRUE)) %>%
  group_by(year,site) %>%
  summarise( kelp = mean(kelp_mn, na.rm=TRUE) , 
             urchins = mean(urchin_mn, na.rm=TRUE)) %>%
  mutate( pch = substring(year,4,4)) %>%
  group_by(year) %>%
  summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))

df_macro <- df_transect %>% filter( kelp_spp == 'Macrocystis') %>%
  group_by(year,site, area, zone, transect) %>%
  summarise( kelp_mn = mean(kelp_density, na.rm = TRUE),
             urchin_mn = mean(urchin_density, na.rm = TRUE)) %>%
  group_by(year,site) %>%
  summarise( kelp = mean(kelp_mn, na.rm=TRUE) , 
             urchins = mean(urchin_mn, na.rm=TRUE)) %>%
  mutate( pch = substring(year,4,4)) %>%
  group_by(year) %>%
  summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))

df_ptero <- df_transect %>% filter( kelp_spp == 'Pterygophora') %>%
  group_by(year,site, area, zone, transect) %>%
  summarise( kelp_mn = mean(kelp_density, na.rm = TRUE),
             urchin_mn = mean(urchin_density, na.rm = TRUE)) %>%
  group_by(year,site) %>%
  summarise( kelp = mean(kelp_mn, na.rm=TRUE) , 
             urchins = mean(urchin_mn, na.rm=TRUE)) %>%
  mutate( pch = substring(year,4,4)) %>%
  group_by(year) %>%
  summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))

df_other <- df_transect %>% filter( kelp_spp == 'Other') %>%
              group_by(year,site, area, zone, transect) %>%
              summarise( kelp_mn = mean(kelp_density, na.rm = TRUE),
                         urchin_mn = mean(urchin_density, na.rm = TRUE)) %>%
              group_by(year,site) %>%
              summarise( kelp = mean(kelp_mn, na.rm=TRUE) , 
                         urchins = mean(urchin_mn, na.rm=TRUE)) %>%
              mutate( pch = substring(year,4,4)) %>%
              group_by(year) %>%
              summarise(Kelp = mean(kelp, na.rm=TRUE) , Urchins = mean(urchins, na.rm=TRUE))




df_tot$group = 'Total'
df_canopy$group = 'Canopy'
df_macro$group = "Macrocystis"
df_nereo$group = "Nereocystis"
df_ptero$group = "Pterygophora"
df_other$group = "Other"

df = rbind(df_tot,df_canopy,df_macro,df_nereo,df_ptero, df_other)
df$group = factor(df$group , levels = c('Total','Canopy','Macrocystis',"Nereocystis","Pterygophora","Other"))


# Differencing Here - none previously

df_diff = df %>% group_by(group) %>%
  mutate( Kelp = Kelp - lag(Kelp, default = Kelp[1]),
          Urchins = Urchins - lag(Urchins, default = Urchins[1])) %>%  filter(year>2015) %>%
  filter(group %in% c('Macrocystis','Nereocystis','Pterygophora','Other', 'Total') )

# Total kelp stipes
dfx = na.omit(df_diff %>% filter(., group == "Total"))
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement0 = paste0('r = ' , round(R$estimate,2) )
pstatement0 = paste0('p = ' , round(R$p.value,2) )

# Macro
dfx = na.omit(df_diff %>% filter(., group == "Macrocystis"))
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement1 = paste0('r = ' , round(R$estimate,2) )
pstatement1 = paste0('p = ' , round(R$p.value,2) )

# Nereo
dfx = na.omit(df_diff %>% filter(., group == "Nereocystis"))
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement2 = bquote(italic(r)^2 == .(format(R$estimate^2, digits = 2)))
pstatement2 = paste0('p = ' , round(R$p.value,2) )

# Pterygophora
dfx = na.omit(df_diff %>% filter(., group == "Pterygophora"))
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement3 = bquote(italic(r)^2 == .(format(R$estimate^2, digits = 2)))
pstatement3 = paste0('p = ' , round(R$p.value,2) )

# other
dfx = na.omit(df_diff %>% filter(., group == "Other"))
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement4 = bquote(italic(r)^2 == .(format(R$estimate^2, digits = 2)))
pstatement4 = paste0('p = ' , round(R$p.value,2) )

stats.x = -0.2



df_diff = df %>% group_by(group) %>%
  mutate( Kelp = Kelp - lag(Kelp, default = Kelp[1]),
          Urchins = Urchins - lag(Urchins, default = Urchins[1])) %>%  filter(year>2015) %>%
  filter(group %in% c('Macrocystis','Nereocystis','Pterygophora','Other') )

# plot
plot1 <- ggplot( df_diff, aes(x = Urchins, y = Kelp, color=group)) +
  geom_point(size = 2) +
  scale_color_manual(values = site.col$col[c(1,2,3,4)],
                     labels = c(bquote(italic('Macrocystis')),
                              bquote(italic('Nereocystis')),
                              bquote(italic('Pterygophora')),
                              'Other'))+
  xlab( expression(paste('Change in urchins per ', m^2) ) )+
  ylab( expression(paste('Change in stipes per ', m^2) ) )+
  # geom_smooth(data = filter(df_diff,group =='Pterygophora' ),method='lm', se=FALSE) +
   # annotate(geom="text", x=stats.x, y=0.9, 
   #          label = expression(italic("Pterygophora")), hjust = 0, color=site.col$col[3], size = 3) +
   # annotate(geom="text", x=stats.x, y=0.7, 
   #          label = rstatement3, hjust = 0, color=site.col$col[3], size = 3) +
   # annotate(geom="text", x=stats.x, y=0.5, 
   #          label = pstatement3, hjust = 0, color=site.col$col[3], size = 3) +
           theme_bw() + theme_nt 

plot1


```
# Year x Site, separate plots by kelp

## Macro

```{r macro-vs-urchins-site-year-diff, fig.height=2.5, fig.width=3.5}

df = df_transect %>% group_by(site, year, zone, area, transect, kelp_spp) %>%
                      summarise(urchins = mean(urchin_density), kelp = mean(kelp_density) ) %>% 
                      group_by(site,year,kelp_spp) %>%
                      summarise(Urchins = mean(urchins), Kelp = mean(kelp) ) 

df$pch = as.character(substring(df$year,4,4))


df_diff = df %>% group_by(kelp_spp, site) %>%
  mutate( Kelp = Kelp - lag(Kelp, default = Kelp[1]),
          Urchins = Urchins - lag(Urchins, default = Urchins[1])) %>%  
  filter(year>2015,kelp_spp == "Macrocystis")

macro1 <- ggplot(df_diff, aes( x=Urchins , y=Kelp, color=site)) +
         geom_point(size=2.5, pch=df_diff$pch) +
         scale_color_manual(values=site.col$col )+
         xlim(0,NA)+
         ylim(0,NA)+
         xlab( expression(paste(Delta, ' urchins per ', m^2) ) )+
         ylab( expression(paste(Delta, italic(' Macro'), ' stipes per ', m^2)) )+
         theme_bw() + theme_nt 

macro1

```
## Nereo

```{r nereo-vs-urchins-site-year-diff, fig.height=2.5, fig.width=3.5}

df = df_transect %>% group_by(site, year, zone, area, transect, kelp_spp) %>%
                      summarise(urchins = mean(urchin_density), kelp = mean(kelp_density) ) %>% 
                      group_by(site,year,kelp_spp) %>%
                      summarise(Urchins = mean(urchins), Kelp = mean(kelp) )

df$pch = as.character(substring(df$year,4,4))

df_diff = df %>% group_by(kelp_spp, site) %>%
  mutate( Kelp = Kelp - lag(Kelp, default = Kelp[1]),
          Urchins = Urchins - lag(Urchins, default = Urchins[1])) %>%  
  filter(year>2015, kelp_spp=="Nereocystis")

nereo = df_diff

lmp = lm( Kelp ~ Urchins, data = nereo[nereo$site =='Tatoosh Island',])
nereo.s.tat = summary(lmp)
lmp = lm( Kelp ~ Urchins, data = nereo[nereo$site =='Destruction Island',])
nereo.s.des = summary(lmp)

# correlation?
dfx = na.omit(df_diff)
X = dfx$Urchins
Y = dfx$Kelp
R = cor.test(X,Y)
rstatement4 = bquote(italic(r)^2 == .(format(R$estimate^2, digits = 2)))
pstatement4 = paste0('p = ' , round(R$p.value,2) )

# no obvious correlations...

nereo1 <- ggplot(df_diff, aes( x=Urchins , y=Kelp, color=site)) +
         geom_point(size=2.5, pch=df_diff$pch) +
         scale_color_manual(values=site.col$col )+
         # geom_smooth(data = filter(nereo,site =='Tatoosh Island' ), method='lm', se=FALSE) +
         # geom_smooth(data = filter(nereo,site =='Destruction Island' ), method='lm', se=FALSE) +
         xlab( expression(paste(Delta, ' urchins per ', m^2) ) )+
         ylab( expression(paste(Delta, italic(' Nereo'), ' stipes per ', m^2)) )+
         theme_bw() + theme_nt 

nereo1

```

## Ptery
```{r ptery-vs-urchins-site-year-diff, fig.height=2.5, fig.width=3.5}

df = df_transect %>% group_by(site, year, zone, area, transect, kelp_spp) %>%
                      summarise(urchins = mean(urchin_density), kelp = mean(kelp_density) ) %>% 
                      group_by(site,year,kelp_spp) %>%
                      summarise(Urchins = mean(urchins), Kelp = mean(kelp) )

df$pch = as.character(substring(df$year,4,4))

df_diff = df %>% group_by(kelp_spp, site) %>%
  mutate( Kelp = Kelp - lag(Kelp, default = Kelp[1]),
          Urchins = Urchins - lag(Urchins, default = Urchins[1])) %>%  
  filter(year>2015, kelp_spp == "Pterygophora")

ptery = df_diff

lmp = lm( Kelp ~ Urchins, data = ptery[ptery$site =='Tatoosh Island',])
pter.s.tat = summary(lmp)
lmp = lm( Kelp ~ Urchins, data = ptery[ptery$site =='Destruction Island',])
pter.s.des = summary(lmp)

r2 =  round( pter.s.tat$r.squared , 2 )
pval = round( pter.s.tat$coefficients[2,4] , 2 )

rstatement = bquote(italic(r)^2 == .(format( pter.s.tat$r.squared, digits = 2)))
# rstatement = paste0("r = ", round(sqrt(r2),2))
pstatement = paste0("p = ", pval)


ptery1 <- ggplot(df_diff, aes( x=Urchins , y=Kelp, color=site)) +
         geom_point(size=2.5, pch=df_diff$pch) +
         scale_color_manual(values=site.col$col )+
         xlim(0,NA)+
         ylim(0,NA)+
         # geom_smooth(data = filter(df_diff,site =='Tatoosh Island' ), method='lm', se=FALSE) +
         # # geom_smooth(data = filter(ptery,site =='Destruction Island' ), method='lm', se=FALSE) +
         # annotate(geom="text", x=0.5, y=2.0, label = "Tatoosh", 
         #          hjust = 0, color=site.col$col[2], size = 3) +
         # annotate(geom="text", x=0.5, y=1.7, label = rstatement, hjust = 0, color=site.col$col[2], 
         #          size = 3) +
         # annotate(geom="text", x=0.5, y=1.4, label = pstatement, hjust = 0, color=site.col$col[2], 
         #          size = 3) +
         xlab( expression(paste(Delta, ' urchins per ', m^2) ) )+
         ylab( expression(paste(Delta, italic(' Ptery'), ' stipes per ', m^2)) )+
         theme_bw() + theme_nt 

ptery1


```
## Other
```{r other-vs-urchins-site-year-diff, fig.height=2.5, fig.width=3.5}

df = df_transect %>% group_by(site, year, zone, area, transect, kelp_spp) %>%
                      summarise(urchins = mean(urchin_density), kelp = mean(kelp_density) ) %>% 
                      group_by(site,year,kelp_spp) %>%
                      summarise(Urchins = mean(urchins), Kelp = mean(kelp) )
df$pch = as.character(substring(df$year,4,4))

df_diff = df %>% group_by(kelp_spp, site) %>%
  mutate( Kelp = Kelp - lag(Kelp, default = Kelp[1]),
          Urchins = Urchins - lag(Urchins, default = Urchins[1])) %>%  
  filter(year>2015, kelp_spp == "Other")



other = df

lmp = lm( Kelp ~ Urchins, data = other[other$site =='Tatoosh Island',])
other.s.tat = summary(lmp)

r2 =  round( other.s.tat$r.squared , 2 )
pval = round( other.s.tat$coefficients[2,4] , 2 )

rstatement = bquote(italic(r)^2 == .(format( other.s.tat$r.squared, digits = 2)))
pstatement = paste0("p = ", pval)



other1 <- ggplot(df_diff, aes( x=Urchins , y=Kelp, color=site)) +
         geom_point(size=2.5, pch=df_diff$pch) +
         scale_color_manual(values=site.col$col )+
         # geom_smooth(data = filter(other,site =='Tatoosh Island' ), method='lm', se=FALSE) +
         # annotate(geom="text", x=1.3, y=1.0, label = "Tatoosh", 
         #          hjust = 0, color=site.col$col[2], size = 3) +
         # annotate(geom="text", x=1.3, y=0.6, label = rstatement, hjust = 0, color=site.col$col[2], 
         #          size = 3) +
         # annotate(geom="text", x=1.3, y=0.2, label = pstatement, hjust = 0, color=site.col$col[2], 
         #          size = 3) +
         xlab( expression(paste(Delta, ' urchins per ', m^2) ) )+
         ylab( expression(paste(Delta, ' other', ' stipes per ', m^2)) )+
         theme_bw() + theme_nt 

other1


```

# Macro and Ptery at Neah Bay
```{r neah-year-diff, fig.height=2.5, fig.width=3.5}

df = df_transect %>% group_by(site, year, zone, area, transect, kelp_spp) %>%
                      summarise(urchins = mean(urchin_density), kelp = mean(kelp_density) ) %>% 
                      group_by(site,year,kelp_spp) %>%
                      summarise(Urchins = mean(urchins), Kelp = mean(kelp) ) %>%
                      filter(site %in% c('Neah Bay'))

df$pch = as.character(substring(df$year,4,4))
# difference
dfn = df %>% filter( kelp_spp %in% c("Macrocystis", "Pterygophora") & site == 'Neah Bay') %>% 
  group_by(kelp_spp) %>%   mutate( Kelp = Kelp - lag(Kelp, default = Kelp[1]),
          Urchins = Urchins - lag(Urchins, default = Urchins[1])) %>%  filter(year>2015)

# macro
mx = lm(Kelp ~ Urchins, data = dfn%>%filter(kelp_spp=="Macrocystis"))
smx = summary(mx)
#
r2 =  round( smx$r.squared , 2 )
pval = round( smx$coefficients[2,4] , 2 )

rstatement.m = bquote(italic(r)^2 == .(format( smx$r.squared, digits = 2)))
pstatement.m = paste0("p = ", pval)

# Ptery
px = lm(Kelp ~ Urchins, data = dfn%>%filter(kelp_spp=="Pterygophora"))
spx = summary(px)
#
r2 =  round( spx$r.squared , 2 )
pval = round( spx$coefficients[2,4] , 2 )

rstatement.p = bquote(italic(r)^2 == .(format( spx$r.squared, digits = 2)))
pstatement.p = paste0("p = ", pval)

neah1 <- ggplot(dfn, aes( x=Urchins , y=Kelp, color=kelp_spp)) +
         geom_point(size=2.5, pch=dfn$pch) +
         geom_smooth(data = filter(dfn,kelp_spp =='Macrocystis' ), method='lm', se=FALSE) +
         geom_smooth(data = filter(dfn,kelp_spp =='Pterygophora' ), method='lm', se=FALSE) +
         scale_color_manual(values=site.col$col[c(1,3)] )+
         # macro       
         annotate(geom="text", x=0.02, y=0, label = expression(italic("Macro")), hjust = 0, color=site.col$col[1], size = 3) +
         annotate(geom="text", x=0.02, y=-0.6, label = rstatement.m, hjust = 0, color=site.col$col[1], size = 3) +
         annotate(geom="text", x=0.02, y=-1.2, label = pstatement.m, hjust = 0, color=site.col$col[1], size = 3) +
         # Ptery
         annotate(geom="text", x=-0.06, y=3.00, label = expression(italic("Ptery")), hjust = 0, color=site.col$col[3], size = 3) +
         annotate(geom="text", x=-0.06, y=2.4, label = rstatement.p, hjust = 0, color=site.col$col[3], size = 3) +
         annotate(geom="text", x=-0.06, y=1.8, label = pstatement.p, hjust = 0, color=site.col$col[3], size = 3) +
         xlab( expression(paste(Delta, ' urchins per ', m^2) ) )+
         ylab( expression(paste(Delta,' in stipes per ', m^2) ) )+
         theme_bw() + theme_nt 

neah1


```

## Tatoosh Nereo, Ptery, Other

# Nereo
```{r tatoosh-transect-level-nereo-diff , fig.height=2.5 , fig.width=3.5}

df_tat <- df_transect %>% filter(., site == 'Tatoosh Island') %>% filter(kelp_spp == 'Nereocystis') %>%
                      group_by(year, zone, area, transect) %>%
                      summarize( kelp = mean(kelp_density) , 
                                 urchins = mean(urchin_density) ) 

dfx = df_tat %>% 
  group_by(zone, area, transect) %>%
  mutate( kelp = kelp - lag(kelp, default = kelp[1]),
          urchins = urchins - lag(urchins, default = urchins[1])) %>%  
  filter(year != 2015,
         kelp != 0)

dfx$pch = as.character(substring(dfx$year,4,4))


dfy = na.omit(df_diff)

X = dfy$urchins
Y = dfy$kelp

dfx$zone = as.character(dfx$zone)
dfx$zone[dfx$zone == 5] <- "5 m"
dfx$zone[dfx$zone == 10] <- "10 m"

dfx$zone = factor(dfx$zone, levels = c('5 m','10 m'))

nereo.tat = ggplot(dfx , aes(x = urchins, y = kelp, color=zone)) + 
        #geom_line(data=pr, aes(x=X,y=p2), color='red') +
        geom_text(data=dfx , key_glyph = "point", aes(x=urchins, y=kelp, label=pch), size = 3)+
        scale_color_manual(values=site.col$col[c(1,3)])+
        xlab( expression(paste(Delta, ' urchins per ', m^2) ) )+
        ylab( expression(paste(Delta, italic(' Nereo'), ' stipes per ', m^2)) )+
        # annotate(geom="text", x=-3, y=10, label = rstatement, hjust = 0, color="black", size = 3) +
        # annotate(geom="text", x=-3, y=8, label= pstatement, hjust = 0, color="black", size = 3) +
        theme_bw() + theme_nt

nereo.tat

```

## Ptery

```{r tatoosh-transect-level-ptery-diff , fig.height=2.5 , fig.width=3.5}

df_tat <-df_transect %>% filter(., site == 'Tatoosh Island') %>% filter(kelp_spp == 'Pterygophora') %>%
                      group_by(site,year,transect,area,zone) %>%
                      summarize( kelp = mean(kelp_density) , 
                                 urchins = mean(urchin_density) )

dfx = df_tat %>% 
  group_by(zone, area, transect) %>%
  mutate( kelp = kelp - lag(kelp, default = kelp[1]),
          urchins = urchins - lag(urchins, default = urchins[1])) %>%  
  filter(year != 2015,
         kelp != 0)
dfx$pch = as.character(substring(dfx$year,4,4))

dfy = na.omit(dfx)

X = dfy$urchins
Y = dfy$kelp


dfx$zone = as.character(dfx$zone)
dfx$zone[dfx$zone == 5] <- "5 m"
dfx$zone[dfx$zone == 10] <- "10 m"

dfx$zone = factor(dfx$zone, levels = c('5 m','10 m'))

ptery.tat = ggplot(dfx , aes(x = urchins, y = kelp, color=zone)) + 
        # geom_line(data=pr, aes(x=X,y=p2), color='red') +
        geom_text(data=dfx , key_glyph = "point", aes(x=urchins, y=kelp, label=pch), size = 3)+
        scale_color_manual(values=site.col$col[c(1,3)])+
        xlab( expression(paste(Delta, 'urchins per ', m^2) ) )+
        ylab( expression(paste(Delta, italic(' Ptery'), ' stipes per ', m^2)) )+
        # annotate(geom="text", x=4, y=15, label = rstatement, hjust = 0, color="black", size = 3) +
        # annotate(geom="text", x=4, y=13, label= pstatement, hjust = 0, color="black", size = 3) +
        theme_bw() + theme_nt

ptery.tat

```
## Other
```{r tatoosh-transect-level-other-diff , fig.height=2.5 , fig.width=3.5}

df_tat <-df_transect %>% filter(., site == 'Tatoosh Island') %>% filter(kelp_spp == 'Other') %>%
                      group_by(site,year,transect,area,zone) %>%
                      summarize( kelp = mean(kelp_density) , 
                                 urchins = mean(urchin_density) )

dfx = df_tat %>% 
  group_by(zone, area, transect) %>%
  mutate( kelp = kelp - lag(kelp, default = kelp[1]),
          urchins = urchins - lag(urchins, default = urchins[1])) %>%  
  filter(year != 2015,
         kelp != 0)

dfx$pch = as.character(substring(dfx$year,4,4))

dfy = na.omit(dfx)

X = dfy$urchins
Y = dfy$kelp

dfy = na.omit(dfx)

dfx$zone = as.character(dfx$zone)
dfx$zone[dfx$zone == 5] <- "5 m"
dfx$zone[dfx$zone == 10] <- "10 m"

dfx$zone = factor(dfx$zone, levels = c('5 m','10 m'))

other.tat = ggplot(dfx , aes(x = urchins, y = kelp, color=zone)) + 
        # geom_line(data=pr, aes(x=X,y=p2), color='red') +
        geom_text(data=dfx , key_glyph = "point", aes(x=urchins, y=kelp, label=pch), size = 3)+
        scale_color_manual(values=site.col$col[c(1,3)])+
        xlab( expression(paste(Delta, ' urchins per ', m^2) ) )+
        ylab( expression(paste(Delta, ' other stipes per ', m^2)) )+
        # annotate(geom="text", x=4, y=15, label = rstatement, hjust = 0, color="black", size = 3) +
        # annotate(geom="text", x=4, y=13, label= pstatement, hjust = 0, color="black", size = 3) +
        theme_bw() + theme_nt

other.tat

```
# Combined Figure 5




```{r Figure-5-diff, fig.height=7, fig.width=6.5}
SIZE = 3
plot2  <- plot1 +  theme(legend.position = 'right', 
                         legend.key.size = unit(0.25,'lines'),
                         legend.text = element_text(size=8)) + 
                         annotate(geom="text",x=-0.3,y=1.0, label="a)", size=SIZE)

macro2 <- macro1 + theme(legend.position = c(0.55,0.7)) + 
                         annotate(geom="text",x=0.07,y=3.0, label="b)", size=SIZE)

nereo2 <- nereo1 + theme(legend.position = 'none')+ 
                         annotate(geom="text",x=-1.0,y=3.4, label="c)", size=SIZE)

ptery2 <- ptery1 + theme(legend.position = 'none')+ 
                         annotate(geom="text",x=0.05,y=2, label="d)", size=SIZE)

other2 <- other1 + theme(legend.position = 'none')+ 
                         annotate(geom="text",x=-1.0,y=2, label="e)", size=SIZE)

neah2 <- neah1 +   theme(legend.position = "none") + 
                         annotate(geom="text",x=-0.07,y=3.4, label="f) Neah Bay", size=SIZE,
                                  hjust='left')

nereo.tat2 <- nereo.tat +theme(legend.position = c(0.75,0.2))+ 
                         annotate(geom="text",x=-6.0,y=10, label="g) Tatoosh Is", size=SIZE,
                                  hjust='left')

ptery.tat2 <- ptery.tat +theme(legend.position = 'none')+ 
                         annotate(geom="text",x=-6.0,y=10, label="h) Tatoosh Is", size=SIZE,
                                  hjust='left')

other.tat2 <- other.tat +theme(legend.position = 'none')+ 
                         annotate(geom="text",x=-6.0,y=10, label="i) Tatoosh Is", size=SIZE,
                                  hjust='left')

library(gridExtra)

grid.arrange(
  grobs = list(plot2, 
               macro2, nereo2, 
               ptery2, other2,neah2,
               nereo.tat2, ptery.tat2, other.tat2),
  widths = c(1,1,1,1,1,1),
  # layout_matrix = rbind(c(NA,NA,1,1,1,NA),
  #                       c(NA,2,2,3,3,NA),
  #                       c(4,4,5,5,6,6),
  #                       c(7,7,8,8,9,9))
  layout_matrix = rbind(c(NA,NA,1,1,1,NA),
                        c(2,2,3,3,4,4),
                        c(NA,5,5,6,6,NA),
                        c(7,7,8,8,9,9))
)

```


