---
title: "Multivariate Fish Analysis"
author: Nick Toimieri
date: 
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
fontsize: 11pt

---

```{r SETUP, include=FALSE}
# A bunch of initial set up junk

# rm(list = ls())

##### knitr stuff and libraries ####

HomeFile = getwd()
Fig_Loc = paste0(HomeFile,"/Figures/")
Data_Loc = paste0(HomeFile,"/Data/")
Results_Loc = paste0(HomeFile,"/Results/")
Other_Files = paste0(HomeFile,"/Other Files/")

##### TRIGGERS AND SETTINGS #####

process.data = FALSE # FALSE to speed up and use previously processed data

fish.include = c("OPEL", "HEXA", "EMBI", "AUFL", "BAIT",  
                 "SECA" , "SCMA" ,"SENE", 
                 "SEFL" , "SEME" , "SEMY" , 
                 "SECAy","SEPIy" , "SEMEy","SEBYTy" , "RYOY") 

# min rockfish size for YOY for correcting some files

min_rockfish_size = 6 # higher?

# include 2015 or not
Include.2015 = FALSE

# Delete low vis transects
Min.Vis = 2.0 # NA will use all the data, otherwise set min.

# common transform for all multivar analyses

data.transform = 'none' # 'none', 'log', 'sqrt' , '4th-root'

#########################

if(Include.2015 == TRUE){year.statement <- "NOTE: These analyses include 2015." }
if(Include.2015 == FALSE){year.statement <-  "NOTE: These analyses DO NOT include 2015."}

if(is.na(Min.Vis)){vis.statement <- "NOTE: The data were not corrected for visibility." }
if(!is.na(Min.Vis)){vis.statement <- paste0("NOTE: Transects with visibility < ", Min.Vis, " m were deleted from the analysis.")}
  
###########################


# tools in use 0
library(knitr)
# library(tidyr)
# library(dplyr)
library(tidyverse)
library(stringr)
library(tinytex)
library(RColorBrewer)
library(readxl)

# stats packages etc
library(vegan)
library(BiodiversityR)
library(pracma)
library(factoextra)

# options
knitr::opts_chunk$set(
  fig.path = Fig_Loc,
  cache.path = 'cache/graphics-', 
  echo=FALSE,
  error=FALSE,
  include = TRUE,
  dev='png',
  dpi=300,
  warning=FALSE,
  #out.width = '100%',
  fig.align='center'
  
  )
# opts_knit$set(eval.after = "fig.cap")

# just to keep track of when I ran things.

Last_Run0 = Sys.time()
Last_Run = str_replace(Last_Run0, ":", ".")
Last_Run = str_replace(Last_Run, ":", ".")
capture.output(Last_Run, file = 'Last_Run.txt')

par(ps=10, cex=1)

```

# Intro

This is an up-dated fish file with only the "final" fish multivariate analyses for the Foundational MS.  


# Import and manipulate data

For data processing: 

* Corrected some spp codes
* Make a 'group' variable for lumping taxa/species
* Combined 2015 and 2016-2019 data
  * 2015 data are not necessarily used in analyses
  * 2015 data lack visibility estimates
* No depth data for the 2015 data. 
  * I took the average of the quadrats.
  * I assigned transects with an average depth of >20 ft to the 10-m group.

`r vis.statement`  

`r year.statement`

DATA TRANSFORM: `r data.transform`

I set the data transform to 'none' because there wasn't a lot of variation in spp density among groups.  Some but not a lot. Again, this emphasizes the more abudnant species, which is, I think, ok.

```{r Import_Data , include=FALSE }
#

spp_code =  data.frame(read.csv("spp_codes.csv", header = TRUE))

if(process.data == TRUE){

fishx =  data.frame(read.csv(paste0(Data_Loc, "NWFSC_FISH_ALLYEARS_data_2019.csv"), header = TRUE))


# Bring in 2015 data and combine
df1 <- data.frame(read_excel( paste0(Data_Loc, "2015_OCNMSDataComplete_standardized_122116.xlsx"),
                              sheet = "Master data"))

# use quadrats to apply depth to transects and get zones

id = paste(df1$Site, df1$Transect, df1$Date, sep='_')
x = data.frame(cbind(id,df1$Depth.ft))
colnames(x) <- c('id','depth.ft')
x$depth.ft = as.numeric(x$depth.ft)
depth = aggregate( x$depth.ft ~ id, data=x, FUN = mean, na.action = na.omit)
colnames(depth) <- c('id','ft')

# get just transects
df2 = df1[ df1$data.type == 'swath' ,]
# get just fish
df3 = df2[ df2$PISCO.Classcode %in% spp_code$code , ]
df3$YEAR = substring(df3$Date, 1 , 4)

id3 = paste(df3$Site, df3$Transect, df3$Date, sep='_')
df3$Depth.ft <- depth$ft[ match(id3, depth$id)]

# subset columns to match fish0
df4 = df3[ , c('YEAR', 'Site','Observer','Depth.ft', 'Transect', 'PISCO.Classcode' , 'Count', 'Size.cm')]

# change to fish0 colnames
colnames(df4) <- c('YEAR', 'SITE', 'OBSERVER', 'DEPTH_FT', 'TRANSECT', 'SPECIES' , 'QUANTITY', 'Size_Min')
df4$Size_Max <- df4$Size_Min

# populate data frame for 2015 that matches 2016-2019
dfx =  data.frame(array(NA, dim=c(nrow(df4),ncol(fishx))))
colnames(dfx)<- colnames(fishx)
x <- colnames(df4)

for(i in 1:length(x)){
  dfx[,x[i]] = df4[,x[i]]  
}
# correct zone for 2015
dfx$ZONE <- ifelse(dfx$DEPTH_FT > 20, 10, 5)

fish_all <- rbind(dfx,fishx)

# transfer some names
fish_all$group = spp_code$group[ match( fish_all$SPECIES , spp_code$code)]
fish_all$spp = spp_code$spp[ match( fish_all$SPECIES , spp_code$code)]
fish_all$com_name = spp_code$com_name[ match( fish_all$SPECIES , spp_code$code)]

# get just 2016-2019 sites for analysis and later plotting

fish_all$Size_Max = as.numeric(fish_all$Size_Max)

# fish_all$SPECIES <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$SPECIES == "SECA" ,  "SECAy", fish_all$SPECIES)
# fish_all$SPECIES <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$SPECIES == "SEPI" ,  "SEPIy",  fish_all$SPECIES)
# fish_all$SPECIES <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$SPECIES == "SEME" ,  "SEMEy",  fish_all$SPECIES)
# fish_all$SPECIES <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$SPECIES == "SEFL" ,  "SEBYTy",  fish_all$SPECIES)

fish_all$group <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$group == "SECA" ,  "SECAy",  fish_all$group)
fish_all$group <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$group == "SEPI" ,  "SEPIy",  fish_all$group)
fish_all$group <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$group == "SEME" ,  "SEMEy",  fish_all$group)
fish_all$group <- ifelse(fish_all$Size_Max <= min_rockfish_size & fish_all$group == "SEFL" ,  "SEBYTy",  fish_all$group)

write.csv(fish_all , paste0(Data_Loc,"Data_Fish_All_2015_2019.csv") , row.names = FALSE)

} # END IMPORT AND PROCESS DATA

if(process.data == FALSE){ # bring in data file again
  # just to get sites
  fishx =  data.frame(read.csv(paste0(Data_Loc, "NWFSC_FISH_ALLYEARS_data_2019.csv"), header = TRUE))
  # data for analysis
  fish_all = data.frame(read.csv(paste0(Data_Loc,"Data_Fish_All_2015_2019.csv"), header = TRUE))}

site =  data.frame(matrix(c(levels(as.factor(fishx$SITE)),'black','darkgrey','green','red','blue'),  ncol = 2))

fish0 <- fish_all[ fish_all$group %in% fish.include, ]
fish0 <- fish0[fish0$SITE %in% site[,1], ]




############# some common plotting stuff #####

years = 2015:2019
PCH = c(8, 21, 22, 24, 25)
if( !is.na(Min.Vis) ){fish0 <- fish0[ fish0$VIS_M >= Min.Vis , ] }
if( Include.2015 == FALSE ){ fish0 <- fish0[ fish0$YEAR != 2015 , ] ; 
                             years = 2016:2019 
                             PCH = c(21, 22, 24,25) }
year.pch = data.frame(cbind(years, PCH))

colnames(site) <- c('site', 'col')
site$col = as.character(site$col)
site$order = c( 3, 2 , 1 , 5, 4)
site = site[ order(site$order , decreasing = TRUE) ,]



```

NOTES

* Some rockfish were recorded in the 'adult' category (eg SECA instead of SECAy) but were small.  
* Set minimum size for rockfishes to `r min_rockfish_size` cm. Fish <= to `r min_rockfish_size` cm were classified as YOY.
* Maintained their original desigation in SPECIES column.  New designation in the 'group' column.

```{r Functions}
# removes rows with zero data and columns with too few observations

Sum_Stats <- function( form1 , data.file ){
  # calculate a bunch of summary stats and put in one file
  # text in quotes, form1 looks like: "Y = site + year"
  # no spaces allowed in column titles
  
  library('stringr')
  data.file$xNx = 1

  mn = aggregate( as.formula( form1 ) , data = data.file , FUN = mean )
  s2 = aggregate( as.formula( form1 ) , data = data.file , FUN = var )
  
  x = stringr::str_split( form1 , "~", simplify = TRUE )[2]
  form2 = paste0( 'xNx ~ ', x )
  n  = aggregate( as.formula( form2 ) , data = data.file , FUN = sum )
  
  BY = stringr::str_split( x , "\\+" , simplify = TRUE ) 
  BY = stringr::str_remove_all(BY," ")
  BY = as.vector(BY , mode = 'character')
  
  a = merge( mn , s2 , by = BY, all = TRUE )
  z = merge( a , n , by = BY , all = TRUE )
  colnames( z )[(ncol(z)-2):ncol(z)] <- c( 'mean' , 'var' , 'n' ) 
  
  z$sd = sqrt(z$var)
  z$se = z$sd/sqrt((z$n-1))
  
  return( z )
  
}

```

Because there were a lot of zeros for some species, we lumped some species or taxa them into higher taxonomic categories. Species higher grouping used in the analyses are:

```{r Species_Included}


spp.inlcluded = spp_code[spp_code$group %in% fish.include , c(1,2,4)]
colnames(spp.inlcluded) <- c('Spp_Code', 'Spp_Group', 'Common_Name')
spp.inlcluded


```
## Year x Site x Depth  

No text here. Just calculating averages by year x site x depth and outputting data. 

```{r Group_Year_Site_Depth , include = FALSE}
# aggregate for analysis
fish0$QUANTITY <- as.numeric(fish0$QUANTITY)

l.fish.ysz = aggregate( QUANTITY ~ YEAR + SITE + ZONE + group , data = fish0 , FUN = mean )
# change to wide format
w.fish.ysz = spread(l.fish.ysz , group , QUANTITY )
w.fish.ysz[is.na(w.fish.ysz)] <- 0

# for plotting
fish_info = w.fish.ysz[,c('YEAR','SITE','ZONE')]
fish_info$color = site$col[match(fish_info$SITE, site$site)]
fish_info$pch = ifelse(fish_info$ZONE == 5, 19, 21)
fish_info$id = paste(fish_info$SITE, fish_info$YEAR, fish_info$ZONE, sep="_")
w.fish.ysz = cbind(fish_info[,c('color','pch','id')] , w.fish.ysz)

# save out
write.csv(l.fish.ysz , paste0(Data_Loc, "D_Fish_Group_Site_Year_Zone_long.csv") , row.names = FALSE)
write.csv(w.fish.ysz , paste0(Data_Loc, "D_Fish_Group_Site_Year_Zone_wide.csv") , row.names = FALSE)

```

## Year x Site

No text here. Just calculating averages by year x site and outputting data.  

```{r Group_Year_Site , include = FALSE}
# aggregate for analysis
l.fish.ys = aggregate( QUANTITY ~ YEAR + SITE + group , data = fish0 , FUN = mean )
# change to wide format
w.fish.ys = spread(l.fish.ys , group , QUANTITY )
w.fish.ys[is.na(w.fish.ys)] <- 0

# for plotting
fish_info = w.fish.ys[,c('YEAR','SITE' )]
fish_info$color = site$col[match(fish_info$SITE, site$site)]
fish_info$id = paste(fish_info$SITE, fish_info$YEAR , sep="_")
w.fish.ys = cbind(fish_info[, c('color', 'id') ] , w.fish.ys)

# save out
write.csv(l.fish.ys , paste0(Data_Loc, "D_Fish_Group_Site_Year_long.csv") , row.names = FALSE)
write.csv(w.fish.ys , paste0(Data_Loc, "D_Fish_Group_Site_Year_wide.csv") , row.names = FALSE)

```


# Constrained Ordinations 



```{r Set_Fish_Transform}
# spread to WIDE
# transform data

# set fish to include - remove YOY

cap.fish =  c("OPEL", "HEXA", "EMBI", "AUFL",  "BAIT",  
                 "SECA" , "SCMA" ,"SENE",  "SEFL" , "SEME" ) 

l.fish  = aggregate( QUANTITY ~ YEAR + SITE + ZONE + TRANSECT + group , data = fish0 , FUN = mean )
# change to wide format
w.fish1 = spread(l.fish , group , QUANTITY )
w.fish1[is.na(w.fish1)] <- 0
rownames(w.fish1) <- paste0("Tran_", 1:nrow(w.fish1))

x = rowSums(w.fish1[ , cap.fish])
w.fish <- w.fish1[ x != 0 , ]

if(data.transform == "none"){df_fish = w.fish[, cap.fish]}
if(data.transform == "sqrt"){df_fish = w.fish[, cap.fish]^(1/2)}
if(data.transform == "4th-root"){df_fish = w.fish[, cap.fish]^(1/4)}
if(data.transform == "log"){df_fish = log(w.fish[, cap.fish]+1) }

# remove transects with zero fish

```

Note: there are no rockfish YOY in these ordinations and analyses. 

DATA TRANSFORM: `r data.transform`


We deleted YOY from the analyses because they were highly variable and often appeared in only one year. Taxa included in the actual analyses were:

Taxa Included: `r cap.fish`



## Basic - Discriminat CAP Site x Depth x Year

### Permanova  

In the PerMANVOA everything is significant.  I think this is OK.  Things are messy.  

The associated ordination is more clear. This ordination uses site x year x depth as groups and transects as replicates.  I have, however, calculated the centroids for plotting.

FYI removing BAIT doesn't do much.

```{r PerManova_Transect}

pm1 = adonis( df_fish ~ SITE*ZONE*YEAR, data = w.fish, method = 'bray')

# summary(pm1)
pm1$aov.tab

saveRDS(pm1, file = paste0(Results_Loc,'Fish_PerMANOVA_transect_all.rds'))
capture.output(pm1, file =  paste0(Results_Loc,'Fish_PerMANOVA_transect_all.txt'))
```
### CAP 


### capscale - site x year x depth

To match PerMANOVA

```{r capscale1 , include = FALSE}

CAPid = as.factor(paste(w.fish$SITE , w.fish$ZONE , w.fish$YEAR, sep = "_"))
CAPfish = data.frame(cbind(CAPid, w.fish))
# fish.start = grep('AUFL', colnames(CAPfish))
# df_fish <- (CAPfish[, fish.start:ncol(CAPfish) ] )^(1/4)
df_fish <- as.matrix(df_fish)
cap2 = capscale( df_fish  ~  as.factor(CAPid) , data=CAPfish , dist = 'bray' )
# cap2
cap2_scores = scores(cap2)

cap2_df = data.frame(cbind( cap2_scores$sites , w.fish ))

capture.output( cap2, paste0(Results_Loc,"R_capscale_site_depth_year.txt"))
saveRDS( cap2, paste0(Results_Loc,"R_capscale_site_depth_year.rds"))
```

```{r Fish_capscale1_plot_YSD , fig.width=6, fig.height= 6, fig.cap="Ordination plot for CAP based on site x year x depth.  Error bars are +/- 1.0 s.e."}
cap2_df$col = site$col[ match(cap2_df$SITE , site$site)]
cap2_df$pch = site$pch[ match(cap2_df$ZONE , site$site)]

par( mfrow = c(2,2), pty='s' , mar = c(4,4,1,1))

Xlim = c(-4, 4)
Ylim = c(-4,4)

plot(cap2_df$CAP1 , cap2_df$CAP2, pch = 19 , col = cap2_df$col, xlim = Xlim, ylim = Ylim )
# summarize
cap2_df1 = Sum_Stats( 'CAP1 ~ SITE + YEAR + col', cap2_df)
cap2_df2 = Sum_Stats(' CAP2 ~ SITE + YEAR + col', cap2_df)

cap2_df1$pch = year.pch$PCH[ match(cap2_df1$YEAR , year.pch$years)]

segments( par()$usr[1],0,par()$usr[2],0, lty = 'dotted', lwd = 0.5)
segments( 0, par()$usr[3],0,par()$usr[4], lty = 'dotted', lwd = 0.5)
legend('topright' , legend = site$site, col = site$col , pch = 19, bty = 'n', cex = 0.8)

plot(cap2_df1$mean , cap2_df2$mean , pch = cap2_df1$pch, col=cap2_df1$col , bg = cap2_df1$col , xlim = c(-2,2) , ylim = c(-2,2), cex = 1.2)
arrows( cap2_df1$mean+cap2_df1$se , cap2_df2$mean,  cap2_df1$mean-cap2_df1$se , cap2_df2$mean, col = cap2_df1$col, length = 0)
arrows( cap2_df1$mean, cap2_df2$mean+cap2_df2$se ,  cap2_df1$mean , cap2_df2$mean-cap2_df2$se, col = cap2_df1$col, length = 0)
segments( par()$usr[1],0,par()$usr[2],0, lty = 'dotted', lwd = 0.5)
segments( 0, par()$usr[3],0,par()$usr[4], lty = 'dotted', lwd = 0.5)
# legend('bottomright' , legend = years,  pch = PCH, bty = 'n', cex = 0.8)


spp_scores = data.frame(cap2_scores$species)
plot(spp_scores[,1] , spp_scores[,2], pch="", xlim = Xlim, ylim = Ylim )
text(spp_scores[,1] , spp_scores[,2], rownames(spp_scores), cex=0.8)
segments( par()$usr[1],0,par()$usr[2],0, lty = 'dotted', lwd = 0.5)
segments( 0, par()$usr[3],0,par()$usr[4], lty = 'dotted', lwd = 0.5)


plot(1:10, 1:10, pch="", xlab=NA, ylab=NA, xaxt='n', yaxt='n', bty='n')
legend('left' , legend = site$site, col = site$col , pch = 19, bty = 'n', cex = 1)
legend('right' , legend = years, pch = PCH, bty = 'n', cex = 1)
```



Same data replotted on one pane. We can see some clear site differences:

* Cape Alava and Cape Johnson do not overlap
* Tatoosh and Neah Bay are intermediate between the former two
* Destruction is all over the place



```{r Fish_capscale1_plot_YSD_1pane , fig.width=6, fig.height= 6, fig.cap="Ordination plot for CAP based on site x year x depth.  Error bars are +/- 1.0 s.e."}
cap2_df$col = site$col[ match(cap2_df$SITE , site$site)]
cap2_df$pch = site$pch[ match(cap2_df$ZONE , site$site)]

par(pty='s')

# cal means
cap2_df1 = Sum_Stats( 'CAP1 ~ SITE + YEAR + col', cap2_df)
cap2_df2 = Sum_Stats(' CAP2 ~ SITE + YEAR + col', cap2_df)
cap2_df1$pch = year.pch$PCH[ match(cap2_df1$YEAR , year.pch$years)]

xy = 2.5

plot(cap2_df1$mean , cap2_df2$mean , pch = cap2_df1$pch, col=cap2_df1$col , bg = cap2_df1$col , xlim = c(-xy,xy) , ylim = c(-xy,xy), cex = 1.2)
arrows( cap2_df1$mean+cap2_df1$se , cap2_df2$mean,  cap2_df1$mean-cap2_df1$se , cap2_df2$mean, col = cap2_df1$col, length = 0)
arrows( cap2_df1$mean, cap2_df2$mean+cap2_df2$se ,  cap2_df1$mean , cap2_df2$mean-cap2_df2$se, col = cap2_df1$col, length = 0)
segments( par()$usr[1],0,par()$usr[2],0, lty = 'dotted', lwd = 0.5)
segments( 0, par()$usr[3],0,par()$usr[4], lty = 'dotted', lwd = 0.5)

spp_scores = data.frame(cap2_scores$species)
spp_scores1 = spp_scores[ abs(spp_scores$CAP1) > 0.2 | abs(spp_scores$CAP2) > 0.2, ]


text(spp_scores1[,1] , spp_scores1[,2], rownames(spp_scores1), cex=0.8 , col='red', font=2)

legend('topleft' , legend = site$site, col = site$col , pch = 19, bty = 'n', cex = 1)
legend('topright' , legend = years, pch = PCH, bty = 'n', cex = 1)
```

```{r CAPDiscrim , include = FALSE}
# for some reason, there are no species scores.  Could calculate correlations instead.

CAPid = data.frame(as.factor(paste(w.fish$SITE , w.fish$ZONE , w.fish$YEAR, sep = "_")))
CAPid = data.frame(as.factor(w.fish$SITE ))
colnames(CAPid) = 'id'
fish.div = vegdist( df_fish , method = "bray")
cap_dis <- CAPdiscrim( fish.div ~ id , data = CAPid)

ordiplot(cap_dis)

```

```{r Import_UPC_Data, include=FALSE}

# bring in data

# algae

algae_zone_long = data.frame(read.csv( paste0(Data_Loc , "algae_swath_zonelevel_lump.csv"), header = TRUE))
algae_zone <- spread(algae_zone_long[, c('year','site','zone','classcode','dens_weighted')], classcode, dens_weighted)
colnames(algae_zone)[1:3] <- c('YEAR', 'SITE','ZONE')



# UPC ####
upc_site_long = data.frame(read.csv( paste0(Data_Loc , "UPC_Year_Site_Level_Summary.csv"), header = TRUE))
upc_zone_long = data.frame(read.csv( paste0(Data_Loc , "UPC_Zone_Level_Summary.csv"), header = TRUE))
# convert to wide 
upc_site <- spread(upc_site_long[ ,c("SITE", "YEAR", "Functional_Group","MEAN")], Functional_Group, MEAN)
upc_zone <- spread(upc_zone_long[ ,c("SITE", "YEAR", "ZONE", "Functional_Group","MEAN")], Functional_Group, MEAN)

# Substrate ####

rock_site_long = data.frame(read.csv( paste0(Data_Loc , "SUBSTRATE summary by YEAR-SITE 2016-2019.csv"), header = TRUE))
rock_zone_long = data.frame(read.csv( paste0(Data_Loc , "SUBSTRATE summary by YEAR-SITE-DEPTH 2016-2019.csv"), header = TRUE))
# convert to wide
rock_site <- spread( rock_site_long[, c("YEAR", "SITE", "CLASSCODE", "MEAN")] , CLASSCODE , MEAN)
rock_zone <- spread( rock_zone_long[, c("YEAR", "SITE", "ZONE", "CLASSCODE", "MEAN")] , CLASSCODE , MEAN)
rock_zone$pch = year.pch$PCH[ match(rock_zone$YEAR , year.pch$years)]
rock_zone$col = site$col[ match(rock_zone$SITE , site$site)]  
# Relief ####

relief.div_site_long = data.frame(read.csv( paste0(Data_Loc , "RELIEF_DIVERSITY_Year_Site_Level_Summary.csv"), header = TRUE))
relief.div_zone_long = data.frame(read.csv( paste0(Data_Loc , "RELIEF_DIVERSITY_Zone_Level_Summary.csv"), header = TRUE))

# Relief  ####

relief_site_long = data.frame(read.csv( paste0(Data_Loc , "RELIEF_Year_Site_Level_Summary.csv"), header = TRUE))

relief_zone_long = data.frame(read.csv( paste0(Data_Loc , "RELIEF_Zone_Level_Summary.csv"), header = TRUE))
# convert to wide
relief_site <- spread( relief_site_long[, c("YEAR", "SITE", "CLASSCODE", "MEAN")] , CLASSCODE , MEAN)
relief_zone <- spread( relief_zone_long[, c("YEAR", "SITE", "ZONE", "CLASSCODE", "MEAN")] , CLASSCODE , MEAN)


#### combine all substrate data for pca


hab1 = merge(rock_zone, relief_zone , by = c("YEAR", "SITE", "ZONE"))
hab = merge(hab1, relief.div_zone_long[,c("YEAR", "SITE", "ZONE","MEAN")], by = c("YEAR", "SITE", "ZONE"))
colnames(hab)[ncol(hab)] <- "rel.div"

```


* PCA the various habitat variables to reduces dimensions
* Do separately for each variable type
* Initial analysis by depth zone to maintain the most replicates
  * not necessarily interested in depth for the analyses

## Substrate PCA - YEAR x SITE x DEPTH means

PC1 mostly distinguishes between bedrock and boulder. Boulder areas had higher relief (>2m) and higher relief diversity. PC2 tends to distinguish between the two mid-complexity categories.

While there is some variation among years within sites (spread), it isn't bad.  Mostly, there are some obvious differences among sites with regard to habitat. 

This might be a good initial ordination figure just describing the sites and showing that the physcial habitat differs among sites. Depths to a lesser extent.

* Tatoosh and Destruction Island have more bedrock and higher relief than other areas. 
* Neah Bay is bascially intermediate for everything.
* Cape Alava and Cape Johnson have more boulder and kind of intermediate relief.



```{r pca_rock , fig.width=5, fig.height = 5, fig.cap= "Substrate PCA. Open circles are 5-m depth zone; closed circles 10-m depth zone."}  

Vars = c("BEDRK","BOULD","COB","SAND",">2m","0-10cm","10cm-1m","1m-2m","rel.div")

hab_pca <- princomp( hab[, Vars] )

par(pty='s', mar = c(4,3,2,1))

# screeplot(rock_pca)
fviz_eig(hab_pca)


# # biplot(rock_pca , arrow.len = 0.05)
# 
# fviz_pca_biplot(hab_pca, repel = TRUE,
#                 geom.ind = c('point'),
#                 col.var = "#2E9FDF", # Variables color
#                 col.ind =  hab$col, # Individuals color
#                 #addEllipses = TRUE,
#                 title = "Substrate Biplot"
#                 )
# 

# good old ordiplot

hab$col2 = ifelse( hab$ZONE == 5 , NA, hab$col)
par( pty = 's', ps = 10, cex = 1.0)
ordiplot(hab_pca, display = 'species' , type = 'text' , xlim = c(-1 , 1) , ylim = c(-1 , 1))

pcs_hab = scores(hab_pca)
points(pcs_hab[,1] , pcs_hab[,2], pch = hab$pch , col = hab$col, bg = hab$col2 , lwd = 1.5 )

segments( par()$usr[1],0,par()$usr[2],0, lty = 'dotted', lwd = 0.5)
segments( 0, par()$usr[3],0,par()$usr[4], lty = 'dotted', lwd = 0.5)

legend('topright' , legend = site$site, pch=19, col=site$col, bty='n', cex = 0.8)
legend('topleft' , legend = years, pch=PCH , bty = 'n', cex= 0.8)

x = data.frame(hab[,c('YEAR','SITE','ZONE','pch','col')])
abiotic = data.frame(cbind(x,pcs_hab))

write.csv(abiotic , paste0(Data_Loc,"PCs_Abiotic_Habitat.csv") , row.names = TRUE)

```

### Fish vs abiotic habitat and kelp

I combined the PCs from the habitat analysis with the kelp data.  I then ran and RDA style constrained analysis.    

This constrained habitat vs. fish analysis is ugly. In fact, it is non-significant! So while there are differences among sites x depth x year in fish abundance, they don't seem related to habitat directly.  

This actually makes sense when you compare the separate fish and habitat ordinations. For example, Cape Alava and Cape Johnson have different fish fauna, but similar habitat. 


```{r Can_Cor_Type_RDA}

# merge data

hab_vars0 = data.frame(merge(abiotic, algae_zone , by=c("YEAR", "SITE", "ZONE")))
hab_vars1 = apply( hab_vars0[, c('Comp.1','Comp.2','MACPYR','NERLUE','PTECAL','OTHER') ] , 2, scale)
hab_vars = cbind(hab_vars0[,1:5] , hab_vars1)

cap_data = data.frame(merge(w.fish.ysz , hab_vars, by=c("YEAR", "SITE", "ZONE")))

write.csv( cap_data , paste0(Data_Loc, 'D_cap_data_test.csv'), row.names = FALSE)
cap_data = read.csv(paste0(Data_Loc, 'D_cap_data_test.csv'), header=TRUE)
cap4 = capscale( cap_data[,cap.fish] ~ Comp.1 + Comp.2 + MACPYR + NERLUE + OTHER + PTECAL, 
                 data = cap_data)
cap4
cap_sum <- summary(cap4)

scores1 = scores(cap4)
anova(cap4)

```



```{r Hab_Analysis_capscale_plot}


df = data.frame(scores1$sites)
xy = max(df)
ordiplot(cap4 , display = 'species' , type = 'text' , xlim = c(-xy,xy) , ylim = c(-xy,xy) , col = 'blue' )
points( df$CAP1 , df$CAP2 ) 
hab_scores = data.frame(cap_sum$biplot)
text(hab_scores$CAP1, hab_scores$CAP1 , rownames(hab_scores) , col = 'red')

```
***

# Univariate plots by group


Note: I have plotted YOY here but they are not in the above ordinations.  



```{r Fish_Spp_Plots_Site_Year , fig.width=6, fig.height=8, fig.cap="Fish abundance by site and year."}


par(  mfrow = c(4,2) , mar=c(3,4,1,1))

for(i in 1:length(fish.include)){
  form1 = paste0(fish.include[i] , " ~ YEAR + SITE" )
  df1 = Sum_Stats( form1 , w.fish)
  df1$seUP = df1$mean + df1$se
  df1$seLO= df1$mean - df1$se
  Ymax = max(df1$seUP)
  
  # blank plot
  plot( years , seq(0,Ymax, length.out = length(years)) , pch ="", ylim = c(0, Ymax) , 
        xlab = NA, ylab = "Number")
  legend('topleft', legend = fish.include[i], bty='n')
  for(k in 1:nrow(site)){
    df2 = df1[ df1$SITE == site$site[k] , ]
    points( df2$YEAR , df2$mean , type = "b", pch =19, col = site$col[k] )
    
    if(i %in% c(1,9)){ legend('topright', legend = site$site, pch=19 , col=site$col , bty='n' , cex=0.8)}
   
   
  }
  
  
}














