---
title: "Summarise UPC Data OCNMS July 2020"
author: Genoa Sullaway, Jameal Samhouri
date: "July-August 2020"
output: 
  html_document:
      toc: true
      toc_float: true
---

## Approach

\n This uses 2016-2019 data, 2015 UPC uses quadrats and is inconsistent with later methods, calculates mean and standard deviation (specific to beta distributed data) for a few different flavors of data summaries, below. 

\n All data sets are saved in: OCNMS/Data/Summarized data

\n There are three **percent cover** data sets:

\n 1. dat.transect.summary = UPC_Transect_Level_Summary.csv: This has percent cover across TRANSECT/ZONE/SIDE/SITE/YEAR. Transect is lowest level of replication here. There are only 11 transects in whole dataset that have less than 30 upc counts (aka an incomplete transect)... In the cases where it is less than 30 I get total % cover from the total points sampled for that transect.

\n 2. dat.zone.summary = UPC_Zone_Level_Summary.csv: This has mean percent cover and variance across ZONE/SITE/YEAR.

\n 3. dat.year.site.summary = UPC_Year_Site_Level_Summary.csv: This has mean percent cover and variance across SITE/YEAR. 

\n 4. dat.site.summary = UPC_Site_Level_Summary.csv: This has mean percent cover and variance across SITES. 

\n There are four **Simspons Diversity** data sets:

\n 1. dat.diversity = UPC_DIVERSITY_Transect_Level_Summary.csv: This has Simpsons Diversity across TRANSECT/ZONE/SIDE/SITE/YEAR. Transect is lowest level of replication here.

\n 2. dat.diversity.zone.summary = UPC_DIVERSITY_Zone_Level_Summary.csv: This has mean Simpsons Diversity and variance across ZONE/SITE/YEAR.

\n 3. dat.diversity.year.site.summary= UPC_DIVERSITY_Year_Site_Level_Summary.csv: This has mean Simpsons Diversity and variance across SITE/YEAR.

\n 4. dat.diversity.site.summary = UPC_DIVERSITY_Site_Level_Summary.csv: This has Simpsons Diversity and variance across SITES. 

## Summary Thoughts

- Two northern sites, Neah Bay and Tattoosh, appear to have more diversity in percent cover per transect than the three southern sites. There dont appear to be huge differences among years or zones, biggest differences are among Sites. This is reflected in Simpsons Index and in the stacked bar plots using percent cover. 


```{r setup,message=FALSE, warning=FALSE}
#to load data make sure directory is set to knit from project directory
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
 
library(here)
library(tidyverse)
library(knitr)
library(viridis)
 
substrate_codes <- read_csv(here("Data","CSV_2015_on","substrate_codes.csv"))
 
#2016 --> 2019
dat.upc <- read_csv(here("Data","CSV_2015_on","NWFSC_UPC_ALLYEARS_data_2019.csv"), col_types = cols(SIDE = col_character())) %>% 
              filter(CATEGORY == "COVER") %>%
              select(1:20) %>%
              rename(DEPTH.FT= "DEPTH (FT)")  %>%
              #mutate(QUADRAT = NA, PERCENT.COVER = NA) %>% #will circle back tothis and get % cover for 2016 to 2019 data 
              select(SITE, MONTH, DAY, YEAR, OBSERVER, TRANSECT , CLASSCODE, COUNT,DEPTH.FT,SIDE,ZONE,TRANSECT,BUDDY,SEGMENT) %>%
#dat.upc <- rbind(dat.2015, dat.2016) %>%
 mutate(COUNT = replace_na(COUNT, 0)) %>%
 left_join(substrate_codes) %>% #rename two that arent in substrate key
 mutate(NAME = case_when(CLASSCODE == "ptero" | NAME == "Pterogophora" ~ "Pterygophora",
                          CLASSCODE == "shell" ~ "various bivalves",
                          TRUE ~ NAME),
#ASSIGN FUNCTIONAL GROUPS 
         Functional_Group = case_when(
                                    NAME %in% c("Red algae (branching flat blade)", "Red algae (cylindrical branches)", "Red algae (lacy branching)", "Red algae (leaf-like)") ~ "Red Algae",
                                    NAME %in% c("articulated coralline","Encrusting hydrocoral (eg. Stylantheca)","Red algae -Encrusting", "Coralline algae -Crustose") ~ "Encrusting Species",
                                    NAME %in% c("Pterygophora", "Brown algae understory prone","Nereocystis holdfast (alive)", "Macrosystis holdfast (alive)", "Laminariales holdfast (alive)") ~ "Brown Algae",
                                    NAME %in% c("Anemone", "Bryozoan", "Cup Coral","Tunicate -Colonial,compund,social","Tunicate -Solitary", "Hydroid", "Sponge", "Barnacle","Tubeworm") ~ "Non-mobile Invertebrate",
                                     NAME %in% c("various bivalves", "Scallop","urchin", "Cucumaria spp.' sea cucumber") ~ "Mobile Invertebrate",
                                    NAME %in% c("Substrate: Bedrock","Shell Debris","Bare Sand", "Substrate: Cobble","Sediment/Mud", "Substrate: Sand","Bare Rock" ) ~ "Non-living Substrate",
                                    NAME %in% c("Zostera marina, eelgrass" , "Phyllospadix spp.' surfgrass") ~ "Eelgrass_Surfgrass",
                                    TRUE ~ NAME),
 #ASSIGN ORGANISM GROUPS         
        Organism_Group = case_when(NAME %in% c("Red algae (branching flat blade)", "Red algae (cylindrical branches", "Red algae (lacy branching)","Pterygophora", "Brown algae understory prone","Red algae -Encrusting", "Coralline algae -Crustose","Nereocystis holdfast (alive)","Macrosystis holdfast (alive)","Phyllospadix spp.' surfgrass", "Zostera marina, eelgrass","Encrusting hydrocoral (eg. Stylantheca)","Green algae","Laminariales holdfast (alive)","articulated coralline", "Red algae (cylindrical branches)","Red algae (leaf-like)") ~ "Plant_Algae",
                                    NAME %in% c("various bivalves", "Scallop","Anemone", "Cup Coral","Barnacle","Tunicate -Colonial,compund,social", "Tunicate -Solitary","urchin","Bryozoan","Diatom Layer","Sponge","Tubeworm","Hydroid", "Cucumaria spp.' sea cucumber") ~ "Invertebrate",
                                    NAME %in% c("Substrate: Bedrock","Shell Debris","Bare Sand", "Substrate: Cobble","Sediment/Mud", "Substrate: Sand","Bare Rock" ) ~ "Non-living Substrate", 
                                    TRUE ~ NAME))  

 

```

```{r , echo=FALSE}
summary_table<- dat.upc %>%
  group_by(Organism_Group,Functional_Group, NAME) %>%
  summarise(species_occurence = sum(COUNT))  
 
kable(summary_table)
```

## Species Occurence and Grouping Plots 
```{r, echo=FALSE}
dat.upc %>%
  group_by(NAME) %>%
  summarise(species_occurence = sum(COUNT)) %>%
  ggplot() +
  geom_bar(aes(x=reorder(NAME, -species_occurence), y = species_occurence), stat = "identity")+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Original Species Groups") +
  ylab("Species Count")
```

```{r,echo=FALSE}
dat.upc %>%
  group_by(Functional_Group) %>%
  summarise(species_occurence = sum(COUNT)) %>%
  ggplot() +
  geom_bar(aes(x=reorder(Functional_Group, -species_occurence), y = species_occurence), stat = "identity")+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Functional Groups") +
  ylab("Species Count")
```

```{r,echo=FALSE}
dat.upc %>%
  group_by(Organism_Group) %>%
  summarise(species_occurence = sum(COUNT)) %>%
  ggplot() +
  geom_bar(aes(x=reorder(Organism_Group, -species_occurence), y = species_occurence), stat = "identity")+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("Organism Groups") +
  ylab("Species Count")
```

## Percent Cover
### Create data frames
```{r}
dat.transect.summary <- dat.upc %>%
  filter(YEAR > 2015) %>%
  group_by(YEAR,SITE,SIDE,ZONE,TRANSECT,DEPTH.FT,Functional_Group) %>%
  summarise(COUNT = sum(COUNT)) %>%
  spread(Functional_Group, COUNT) %>%
  ungroup() %>%
  replace(is.na(.), 0) %>%
  dplyr::mutate(Sum = rowSums(.[,7:15])) %>%
  tidyr::gather(7:15, key = "Functional_Group", value = "COUNT") %>%
  mutate(percent_cover = COUNT/Sum) 

dat.zone.summary <- dat.transect.summary %>%
  group_by(YEAR,SITE,ZONE,Functional_Group) %>% 
 summarise(
      N=length(percent_cover),
      MEAN=mean(percent_cover),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N)) %>%
  ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay")))

# added by JS 081820
dat.year.site.summary <- dat.transect.summary %>%
  group_by(SITE,YEAR,Functional_Group) %>% 
  summarise(
      N=length(percent_cover),
      MEAN=mean(percent_cover),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N)) %>%
  ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay")))
  
dat.site.summary <- dat.transect.summary %>%
  group_by(SITE,Functional_Group) %>% 
  summarise(
      N=length(percent_cover),
      MEAN=mean(percent_cover),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N)) %>%
  ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay")))
  
  write.csv(dat.transect.summary, here("Data","Summarized data","UPC_Transect_Level_Summary.csv"))
  write.csv(dat.zone.summary, here("Data","Summarized data","UPC_Zone_Level_Summary.csv"))
  write.csv(dat.year.site.summary, here("Data","Summarized data", "UPC_Year_Site_Level_Summary.csv"))
  write.csv(dat.site.summary, here("Data","Summarized data","UPC_Site_Level_Summary.csv"))
  
```

### Plots  
```{r, echo=FALSE}
ggplot(dat.zone.summary,aes(x=YEAR)) +
  geom_bar(aes(y=MEAN, fill= Functional_Group), position="fill", stat="identity",width = 0.95
           )+
  scale_fill_viridis(discrete = T) + #,  guide = guide_legend(reverse = TRUE), name = "Stock") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "left",aspect.ratio = 3/1)  +
  theme_classic()+
  ylab("Mean Percent Cover") +
  xlab("Year") +
  facet_grid(ZONE~SITE) +
 # coord_flip() +
  scale_x_continuous(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0)) +
  theme(legend.position="left") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), strip.text.x = element_text(size = 7))
```


Also see what it looks like if I lump across years and zones, largest differences across zones seem to be increased brown algae in 5m zone and sligthly higher functional group richness in shallower areas. 
 
```{r, echo=FALSE}
ggplot(dat.site.summary,aes(x=SITE)) +
  geom_bar(aes(y=MEAN, fill= Functional_Group), position="fill", stat="identity",width = 0.95
           )+
  scale_fill_viridis(discrete = T) + #,  guide = guide_legend(reverse = TRUE), name = "Stock") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "left",aspect.ratio = 3/1)  +
  theme_classic()+
  ylab("Mean Percent Cover") +
  xlab("Year") +
  #facet_wrap(~ZONE, nrow =1) +
  scale_x_discrete(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0)) +
  theme(legend.position="left") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), strip.text.x = element_text(size = 7))
```

## Simpsons Diversity Index 
### Create data frames
```{r calculate Simpsons Diversity}
#use dat.relief to get Simpsons index for every transect and save data frames 
dat.diversity<- dat.transect.summary %>% 
    group_by(YEAR,SITE,ZONE,SIDE,TRANSECT,Functional_Group) %>%
    dplyr::summarise(
            PROPORTION=sum(percent_cover),
            PROPORTION_2=sum(PROPORTION)^2,
             .groups = "keep") %>%
      group_by(YEAR,SITE,ZONE,SIDE,TRANSECT) %>%
    dplyr::summarise(    
            DIVERSITY= sum(PROPORTION_2),
            RELIEF_SIMPSONS_DIVERSITY= 1 - DIVERSITY, # simpsons diversity
            .groups = "keep"
    ) %>%
    select(-DIVERSITY)

# calculate substrate diversity for each site-year-zone
  dat.diversity.zone.summary <- dat.diversity %>% 
    group_by(YEAR,SITE,ZONE) %>%
    dplyr::summarise(
      MEAN=mean(RELIEF_SIMPSONS_DIVERSITY),
      SD= sd(RELIEF_SIMPSONS_DIVERSITY),
      N=length(RELIEF_SIMPSONS_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )

# calculate substrate diversity for each site
  dat.diversity.year.site.summary <- dat.diversity %>% 
    group_by(YEAR,SITE) %>%
    dplyr::summarise(
      MEAN=mean(RELIEF_SIMPSONS_DIVERSITY),
      SD= sd(RELIEF_SIMPSONS_DIVERSITY),
      N=length(RELIEF_SIMPSONS_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  
  # calculate substrate diversity for each site
  dat.diversity.site.summary <- dat.diversity %>% 
    group_by(SITE) %>%
    dplyr::summarise(
      MEAN=mean(RELIEF_SIMPSONS_DIVERSITY),
      SD= sd(RELIEF_SIMPSONS_DIVERSITY),
      N=length(RELIEF_SIMPSONS_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  
  write.csv(dat.diversity, here("Data","Summarized data","UPC_DIVERSITY_Transect_Level_Summary.csv"))
  write.csv(dat.diversity.zone.summary, here("Data","Summarized data","UPC_DIVERSITY_Zone_Level_Summary.csv"))
  write.csv(dat.diversity.site.summary, here("Data","Summarized data","UPC_DIVERSITY_Site_Level_Summary.csv"))
  write.csv(dat.diversity.year.site.summary, here("Data","Summarized data","UPC_DIVERSITY_Year_Site_Level_Summary.csv"))

```

### Plots 
```{r,echo=FALSE}
dat.diversity.zone.summary %>%
  ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay"))) %>%
  ggplot(aes(x=YEAR, y =MEAN, group = as.factor(ZONE), color = as.factor(ZONE))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=MEAN -SE, ymax=MEAN+SE), width = 0.1)+
  facet_wrap(~SITE)+
  theme_classic() +
  ylab("Simpsons Diversity Index - UPC")


```

```{r,echo=FALSE}
dat.diversity.year.site.summary %>%
    ungroup() %>%
   mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay"))) %>%
  ggplot(aes(x=YEAR, y =MEAN))+
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=MEAN -SE, ymax=MEAN+SE), width = 0.1)+
  facet_wrap(~SITE) + 
  theme_classic() +
  ylab("Simpsons Diversity Index - UPC")
```

```{r,echo=FALSE}
dat.diversity.site.summary %>%
    ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay"))) %>%
  ggplot(aes(x=SITE, y =MEAN))+
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=MEAN -SE, ymax=MEAN+SE), width = 0.1)+
  theme_classic() +
  ylab("Simpsons Diversity Index - UPC")
```






 


 
