---
title: "Summarise Substrate Data"
author: "Jameal Samhouri"
date: "7/30/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---

# Summary

## Approach

This analysis represents a summary of the substrate data collected on uniform point contact (UPC) surveys of the Olympic Coast National Marine Sanctuary from 2016-2019. These annual surveys are conducted by NMFS and NMS staff at Destruction Island, Cape Johnson, Cape Alava, Tatoosh Island, and Neah Bay.

The outputs include data frames (.csv, .rds) and plots of the proportion of each substrate category (bedrock, boulder, cobble, sand) and the Simpson's diversity of susbtrate. They represent summaries at the following levels:
- year-site-side-zone-transect
- year-site-zone
- site-zone
- year-site
- site.

## Results overview

The predominant substrate types at each site are:

Site               | Dominant Substrate
-------------------|-------------------
Destruction Island | Bedrock
Cape Johnson       | Boulder
Cape Alava         | Boulder
Tatoosh Island     | Bedrock
Neah Bay           | Bedrock

While these are the dominant substrate types, there is a fair amount of spatial variability among transects and zones at each site, as well as interannual variability within each site. Specific examples include the wide range of proportion boulder across transects at Cape Alava in 2019, and trends showing we are surveying increasing boulder habitat from 2016-19 at Cape Johnson and Neah Bay. **Therefore, it may be useful to use both the mean and the variability in the proportion of each substrate category across a site as predictors of community composition or species abundance in future analyses.**

Most sites show similar substrate composition between the two depth zones (5m and 10m). Variability between depth zones in substrate composition at a site is much lower than variability among sites in substrate composition. 

**The diversity of substrate composition is highest at Cape Johnson, Cape Alava, and Neah Bay, and lowest at Destruction Island and Tatoosh Island. Substrate diversity is much higher in the 10m depth zone than the 5m depth zone at Destruction Island, Cape Johnson, and Tatoosh Island, but substrate diversity is similar in the 5m and 10m depth zones at Cape Alava and Neah Bay.**

We plan to consider the correlation between proportion cover of each substrate category in relation to the proportion cover of each relief category, as well as the correlation between substrate diversity and relief diversity. At a first cut, these analyses are probably most informative at the year-site-zone scale.


# Begin

Read in libraries and 2016-2019 data; 2015 UPC uses quadrats and is inconsistent with later methods.

```{r setup, message=FALSE, warning=FALSE}
#to load data make sure directory is set to knit from project directory

HomeFile = getwd()
DataLoc = paste0(HomeFile,"/Data/")
Fig_Loc = paste0(HomeFile,"/Figures/")

knitr::opts_chunk$set(
  fig.path = Fig_Loc,
  cache.path = 'cache/graphics-', 
  echo=FALSE,
  error=FALSE,
  include = TRUE,
  dev='png',
  dpi=300,
  warning=FALSE,
  #out.width = '100%',
  fig.align='center'
  
  )

# Libraries
library(tidyverse)
library(ggplot2)
library(viridis)
library(reshape2)
library(janitor)
library(magrittr)
library(vegan)

#base.dir <- "/Users/ole.shelton/GitHub/OCNMS"
base.dir <- HomeFile
data.dir <- DataLoc
data.summary.output.dir <- DataLoc

dat.2016.on.upc <- clean_names(read.csv(paste0(DataLoc,"NWFSC_UPC_ALLYEARS_data_2019.csv")), case = "all_caps") %>%
  remove_empty(c("rows", "cols")) 
# fix data entry issues
dat.2016.on.upc %<>% mutate(
  CLASSCODE = toupper(CLASSCODE)
)
unique(dat.2016.on.upc$CLASSCODE)
unique(dat.2016.on.upc$SITE)
glimpse(dat.2016.on.upc)

substrate.codes <- clean_names(read.csv("substrate_codes.csv"), case = "all_caps") %>%
  remove_empty(c("rows", "cols")) # js checked on 072720 and codes are same in 2019 data
  
substrate <- dat.2016.on.upc %>% filter(CATEGORY=="SUBSTRATE")
#relief <- dat.2016.on.upc %>% filter(CATEGORY=="RELIEF")
#cover <- dat.2016.on.upc %>% filter(CATEGORY=="COVER")

```

# Analyze substrate data

Summarise raw data into data frames that: 1) sum segments for each transect. 2) calculate summary stats for site-year-zone. 3) calculate summary stats for site-year. 4) calculate summary stats for site-zone. 5) calculate summary stats for site.

```{r}

# Pad with zero observations. i.e., complete the data so that each segment has all 4 substrate categories
  cat <- data.frame(merge(unique(substrate$SITE),unique(substrate$CLASSCODE))); colnames(cat) <- c("SITE","CLASSCODE")
  dat.sub <- substrate %>% 
    group_by(YEAR,SITE,SEGMENT,TRANSECT,SIDE,ZONE,OBSERVER) %>% 
    summarise(a=length(SEGMENT)) %>% 
    full_join(.,cat,by="SITE") %>% 
    dplyr::select(-a) %>% 
    left_join(.,substrate)
  dat.sub$COUNT[is.na(dat.sub$COUNT)==TRUE] <- 0
  
  # order sites from south to north
  dat.sub$SITE <- factor(dat.sub$SITE,
               levels=c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island","Neah Bay"))
  
  # make year a factor
  dat.sub$YEAR <- as.factor(dat.sub$YEAR)
  
  # not sure if we have to account for SIDE 
  # (eg, multiple observers on a segment)? looks like no.
  tmp<-substrate %>% 
    group_by(YEAR,SITE,SEGMENT,TRANSECT,SIDE,ZONE,OBSERVER) %>% 
    summarise(a=length(SEGMENT), b=length(OBSERVER))
  which(tmp$a != tmp$b) # integer(0)
  # with(dat.sub, table(SITE, ZONE, TRANSECT, SIDE))
  # unique(dat.sub$SIDE)
  # length(which(dat.sub$SIDE != 1))/dim(dat.sub)[1]
  
  # but it does look like there are some issues with duplicate entries. so fix those
  dat.sub <- dat.sub %>% 
    group_by(YEAR,SITE,ZONE,SIDE,TRANSECT) %>%
    distinct(CLASSCODE, SEGMENT, COUNT, .keep_all = TRUE) %>% # remove duplicate entries based on CLASSCODE, SEGMENT, COUNT
    mutate(
      TOTAL_COUNT = sum(COUNT),
      PROPORTION = COUNT/TOTAL_COUNT
    )
  unique(dat.sub$TOTAL_COUNT) # 30 29
  dat.sub[which(dat.sub$TOTAL_COUNT !=30),]
  glimpse(dat.sub)
  
  # create summary df. 1) sum segments for each transect. 2) calculate summary stats for site-year-zone. 3) calculate summary stats for site-year. 4) calculate summary stats for site-zone. 5) calculate summary stats for site
  #proportion data so use binomial distribution for summary stats. https://sites.warnercnr.colostate.edu/gwhite/wp-content/uploads/sites/73/2017/04/BinomialDistribution.pdf
  
  # 1a) sum segments for each transect to get total proportion of each substrate CLASSCODE
  sub.summary.year.transect <- dat.sub %>% 
    group_by(YEAR,SITE,ZONE,SIDE,TRANSECT,CLASSCODE) %>%
    dplyr::summarise(
      PROPORTION=sum(PROPORTION),
      PROPORTION_2=sum(PROPORTION)^2,
      .groups = "keep"
    )
  sub.summary.year.transect$YEAR <- as.factor(sub.summary.year.transect$YEAR)
  # order sites from south to north
  sub.summary.year.transect$SITE <- factor(sub.summary.year.transect$SITE,
                                       levels=c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island","Neah Bay"))
  glimpse(sub.summary.year.transect)
  
  # 1b) calculate substrate diversity for each transect
  sub.summary.year.transect.div <- sub.summary.year.transect %>%
    #dat.sub %>%
    #filter(!is.na(COUNT)) %>%
    #pivot_wider(names_from = CLASSCODE, values_from = COUNT, values_fill = list(COUNT=0)) 
  #sub.summary.year.transect.div2 <- sub.summary.year.transect.div1 %>%
    group_by(YEAR,SITE,ZONE,SIDE,TRANSECT) %>%
    dplyr::summarise(
      DIVERSITY= sum(PROPORTION_2),
      SUBSTRATE_DIVERSITY= 1 - DIVERSITY, # simpsons diversity
      #SUBSTRATE_DIVERSITY=diversity(sub.summary.year.transect.div1[,c('BEDRK', 'BOULD', 'SAND', 'COB')], index = "simpson"),
      .groups = "keep"
    ) %>%
    select(-DIVERSITY)
  glimpse(sub.summary.year.transect.div)
  
  # ggplot(data=sub.summary.year.transect.div) +
  #   geom_density(aes(SUBSTRATE_DIVERSITY)) +
  #   theme_bw()
  
  # 2a) calculate summary stats for site-year-zone
  ######## corrected SE ##################
  
  sub.summary.year.zone <- sub.summary.year.transect %>% 
    group_by(YEAR,SITE,ZONE,CLASSCODE) %>%
    dplyr::summarise(
      N=length(PROPORTION),
      MEAN=mean(PROPORTION),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      #SE=SD/sqrt(N),
      SE = sqrt(MEAN*(1-MEAN)/(N-1)),
      .groups = "keep"
      )
  
  ##########################################
  #sub.summary.year.zone$YEAR <- as.factor(sub.summary.year.zone$YEAR)
  # order sites from south to north
  sub.summary.year.zone$SITE <- factor(sub.summary.year.zone$SITE,
                         levels=c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island","Neah Bay"))
  glimpse(sub.summary.year.zone)
  
  # 2b) calculate substrate diversity for each site-year-zone
  sub.summary.year.zone.div <- sub.summary.year.transect.div %>% 
    group_by(YEAR,SITE,ZONE) %>%
    dplyr::summarise(
      MEAN=mean(SUBSTRATE_DIVERSITY),
      SD= sd(SUBSTRATE_DIVERSITY),
      N=length(SUBSTRATE_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  glimpse(sub.summary.year.zone.div)
  
  # 3a) calculate summary stats for site-year
  sub.summary.year <- sub.summary.year.transect %>% 
    group_by(YEAR,SITE,CLASSCODE) %>%
    dplyr::summarise(
      N=length(PROPORTION),
      MEAN=mean(PROPORTION),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N),
      .groups = "keep"
    )
  #sub.summary.year$YEAR <- as.factor(sub.summary.year$YEAR)
  sub.summary.year$SITE <- factor(sub.summary.year$SITE,
                                       levels=c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island","Neah Bay"))
  glimpse(sub.summary.year)
  
    # 3b) calculate substrate diversity for each site-year-zone
  sub.summary.year.div <- sub.summary.year.transect.div %>% 
    group_by(YEAR,SITE) %>%
    dplyr::summarise(
      MEAN=mean(SUBSTRATE_DIVERSITY),
      SD= sd(SUBSTRATE_DIVERSITY),
      N=length(SUBSTRATE_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  glimpse(sub.summary.year.div)
  
  # 4a) calculate summary stats for site-zone
  sub.summary.zone <- sub.summary.year.transect %>% 
    group_by(SITE,ZONE,CLASSCODE) %>%
    dplyr::summarise(
      N=length(PROPORTION),
      MEAN=mean(PROPORTION),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N),
      .groups = "keep"
    )
  sub.summary.zone$SITE <- factor(sub.summary.zone$SITE,
                                  levels=c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island","Neah Bay"))
  glimpse(sub.summary.zone)
  
  # 4b) calculate substrate diversity for each site-year-zone
  sub.summary.zone.div <- sub.summary.year.transect.div %>% 
    group_by(SITE,ZONE) %>%
    dplyr::summarise(
      MEAN=mean(SUBSTRATE_DIVERSITY),
      SD= sd(SUBSTRATE_DIVERSITY),
      N=length(SUBSTRATE_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  glimpse(sub.summary.zone.div)
  
  # 5a) calculate summary stats for site
  sub.summary.site <- sub.summary.year.transect %>% 
    group_by(SITE,CLASSCODE) %>%
    dplyr::summarise(
      N=length(PROPORTION),
      MEAN=mean(PROPORTION),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N),
      .groups = "keep"
    )
  sub.summary.site$SITE <- factor(sub.summary.site$SITE,
                                  levels=c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island","Neah Bay"))
  glimpse(sub.summary.site)
  
  # 5b) calculate substrate diversity for each site-year-zone
  sub.summary.site.div <- sub.summary.year.transect.div %>% 
    group_by(SITE) %>%
    dplyr::summarise(
      MEAN=mean(SUBSTRATE_DIVERSITY),
      SD= sd(SUBSTRATE_DIVERSITY),
      N=length(SUBSTRATE_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  glimpse(sub.summary.site.div)

  #### END BASIC SUBSTRATE SUMMARIES.



```


Write out the data frames we just made.


```{r }
### WRITE OUT SUBSTRATE DATA FRAMES
  
  
  write_csv(sub.summary.year.transect, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE-DEPTH-SIDE-TRANSECT 2016-2019.csv"))
  write_rds(sub.summary.year.transect, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE-DEPTH-SIDE-TRANSECT 2016-2019.rds"))
  write_csv(sub.summary.year.transect.div, paste0(data.summary.output.dir, "/SUBSTRATE diversity summary by YEAR-SITE-DEPTH-SIDE-TRANSECT 2016-2019.csv"))
  write_rds(sub.summary.year.transect.div, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE-DEPTH-SIDE-TRANSECT 2016-2019.rds"))
  
  write_csv(sub.summary.year.zone, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE-DEPTH 2016-2019.csv"))
  write_rds(sub.summary.year.zone, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE-DEPTH 2016-2019.rds"))
  write_csv(sub.summary.year.zone.div, paste0(data.summary.output.dir, "/SUBSTRATE diversity summary by YEAR-SITE-DEPTH 2016-2019.csv"))
  write_rds(sub.summary.year.zone.div, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE-DEPTH 2016-2019.rds"))
  
  write_csv(sub.summary.year, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE 2016-2019.csv"))
  write_rds(sub.summary.year, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE 2016-2019.rds"))
  write_csv(sub.summary.year.div, paste0(data.summary.output.dir, "/SUBSTRATE diversity summary by YEAR-SITE 2016-2019.csv"))
  write_rds(sub.summary.year.div, paste0(data.summary.output.dir, "/SUBSTRATE summary by YEAR-SITE 2016-2019.rds"))
  
  write_csv(sub.summary.zone, paste0(data.summary.output.dir, "/SUBSTRATE summary by SITE-DEPTH 2016-2019.csv"))
  write_rds(sub.summary.zone, paste0(data.summary.output.dir, "/SUBSTRATE summary by SITE-DEPTH 2016-2019.rds"))
  write_csv(sub.summary.zone.div, paste0(data.summary.output.dir, "/SUBSTRATE diversity summary by SITE-DEPTH 2016-2019.csv"))
  write_rds(sub.summary.zone.div, paste0(data.summary.output.dir, "/SUBSTRATE summary by SITE-DEPTH 2016-2019.rds"))
  
  write_csv(sub.summary.site, paste0(data.summary.output.dir, "/SUBSTRATE summary by SITE 2016-2019.csv"))
  write_rds(sub.summary.site, paste0(data.summary.output.dir, "/SUBSTRATE summary by SITE 2016-2019.rds"))
  write_csv(sub.summary.site.div, paste0(data.summary.output.dir, "/SUBSTRATE diversity summary by SITE 2016-2019.csv"))
  write_rds(sub.summary.site.div, paste0(data.summary.output.dir, "/SUBSTRATE diversity summary by SITE 2016-2019.rds"))
  
  
  #### END WRITING OUT SUBSTRATE DATA FRAMES
```


# Make substrate plots

## 1) plots for site-year-zone.

```{r}

  # a) substrate type on x axis, site as facets
  sub.plot.year.zone_sub <- ggplot() +
    geom_jitter(data=sub.summary.year.transect,aes(y=PROPORTION,x=CLASSCODE,colour=YEAR),alpha=0.5, position=position_dodge(width=0.5))+
    geom_point(data=sub.summary.year.zone,aes(y=MEAN,x=CLASSCODE,fill=YEAR),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=CLASSCODE, colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(SITE~ZONE) +
    theme_bw()
 plot(sub.plot.year.zone_sub)

  # b) site on x axis, substrate type as facets
  sub.plot.year.zone_site <- ggplot() +
    geom_jitter(data=sub.summary.year.transect,aes(y=PROPORTION,x=SITE, colour=YEAR),alpha=0.5, position=position_dodge(width=0.5) )+
    geom_point(data=sub.summary.year.zone,aes(y=MEAN,x=SITE,fill=YEAR),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(CLASSCODE~ZONE) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.year.zone_site)
  
  sub.plot.year.zone_site_div <- ggplot() +
    geom_jitter(data=sub.summary.year.transect.div,aes(y=SUBSTRATE_DIVERSITY,x=SITE, colour=YEAR),alpha=0.5, position=position_dodge(width=0.5) )+
    geom_point(data=sub.summary.year.zone.div,aes(y=MEAN,x=SITE,fill=YEAR),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.zone.div,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(.~ZONE) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.year.zone_site_div)
  
  # c) compare zones, years as facets, lines connecting 5m v 10m
  
  sub.plot.year.zone_compare1 <- ggplot() +
    geom_jitter(data=sub.summary.year.transect,aes(y=PROPORTION,x=ZONE, colour=SITE),alpha=0.5, position=position_dodge(width=0.5))+
    geom_point(data=sub.summary.year.zone,aes(y=MEAN,x=ZONE,colour=SITE),size=2, position=position_dodge(width=0.5))+
    geom_line(data=sub.summary.year.zone,aes(y=MEAN,x=ZONE,colour=SITE),size=1)+
    geom_errorbar(data=sub.summary.year.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=ZONE,colour=SITE),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    facet_grid(YEAR~CLASSCODE) +
    theme_bw()
  plot(sub.plot.year.zone_compare1)
  
  # d) compare zones, CLASSCODE as facets, no lines connecting 5m v 10m
  sub.plot.year.zone_compare2 <- ggplot() +
    geom_jitter(data=sub.summary.year.transect,aes(y=PROPORTION,x=ZONE,colour=YEAR),alpha=0.3 , position=position_dodge(width=0.5))+
    geom_point(data=sub.summary.year.zone,aes(y=MEAN,x=ZONE,fill=YEAR),size=2,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=ZONE,colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d() +
    scale_fill_viridis_d() +
    facet_grid(CLASSCODE~SITE) +
    theme_bw()
  plot(sub.plot.year.zone_compare2)
  
  sub.plot.year.zone_compare2_div <- ggplot() +
    geom_jitter(data=sub.summary.year.transect.div,aes(y=SUBSTRATE_DIVERSITY,x=ZONE,colour=YEAR),alpha=0.3 , position=position_dodge(width=0.5))+
    geom_point(data=sub.summary.year.zone.div,aes(y=MEAN,x=ZONE,fill=YEAR),size=2,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.zone.div,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=ZONE,colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d() +
    scale_fill_viridis_d() +
    facet_grid(.~SITE) +
    theme_bw()
  plot(sub.plot.year.zone_compare2_div)
```
```{r abiotic-substrate, fig.width=6, fig.height=4}

par(ps=10, cex=1)

zone.labs = c('5 m', '10 m')
names(zone.labs) <- c('5','10')

sub.plot.year.zone_site <- ggplot() +
    geom_jitter(data=sub.summary.year.transect,
                aes(y=PROPORTION,x=SITE, colour=YEAR),alpha=0.5, position=position_dodge(width=0.5) )+
    geom_point(data=sub.summary.year.zone,
                aes(y=MEAN,x=SITE,fill=YEAR),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.zone,
                  aes(ymin=MEAN-1.96*SE,ymax=MEAN+1.96*SE,x=SITE, colour=YEAR),width=0.1, 
                  position=position_dodge(width=0.5))  +
    scale_x_discrete(expand = c(0,0), labels=c("DI","CJ","CA","TI","NB"))+  
    # scale_fill_manual(values = color4)+ 
    # scale_color_manual(values = rep('black',4))+
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(CLASSCODE~ZONE , labeller = labeller(ZONE = zone.labs)) +
    theme_bw() +
    theme( strip.text.x = element_text( size = 10))
    
  
plot(sub.plot.year.zone_site)

```

## 2) plots for site-year. 
```{r}
# a) substrate type on x axis, site as facets
  sub.plot.year_sub <- ggplot() +
    geom_point(data=sub.summary.year,aes(y=MEAN,x=CLASSCODE,fill=YEAR),colour="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=CLASSCODE, colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(SITE~.) +
    theme_bw()
  plot(sub.plot.year_sub)
  

  # b) site on x axis, substrate type as facets
  sub.plot.year_site <- ggplot() +
    geom_point(data=sub.summary.year,aes(y=MEAN,x=SITE,fill=YEAR),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(CLASSCODE~.) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.year_site)
  
  sub.plot.year_site_div <- ggplot() +
    geom_point(data=sub.summary.year.div,aes(y=MEAN,x=SITE,fill=YEAR),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.div,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=YEAR),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    #facet_grid(CLASSCODE~.) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.year_site_div)
  
  # c) time series, CLASSCODE as facets
  
  sub.plot.year_ts1 <- ggplot() +
    geom_point(data=sub.summary.year,aes(y=MEAN,x=YEAR,colour=SITE, group=SITE),size=2, position=position_dodge(width=0.5))+
    geom_line(data=sub.summary.year,aes(y=MEAN,x=YEAR,colour=SITE, group=SITE),size=1, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=YEAR,colour=SITE, group=SITE),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    facet_grid(.~CLASSCODE) +
    theme_bw()
  plot(sub.plot.year_ts1)
  
  # d) time series, SITE as facets
  
  sub.plot.year_ts2 <- ggplot() +
    geom_point(data=sub.summary.year,aes(y=MEAN,x=YEAR,colour=CLASSCODE, group=CLASSCODE),size=2, position=position_dodge(width=0.5))+
    geom_line(data=sub.summary.year,aes(y=MEAN,x=YEAR,colour=CLASSCODE, group=CLASSCODE),size=1, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=YEAR,colour=CLASSCODE, group=CLASSCODE),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    facet_grid(.~SITE) +
    theme_bw()
  plot(sub.plot.year_ts2)
  
  sub.plot.year_ts2_div <- ggplot() +
    geom_point(data=sub.summary.year.div,aes(y=MEAN,x=YEAR,colour=SITE, group=SITE),size=2, position=position_dodge(width=0.5))+
    geom_line(data=sub.summary.year.div,aes(y=MEAN,x=YEAR,colour=SITE, group=SITE),size=1, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.year.div,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=YEAR,colour=SITE, group=SITE),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    #facet_grid(.~SITE) +
    theme_bw()
  plot(sub.plot.year_ts2_div)
```


## 3) plots for site-zone. 

```{r}
  # a) substrate type on x axis, site as facets
  sub.plot.zone_sub <- ggplot() +
    geom_point(data=sub.summary.zone,aes(y=MEAN,x=CLASSCODE,fill=factor(ZONE)),colour="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=CLASSCODE, colour=factor(ZONE)),width=0.1, position=position_dodge(width=0.5))  +
    labs(fill='ZONE', colour='ZONE') +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(SITE~.) +
    theme_bw()
  plot(sub.plot.zone_sub)
  
  
  # b) site on x axis, substrate type as facets
  sub.plot.zone_site <- ggplot() +
    geom_point(data=sub.summary.zone,aes(y=MEAN,x=SITE,fill=factor(ZONE)),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=factor(ZONE)),width=0.1, position=position_dodge(width=0.5))  +
    labs(fill='ZONE', colour='ZONE') +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(CLASSCODE~.) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.zone_site)
  
  sub.plot.zone_site_div <- ggplot() +
    geom_point(data=sub.summary.zone.div,aes(y=MEAN,x=SITE,fill=factor(ZONE)),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.zone.div,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=factor(ZONE)),width=0.1, position=position_dodge(width=0.5))  +
    labs(fill='ZONE', colour='ZONE') +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    #facet_grid(CLASSCODE~.) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.zone_site_div)
  
  # c) compare zones, years as facets, lines connecting 5m v 10m
  
  sub.plot.zone_compare1 <- ggplot() +
    geom_point(data=sub.summary.zone,aes(y=MEAN,x=factor(ZONE),colour=SITE, group=SITE),size=2, position=position_dodge(width=0.5))+
    geom_line(data=sub.summary.zone,aes(y=MEAN,x=factor(ZONE),colour=SITE, group=SITE),size=1, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=factor(ZONE),colour=SITE, group=SITE),width=0.1, position=position_dodge(width=0.5))  +
    xlab("ZONE") +
    scale_colour_viridis_d()+
    facet_grid(.~CLASSCODE) +
    theme_bw()
  plot(sub.plot.zone_compare1)
  
  sub.plot.zone_compare1_div <- ggplot() +
    geom_point(data=sub.summary.zone.div,aes(y=MEAN,x=factor(ZONE),colour=SITE, group=SITE),size=2, position=position_dodge(width=0.5))+
    geom_line(data=sub.summary.zone.div,aes(y=MEAN,x=factor(ZONE),colour=SITE, group=SITE),size=1, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.zone.div,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=factor(ZONE),colour=SITE, group=SITE),width=0.1, position=position_dodge(width=0.5))  +
    xlab("ZONE") +
    scale_colour_viridis_d()+
    #facet_grid(.~CLASSCODE) +
    theme_bw()
  plot(sub.plot.zone_compare1_div)
  
  # d) compare zones, CLASSCODE as facets, no lines connecting 5m v 10m
  
  sub.plot.zone_compare2 <- ggplot() +
    geom_point(data=sub.summary.zone,aes(y=MEAN,x=factor(ZONE),colour=CLASSCODE, group=CLASSCODE),size=2, position=position_dodge(width=0.5))+
    geom_line(data=sub.summary.zone,aes(y=MEAN,x=factor(ZONE),colour=CLASSCODE, group=CLASSCODE),size=1, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.zone,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=factor(ZONE),colour=CLASSCODE, group=CLASSCODE),width=0.1, position=position_dodge(width=0.5))  +
    xlab("ZONE") +
    scale_colour_viridis_d()+
    facet_grid(.~SITE) +
    theme_bw()
  plot(sub.plot.zone_compare2)
  

```


  
## 4) plots for site

```{r}  
  # a) substrate type on x axis, site as facets
  sub.plot.site_sub <- ggplot() +
    geom_point(data=sub.summary.site,aes(y=MEAN,x=CLASSCODE, fill=CLASSCODE),colour="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.site,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=CLASSCODE, colour=CLASSCODE),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(SITE~.) +
    theme_bw()
  plot(sub.plot.site_sub)
  
  
  # b) site on x axis, substrate type as facets
  sub.plot.site_site <- ggplot() +
    geom_point(data=sub.summary.site,aes(y=MEAN,x=SITE,fill=SITE),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.site,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=SITE),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    facet_grid(CLASSCODE~.) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.site_site)
  
  sub.plot.site_site_div <- ggplot() +
    geom_point(data=sub.summary.site.div,aes(y=MEAN,x=SITE,fill=SITE),color="black",size=3,shape=21, position=position_dodge(width=0.5))+
    geom_errorbar(data=sub.summary.site.div,
                  aes(ymin=MEAN-SE-1e-10,ymax=MEAN+SE,x=SITE, colour=SITE),width=0.1, position=position_dodge(width=0.5))  +
    scale_colour_viridis_d()+
    scale_fill_viridis_d()+
    #facet_grid(CLASSCODE~.) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust=1)
    )
  plot(sub.plot.site_site_div)
  
```

PRINT OUT ALL OF THE PLOTS IN A SINGLE PDF.

```{r}


  
  # pdf(file=paste0(base.dir,"/Plots/Substrate v3.pdf"),onefile=T)
  # 
  # print(sub.plot.site_sub)
  # print(sub.plot.site_site)
  # plot(sub.plot.site_site_div)
  # 
  # print(sub.plot.zone_sub)
  # print(sub.plot.zone_site)
  # print(sub.plot.zone_compare1)
  # print(sub.plot.zone_compare2)
  # print(sub.plot.zone_site_div)
  # print(sub.plot.zone_compare1_div)
  # 
  # print(sub.plot.year_sub)
  # print(sub.plot.year_site)
  # print(sub.plot.year_ts1)
  # print(sub.plot.year_ts2)
  # print(sub.plot.year_site_div)
  # print(sub.plot.year_ts2_div)
  # 
  # print(sub.plot.year.zone_sub)  
  # print(sub.plot.year.zone_site)
  # print(sub.plot.year.zone_compare1)
  # print(sub.plot.year.zone_compare2)
  # print(sub.plot.year.zone_site_div)
  # print(sub.plot.year.zone_compare2_div)
  # 
  # invisible(dev.off())
  
```

