---
title: "OCNMS exploratory plots"
author: "Ole Shelton"
date: "November 30, 2015"
output: pdf_document
---

This is a basic file for processing and providing basic plots for the 2015 Olympic Coast National Marine Sanctuary Dive Surveys.

```{r load, message=FALSE, echo=FALSE, warning=FALSE}
  library(rgdal)
  library(ggplot2)
  library(maps)
  library(mapdata)
  library(mapproj)
  library(reshape2)
  library(RColorBrewer)
  source("theme_acs.R")
 # This means don't display the r results in the output file
 # Read in the data, fix some of the data inconsistencies and make summary data.frames and lists

 ## This section is for breaking the dataframe into multiple components
  # quadrats: sessile, and algae
  # quatrats: mobile countable inverts
  # transects: mobile inverts
  # transects: algae stipes
  # transects: fish

 # a useful function
length.unique <-function(X){Y<-length(unique(X)); return(Y)}

 # Go get the NOAA survey data from the appropriate directory.

dat <- read.csv("Data/csv files/OCNMSDataComplete_master.csv")

dat.kvitek <- list()
dat.kvitek$Invert.quad   <- read.csv("Data/csv files/Kvitek.Data.Summary.Invert.quadrats.csv")
dat.kvitek$Invert.trans  <- read.csv("Data/csv files/Kvitek.Data.Summary.Invert.transects.csv")
dat.kvitek$Veg           <- read.csv("Data/csv files/Kvitek.Data.Summary.Veg.csv")

 # Do some name fixing
dat$Site[dat$Site == "Tatoosh"] <- "Tatoosh Island"
dat$Species <-  as.character(dat$Species)
dat$Species[dat$Species == "Mesocentrotus purpuratus"] <- "purple urchin"
dat$Species[dat$Species == "Halocynthia aurantium"]     <- "sea peach"
dat$Species[dat$Species == "Mimulus"]     <- "foliose kelp crab"
dat$Species[dat$Species == "coralline algae"]     <- "Coralline encrusting"
dat$Species[dat$Species == "cobble"]     <- "Cobble"
dat$Species[dat$Species == "tegula"]     <- "Tegula"
dat$Species[dat$Species == "styela"]     <- "Styela"
dat$Species[dat$Species == "sponge"]     <- "Sponge"
dat$Species[dat$Species == "epiactis prolifera"]     <- "Epiactis prolifera"
dat$Species[dat$Species == "Balanophyllia"]     <- "orange cup coral"
dat$Species[dat$Species == "tunicate"]     <- "Tunicate"

 ### Read in species lists for dividing species into algae, inverts, and fish.
sp.names <- read.csv("Data/csv files/OCNMSDataComplete_sp_names.csv")

 #Get rid of "trace"
dat$Percent.cover[dat$Percent.cover =="trace"] <- 1
dat$Percent.cover<- as.numeric(as.character(dat$Percent.cover))

 #QUADRAT
dat.quad <- aggregate(dat$Percent.cover,by=list(Observer=dat$Observer,Site=dat$Site,Transect=dat$Transect,Quadrat=dat$Quadrat,Species=dat$Species), sum,na.rm=T)
dat.quad <- dat.quad[dat.quad$x>0,]

samp.size.quad <- aggregate(dat.quad$Quadrat,
                            by=list(Site=dat.quad$Site,Transect=dat.quad$Transect,Observer=dat.quad$Observer),length.unique)
samp.size.site.quad <- aggregate(samp.size.quad$x,by=list(Site=samp.size.quad$Site),sum) 
samp.size.quad <- samp.size.quad[order(samp.size.quad$Site),]

 #### Quadrat Counts
dat.quad.count <- aggregate(dat$Count,by=list(Observer=dat$Observer,Site=dat$Site,Transect=dat$Transect,Quadrat=dat$Quadrat,Species=dat$Species), sum,na.rm=T)
dat.quad.count <- dat.quad.count[dat.quad.count$x>0,]
samp.size.quad.count <- aggregate(dat.quad.count$Quadrat,
                             by=list(Site=dat.quad.count$Site,Transect=dat.quad.count$Transect,Observer=dat.quad.count$Observer),length.unique)
samp.size.site.quad.count <- aggregate(samp.size.quad.count$x,by=list(Site=samp.size.quad.count$Site),sum) 

 ### Transect Counts
dat.trans <-dat[is.na(dat$Quadrat)==T,]
    dat.trans$NUMB <- 1:nrow(dat.trans)
    
    #### Merge group names in with transects
    dat.trans.1 <- merge(dat.trans,sp.names,by.x="Species",by.y="drop.down")
    dat.trans.1 <- dat.trans.1[order(dat.trans.1$NUMB),]

    temp<- dat.trans[is.na(match(dat.trans$NUMB,dat.trans.1$NUMB))==T,]
    if(nrow(temp)>0){
      print("ERROR ERROR: Unmatched species names")
      stop()
    }

    qual.trans   <- dat.trans.1[is.na(dat.trans.1$Categorical.abundance)==F,]
    dat.trans.1  <- dat.trans.1[is.na(dat.trans.1$Categorical.abundance)==T,]

dat.trans.fish  <- list()
dat.trans.fish$All.fish <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Fish",]

dat.trans.algae  <- list()
dat.trans.algae$All.algae <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Brown algae",]

dat.trans.mammal.bird <- list()
dat.trans.mammal.bird$Bird    <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Birds",]
dat.trans.mammal.bird$Mammal  <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Mammals",]

dat.trans.invert  <- list()
dat.trans.invert$All.invert       <- dat.trans.1[dat.trans.1$Broad.taxonomic.category != "Fish" &
                                           dat.trans.1$Broad.taxonomic.category != "Birds" &
                                           dat.trans.1$Broad.taxonomic.category != "Mammals" &
                                           dat.trans.1$Broad.taxonomic.category != "Brown algae",]
dat.trans.invert$Urchin           <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Urchin",]
dat.trans.invert$Cucumber         <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Sea cucumbers",]
dat.trans.invert$Sea.star         <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Seastars",]
dat.trans.invert$Crab             <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Crabs",]
dat.trans.invert$Chiton           <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Chiton",]
dat.trans.invert$Limpet           <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Limpet",]
dat.trans.invert$Gastropod        <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Gastropod",]
dat.trans.invert$Bivalve          <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Bivalve" ,]
dat.trans.invert$Nudi             <- dat.trans.1[dat.trans.1$Broad.taxonomic.category == "Nudi",]

 ### samp.size.transect
  samp.size.trans <- list()
    samp.size.trans$Invert <-aggregate(dat.trans.invert$All.invert$Transect,
          by=list(Site=dat.trans.invert$All.invert$Site,Observer=dat.trans.invert$All.invert$Observer),length.unique)
    samp.size.trans$Fish <-aggregate(dat.trans.fish$All.fish$Transect,
          by=list(Site=dat.trans.fish$All.fish$Site,Observer=dat.trans.fish$All.fish$Observer),length.unique)

 ###  Make a Species list for each type of data.
sp.quad.list<- list()
sp.quad.count.list<- list()
sp.trans.list <- list()
SITE          <- sort(unique(dat.quad$Site))

sp.quad       <-  data.frame(sp.names = sort(unique(dat.quad$Species)))
  sp.quad$Category  <- sp.names$Broad.taxonomic.category[match(sp.quad$sp.names,sp.names$drop.down)]
sp.quad.count <-  data.frame(sp.names = sort(unique(dat.quad.count$Species)))
  sp.quad.count$Category  <- sp.names$Broad.taxonomic.category[match(sp.quad.count$sp.names,sp.names$drop.down)]
sp.trans      <-  data.frame(sp.names = sort(unique(dat.trans$Species)))
  sp.trans$Category  <- sp.names$Broad.taxonomic.category[match(sp.trans$sp.names,sp.names$drop.down)]

for(i in 1:length(SITE)){
    
    a <- match(sp.quad$sp.names,sort(unique(dat.quad$Species[dat.quad$Site==SITE[i]])))
    sp.quad[,as.character(SITE[i])] <- 0
    sp.quad[which(is.na(a)==F),as.character(SITE[i])] <- 1

    b <- match(sp.quad.count$sp.names,sort(unique(dat.quad.count$Species[dat.quad.count$Site==SITE[i]])))
    sp.quad.count[,as.character(SITE[i])] <- 0
    sp.quad.count[which(is.na(b)==F),as.character(SITE[i])] <- 1

    d <- match(sp.trans$sp.names,sort(unique(dat.trans$Species[dat.trans$Site==SITE[i]])))
    sp.trans[,as.character(SITE[i])] <- 0
    sp.trans[which(is.na(d)==F),as.character(SITE[i])] <- 1
}
```

## The important data.frames and lists for 2015 data are these:

* "dat.quad" has all of the pecent cover data associated with quadrat counts
* "dat.quad.count" has all of the count data of mobile and countable inverts from the quadrat surveys
* "dat.trans.alge" has all the transect observations for algae an
* "dat.trans.mammal.bird" has all the mammal and bird observations
* "dat.trans.fish" has all of the transect observations for fish species.
* "dat.trans.invert" has all of the transect observations for invertebrates.  I made this into a list that has multiple entries for species folk are likely interested in
```{r}
names(dat.trans.invert)
```
* "dat.kvitek" is a list with the historical data available from the 1987,1995, and 1999 surveys.  There are a couple of different kinds of surveys.
```{r}
names(dat.kvitek)
```

All surveys were conducted on these dates:
```{r,echo=FALSE}
 sort(as.character(unique(dat$Date)))
```

\newpage
Here is a map showing the approximate locations of the sites

```{r, echo=FALSE}
  #### READ IN WASHINGTON MARINE SHORELINES (FROM http://www.ecy.wa.gov/services/gis/data/data.htm
  
  #base.dir <- getwd()
#   gis.dir<- "/Users/ole.shelton/GitHub/OCNMS/Data/PG Shore shore_thematic_UTM10/"
#   setwd(gis.dir)
#   # shore.poly	<-	readShapePoly("shore_thematic_UTM10.shp")
#   
#   shore.poly	<-	readOGR(dsn=".","shore_thematic_UTM10")
#   shore		<-	data.frame(shore.poly)
# 
#   shore.dat	<-	 as(shore.poly,"data.frame")
#   shore.dat$NEW	<- 1:nrow(shore.dat)
#   shore.poly@data$id 	= rownames(shore.poly@data)
#   shore.points		<-	fortify(shore.poly,region="id")
#   shore.df			  <- merge(shore.points,shore.poly@data,by="id")
#   write.csv(shore.df,file="shore df.csv",row.names=F)
  shore.df        <- read.csv("Data/PG Shore shore_thematic_UTM10/shore df.csv")
  
  # setwd(base.dir)
  PROJ <- "+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
  loc <-  read.csv("Data/csv files/OCNMSDataComplete_site_loc.csv")
  loc.new				<- cbind(-loc$Longitude, loc$Latitude)
  loc.new.sp		<- SpatialPoints(loc.new,CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))

  loc.new.sp.NAD83	<- spTransform(loc.new.sp, CRS(PROJ))
  new.dat			      <- as(loc.new.sp.NAD83,"data.frame")
  colnames(new.dat)	<- c("LonUTM","LatUTM")
  loc <- data.frame(loc,new.dat)
  loc$Number <- c(2,1,3:10)
  
  x.lim =c(3.3e+05,4.1e+05)
  y.lim =c(52.7e+05,53.7e+05)
  
  RATIO <- (y.lim[2]-y.lim[1])/ (x.lim[2]-x.lim[1])
  FIG.WIDTH <- 5.2
  FIG.HEIGHT <- RATIO * FIG.WIDTH 
```  
  
```{r,fig.width=FIG.WIDTH, fig.height= FIG.HEIGHT , echo=FALSE} 
  #wa <- map_data('worldHires','USA')
  #wa <- map_data("state",region = c('washington'))
  p3	<-	ggplot()+
				theme_acs() +
				theme(legend.key =  element_rect(fill = "white", colour = "white"),
					panel.border = element_rect(fill = NA, colour = "black",size=1.5)) +
		  #  geom_polygon(data=wa,aes(long,lat,group=group),fill=8,color="black")	+
				coord_cartesian(xlim = x.lim,ylim=y.lim)     +
        geom_polygon(data=shore.df[shore.df$POLYTYPE == "Land" | shore.df$POLYTYPE == "Island" ,], fill=grey(0.7),color=NA) +
					aes(long,lat,group=group) +
        xlab("Eastings (UTM)") +
			  ylab("Northings (UTM)") +
        geom_point(data=loc,aes(LonUTM,LatUTM,group=1),color="red",size=3)
	
  p3 <- p3 + annotate("text", x=loc$LonUTM - c(1e4,15e3,9e3,1e4,9e3,12e3,1e4,1e4,-9e3,-8e3), y= loc$LatUTM + c(rep(0,8),3e3,0), label= loc$Site.name,size=3.5)
  print(p3)
```

\newpage
## Methods
Four divers conducted dive surveys at each site (two teams of two). Each diver pair surveyed two 30m long transects. All dives in 2015 were conducted at about 6-8m depth.

I've focused on summarizing the species observed in at each site (for the quadrats, invertebrate transects, and fish transects)

### Sessile creatures (and some benthic categories)
```{r,echo=FALSE,warning=F,message=F} 
  quad.tile        <- melt(sp.quad)
    colnames(quad.tile)[3] <- "Site"
    quad.tile <- merge(quad.tile,loc[,c("Site.name","Number")],by.x="Site",by.y="Site.name")
    quad.tile <- quad.tile[order(quad.tile$Number,quad.tile$Category,quad.tile$sp.names,decreasing=T),]
    
  quad.count.tile  <- melt(sp.quad.count)
    colnames(quad.count.tile)[3] <- "Site"
    quad.count.tile <- merge(quad.count.tile,loc[,c("Site.name","Number")],by.x="Site",by.y="Site.name")
    quad.count.tile <- quad.count.tile[order(quad.count.tile$Number,quad.count.tile$Category,quad.count.tile$sp.names,decreasing=T),]

  trans.tile       <- melt(sp.trans)
    colnames(trans.tile)[3] <- "Site"
    trans.tile <- merge(trans.tile,loc[,c("Site.name","Number")],by.x="Site",by.y="Site.name")
    trans.tile <- trans.tile[order(trans.tile$Number,trans.tile$Category,trans.tile$sp.names,decreasing=T),]
```

```{r,fig.width=7, fig.height= 10, echo=FALSE,warning=F,message=F} 
  COL <- brewer.pal(6,"Greens")[6]
  
  quad.p <- ggplot(data = quad.tile, aes(x=Number, y=sp.names, fill=value)) + 
    scale_fill_gradient(low="white",high=COL,guide="none")+
    geom_tile(color="white") +
    xlab("Site") +
	  ylab("Species") +
    scale_x_discrete( breaks=1:nrow(loc),labels= loc$Site.name[order(loc$Number)]) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(quad.p)
```

```{r,fig.width=7, fig.height= 8, echo=FALSE} 
  quad.count.p <- ggplot(data = quad.count.tile, aes(x=Number, y=sp.names, fill=value)) +
   scale_fill_gradient(low="white",high=COL,guide="none")+
    geom_tile(color="white") +
    xlab("Site") +
	  ylab("Species") +
    scale_x_discrete( breaks=1:nrow(loc),labels= loc$Site.name[order(loc$Number)]) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #print(quad.count.p)
```

```{r,fig.width=7, fig.height= 8, echo=FALSE,warning=F} 
  trans.p <- ggplot(data = trans.tile, aes(x=Number, y=sp.names, fill=value)) +    
   scale_fill_gradient(low="white",high=COL,guide="none")+
    geom_tile(color="white") +
    xlab("Site") +
	  ylab("Species") +
    scale_x_discrete( breaks=1:nrow(loc),labels= loc$Site.name[order(loc$Number)]) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #print(trans.p)
```

### Fish species on transects
```{r,fig.width=7, fig.height= 8, echo=FALSE,warning=F} 
  trans.p <- ggplot(data = trans.tile[trans.tile$Category=="Fish",], aes(x=Number, y=sp.names, fill=value)) +    
   scale_fill_gradient(low="white",high=COL,guide="none")+
    geom_tile(color="white") +
    xlab("Site") +
	  ylab("Species") +
    scale_x_discrete( breaks=1:nrow(loc),labels= loc$Site.name[order(loc$Number)]) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(trans.p)
```

### Mobile invertebrate species on transects
```{r,fig.width=7, fig.height= 8, echo=FALSE,warning=F} 
  trans.p <- ggplot(data = trans.tile[trans.tile$Category=="Crabs" | trans.tile$Category=="Urchin"  | trans.tile$Category=="Chiton" |
                                            trans.tile$Category=="Sea cucumbers" | trans.tile$Category=="Seastars" |
                                            trans.tile$Category=="Limpet" |trans.tile$Category=="Gastropod" ,], 
                    aes(x=Number, y=sp.names, fill=value)) +    
   scale_fill_gradient(low="white",high=COL,guide="none")+
    geom_tile(color="white") +
    xlab("Site") +
	  ylab("Species") +
    scale_x_discrete( breaks=1:nrow(loc),labels= loc$Site.name[order(loc$Number)]) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  print(trans.p)
```

\newpage
## Comparing mobile invertebrates
Data from all sites in all years.  I include only data from 6m and shallower from historical data.

```{r,echo=F}
  # First merge the kvitek and 2015 survey.
  # Only include sites that are directly comparable.
  
  dat.mob.invert  <- expand.grid( Site=unique(loc$Site.name),Year=c(1987,1995,1999))
  dat.mob.invert$match.2015.name <- dat.mob.invert$Site
  dat.mob.invert  <- merge(dat.mob.invert,
                           dat.kvitek$Invert.quad[dat.kvitek$Invert.quad$match.2015.name != "",c(3,10:28)],
                           by=c("match.2015.name","Year"),all=T)
  # dat.mob.invert <- 
  
  
  # Calculate summaries for invertebrate fauna.
   All.transect <- aggregate(dat.trans.invert$All.invert$Transect,
                             by=list(Site=dat.trans.invert$All.invert$Site,
                                Observer=dat.trans.invert$All.invert$Observer,
                                Transect=dat.trans.invert$All.invert$Transect),length.unique)  
  All.transect<-All.transect[order(All.transect$Site),]
    
  # Urchins 
     temp <- aggregate(dat.trans.invert$Urchin$Count,
                       by=list(Site=dat.trans.invert$Urchin$Site,
                               Transect=dat.trans.invert$Urchin$Transect,
                               Observer=dat.trans.invert$Urchin$Observer,
                               Area=dat.trans.invert$Urchin$Transect.area..inverts.),sum)  
    temp$Area <- as.numeric(as.character(temp$Area))
     temp$density.m2 <- temp$x / temp$Area
  
    All.transect.temp <- merge(All.transect,temp[,c("Site", "Observer", "Transect","density.m2")],all=T)
    All.transect.temp[is.na(All.transect.temp ==T)] <- 0
    Mean <- aggregate(All.transect.temp$density.m2,by=list(Site=All.transect.temp$Site),mean)
    SD   <-   aggregate(All.transect.temp$density.m2,by=list(Site=All.transect.temp$Site),sd)
    
    temp.merge <- cbind(Mean,SD[,2])  
    NAME <- names(dat.mob.invert)[grep("urchin",names(dat.mob.invert))]
    colnames(temp.merge) <- c("match.2015.name",NAME)
    all.merge <- temp.merge
  
  page.name <- c("Cucumber","Crab","Bivalve","Limpet","Gastropod","Nudi")  
  col.nom <- c("cuc","crab","scal","limp","gast","nudi")
          
   #remaining species 
   for(i in 1:length(page.name)){
     temp <- aggregate(dat.trans.invert[[page.name[i]]]$Count,
                       by=list(Site=dat.trans.invert[[page.name[i]]]$Site,
                               Transect=dat.trans.invert[[page.name[i]]]$Transect,
                               Observer=dat.trans.invert[[page.name[i]]]$Observer,
                               Area=dat.trans.invert[[page.name[i]]]$Transect.area..inverts.),sum)  
    temp$Area <- as.numeric(as.character(temp$Area))
    temp$density.m2 <- temp$x / temp$Area
  
    All.transect.temp <- merge(All.transect,temp[,c("Site", "Observer", "Transect","density.m2")],all=T)
    All.transect.temp[is.na(All.transect.temp ==T)] <- 0
    Mean <- aggregate(All.transect.temp$density.m2,by=list(Site=All.transect.temp$Site),mean)
    SD   <-   aggregate(All.transect.temp$density.m2,by=list(Site=All.transect.temp$Site),sd)
    
    temp.merge <- cbind(Mean,SD[,2])  
    NAME <- names(dat.mob.invert)[grep(col.nom[i],names(dat.mob.invert))]
    
    colnames(temp.merge) <- c("match.2015.name",NAME)
    all.merge<-cbind(all.merge,temp.merge[,2:3])
  } 

    all.merge$Year <- 2015

    dat.mob.invert <- merge(dat.mob.invert,all.merge,all=T)
    dat.mob.invert$Site <- dat.mob.invert$match.2015.name
    dat.mob.invert  <- merge(dat.mob.invert,loc[,c("Site.name","Number")],by.x="Site",by.y="Site.name",all=T)

    dat.mob.invert$Site.numb <- factor(paste(dat.mob.invert$Number-1,dat.mob.invert$Site,sep="."))
    dat.mob.invert <- dat.mob.invert[order(dat.mob.invert$Site.numb),]
    
  FIG.WIDTH  <- 8
  FIG.HEIGHT <- 9   
  COL <- grey(0.7)
```

\newpage
### Urchins (Mean +/- SD).
Zeros shown as thin lines, missing data shown as blank

```{r,echo=F,fig.width=FIG.WIDTH,fig.height=FIG.HEIGHT,warning=F,message=F}
    
    limits <- aes(ymax = urchins.n.per.m2 + urchins.n.per.m2.sd , ymin=urchins.n.per.m2 )
      p1 <- ggplot(dat.mob.invert, aes(x=factor(Year),y=urchins.n.per.m2)) + 
        theme_acs()+
        ylab(expression("Sea Urchins per m"^2))+
        geom_bar(stat="identity",fill=COL,colour=1) +
        geom_errorbar(limits,width=0.2)+
        facet_wrap(~Site.numb,nrow=5,ncol=2)
      print(p1)
```

\newpage
### Sea Cucumbers (Mean +/- SD).
Zeros shown as thin lines, missing data shown as blank

```{r,echo=F,fig.width=FIG.WIDTH,fig.height=FIG.HEIGHT,warning=F,message=F}
    limits <- aes(ymax = sea.cucs.n.per.m2 + sea.cucs.n.per.m2.sd , ymin=sea.cucs.n.per.m2 )
      ggplot(dat.mob.invert, aes(x=factor(Year),y=sea.cucs.n.per.m2)) + 
        theme_acs()+
        ylab(expression("Sea Cucumbers per m"^2))+
        geom_bar(stat="identity",fill=COL,colour=1) +
        geom_errorbar(limits,width=0.2)+
        facet_wrap(~Site.numb,nrow=5,ncol=2)
```

\newpage
### Scallops (Mean +/- SD).
Zeros shown as thin lines, missing data shown as blank

```{r,echo=F,fig.width=FIG.WIDTH,fig.height=FIG.HEIGHT,warning=F,message=F}

    limits <- aes(ymax = scallops.n.per.m2 + scallops.n.per.m2.sd , ymin=scallops.n.per.m2 )
      ggplot(dat.mob.invert, aes(x=factor(Year),y=scallops.n.per.m2)) + 
        theme_acs()+
        ylab(expression("Scallops per m"^2))+
        geom_bar(stat="identity",fill=COL,colour=1) +
        geom_errorbar(limits,width=0.2)+
        facet_wrap(~Site.numb,nrow=5,ncol=2)
```

\newpage
### Crabs (Mean +/- SD).
Zeros shown as thin lines, missing data shown as blank

```{r,echo=F,fig.width=FIG.WIDTH,fig.height=FIG.HEIGHT,warning=F,message=F}

    limits <- aes(ymax = crabs.n.per.m2 + crabs.n.per.m2.sd , ymin=crabs.n.per.m2 )
      ggplot(dat.mob.invert, aes(x=factor(Year),y=crabs.n.per.m2)) + 
        theme_acs()+
        ylab(expression("Crab per m"^2))+
        geom_bar(stat="identity",fill=COL,colour=1) +
        geom_errorbar(limits,width=0.2)+
        facet_wrap(~Site.numb,nrow=5,ncol=2)
```

\newpage
### Limpets (Mean +/- SD).
Zeros shown as thin lines, missing data shown as blank

```{r,echo=F,fig.width=FIG.WIDTH,fig.height=FIG.HEIGHT,warning=F,message=F}

    limits <- aes(ymax = limpets.n.per.m2 + limpets.n.per.m2.sd , ymin=limpets.n.per.m2 )
      ggplot(dat.mob.invert, aes(x=factor(Year),y=limpets.n.per.m2)) + 
        theme_acs()+
        ylab(expression("Limpets per m"^2))+
        geom_bar(stat="identity",fill=COL,colour=1) +
        geom_errorbar(limits,width=0.2)+
        facet_wrap(~Site.numb,nrow=5,ncol=2)
```

\newpage
### Non-Limpet Gastropods (Mean +/- SD).
Zeros shown as thin lines, missing data shown as blank

```{r,echo=F,fig.width=FIG.WIDTH,fig.height=FIG.HEIGHT,warning=F,message=F}
    
    limits <- aes(ymax = gastropod.n.per.m2 + gastropod.n.per.m2.sd , ymin=gastropod.n.per.m2 )
      ggplot(dat.mob.invert, aes(x=factor(Year),y=gastropod.n.per.m2)) + 
        theme_acs()+
        ylab(expression("Gastropods per m"^2))+
        geom_bar(stat="identity",fill=COL,colour=1) +
        geom_errorbar(limits,width=0.2)+
        facet_wrap(~Site.numb,nrow=5,ncol=2)
```

