---
title: "Seastars"
author: "Genoa Sullaway"
date: "8/27/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---
## Purpose
- to plot seastar patterns in abundance and occurance, look at potential predator prey relationships and look at size distributions (not done yet).  

## Summary Notes
- 2015 data was wonky so removed it, all plots are 2016-2019 data with zone level replication.   
- A lot of YOY seastars in 2017
- Looks like there is more to explore with seastars and prey... Species and site specific relationships ?   

\n Take home:
- No large temporal changes in seastars in the 4 years post-wasting disease onset. Recruitment event in 2017 does not result in subsequent increase in seastars. 
- Increase in urchins with no seastars at Tatoosh (but since we dont have pre-WD data, can't say if this is a change. WIsh we knew what it looked like before)
- Strong top-down control of sestars on prey, this is weaker at destruction... mediated by rugose habitat here?
- Blood stars were supposedly minimally effected by WD and mostly prey on sponges, strong relationship between sponge density and henricia abundance. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(PNWColors)
library(viridis)
 
color2= pnw_palette("Cascades",n=2 ,"discrete")
color5= pnw_palette("Cascades",n=5 ,"discrete")
color6= pnw_palette("Cascades",n=6 ,"discrete")
seastar_spp <- c(#"blood_sea_star",
  "large_sea_star","med_sea_star",
  #                   "sea_star_YOY",
  "large_sea_star","leather_star","P_ochraceous")
site_factor <- c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island", "Neah Bay")
```

## Site-Level Abundance
- Plot patterns of seastar abundance by site and year using site-level data
```{r invert swath site level}
inverts <- read_csv(here('Flagstone paper','Data','invert_swath_sitelevel.csv'))
seastars_site <- inverts %>% 
  mutate(lump = case_when(lump %in% c("brood_sea_star", "blood_star")~ "blood_sea_star",
                          lump %in% c("med_seastar")~ "med_sea_star",
                          TRUE ~ lump )) %>%
  dplyr::filter(lump %in% seastar_spp,
                !year==2015,
         !site %in% c("Anderson Point", "Point of the Arches","Rock 305", "Chibadehl Rocks","Teahwhit Head")) %>%     
  dplyr::select(year,site,lump,dens_weighted,se_w) 

# look at community across sites and years using a stacked bar  
seastars_site %>%
  group_by(site,lump) %>%
  summarise(dens_weighted = mean(dens_weighted)) %>%
  ungroup() %>%
  mutate(site = factor(site, levels =c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island" , "Neah Bay" ))) %>%
ggplot(aes(fill=lump, y=dens_weighted, x=site)) + 
    geom_bar(position="stack", stat="identity") +
  scale_fill_manual(values = color6) +
  ggtitle("2016-2019")+
  theme_classic()

#look at stacked bar for each site across years 
seastars_site %>%  
  group_by(year,site,lump) %>%
  summarise(dens_weighted = mean(dens_weighted)) %>% #take mean across zones 
  ungroup() %>%
  mutate(site = factor(site, levels =c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island" , "Neah Bay" ))) %>%
ggplot(aes(fill=lump, y=dens_weighted, x=year)) + 
    geom_bar(position="stack", stat="identity") +
  scale_fill_manual(values = color6) +
  ggtitle("2016-2019") +
  facet_wrap(~site,scales = "free") +
  theme_classic()

################same as above but remove henricia

#look at stacked bar for each site across years 
seastars_site %>%  
 filter(!lump %in% c( "blood_sea_star","sea_star_YOY")) %>% 
  group_by(year,site,lump) %>%
  summarise(dens_weighted = mean(dens_weighted)) %>% #take mean across zones 
  ungroup() %>%
  mutate(site = factor(site, levels =c("Destruction Island","Cape Johnson","Cape Alava","Tatoosh Island" , "Neah Bay" ))) %>%
ggplot(aes(fill=lump, y=dens_weighted, x=year)) + 
    geom_bar(position="stack", stat="identity") +
  scale_fill_manual(values = color6) +
  ggtitle("2016-2019, Possible urchin predators - remove henricia and YOY") +
  facet_wrap(~site )+#,scales = "free") +
  theme_classic()
```


- Get seastar diets, match them with prey
- Look at transect,zone and site level relationships of species with their specific prey groups 
- Which species is most likely to each urchins? make a plot that focuses on that relatioship
- Right now considering bivalves and urchins as prey -- but need to research this a bit

```{r}
sd_agg <- function(sdvec,nvec){
  sqrt(sum((sdvec*nvec)^2,na.rm=T)/sum(nvec,na.rm=T))
}
```


```{r}
#load data - filter to species, remove some 2015 sites that werent surveyed in other years, fix some species names  
inverts <- read_csv(here('Flagstone paper','Data','invert_swath_transectlevel.csv')) %>% 
    filter(!year == 2015) 

seastars_site <- inverts %>% 
  mutate(lump = case_when(lump %in% c("brood_sea_star", "blood_star")~ "blood_sea_star",
                          lump %in% c("med_seastar")~ "med_sea_star",
                          TRUE ~ lump )) %>%
  dplyr::filter(lump %in% seastar_spp,
         !site %in% c("Anderson Point", "Point of the Arches","Rock 305", "Chibadehl Rocks","Teahwhit Head")) %>%   
  dplyr::select(year,site,zone,lump,transect,dens_weighted,sd_w) %>%
         mutate(pred_prey_group = "seastar") %>%
         group_by(year,site,pred_prey_group) %>%
         summarise(n_obs=n(),
                     n_obs=sum(n_obs),
                     dens_weighted=mean(dens_weighted),
                     sd_w=sd_agg(sd_w,n_obs)
                     ) %>% 
          mutate(se_w=sd_w/sqrt(n_obs)) 
 
#unique(inverts$lump)
seastar_prey_site <- inverts %>% 
  dplyr::filter(lump %in% c("bivalve","green_urchin","shelled_mollusk", "purple_urchin","chiton",
                            "shelled_gastropod","red_urchin"),
                !site %in% c("Anderson Point", "Point of the Arches","Rock 305", "Chibadehl Rocks","Teahwhit Head")) %>%    
  dplyr::select(year,site,zone,lump,transect,dens_weighted,sd_w) %>%
  mutate(pred_prey_group = "prey") %>%
         group_by(year,site,pred_prey_group) %>%
         summarise(n_obs=n(),
                     n_obs=sum(n_obs),
                     dens_weighted=mean(dens_weighted),
                     sd_w=sd_agg(sd_w,n_obs)) %>% 
          mutate(se_w=sd_w/sqrt(n_obs)) 

seastar.pred.prey <- full_join(seastars_site,seastar_prey_site) 

```

## Seastar Predator-Prey
```{r}
#plot predators x prey geom_point
seastar.pred.prey %>%
  select(-n_obs, -sd_w, -se_w) %>%
   spread(pred_prey_group, dens_weighted) %>%
   ungroup() %>%
   ggplot() +
   geom_point(aes(x=seastar, y = prey,color = as.factor(site))) +
   scale_color_viridis(discrete = T)  +
  xlab("Seastar Density (Includes all lumps except blood stars)\nZone Level Reps") +
    facet_wrap(~site ) +
  theme_classic()

seastar.pred.prey %>%
  select(-n_obs, -sd_w, -se_w) %>%
   spread(pred_prey_group, dens_weighted) %>%
   ungroup() %>%
   ggplot() +
   geom_point(aes(x=seastar, y = prey,color = as.factor(site))) +
   scale_color_viridis(discrete = T)  +
  xlab("Seastar Density (Includes all lumps except blood stars)\nZone Level Reps") +
  theme_classic()

```

## Seastar Predator-Prey Timeseries
```{r predator prey time series}
  seastar.pred.prey %>%
     ggplot(aes(x=year, y = dens_weighted, color = as.factor(pred_prey_group))) +
       geom_point() +
       geom_line() +
       # geom_errorbar(aes(ymin=dens_weighted-se_w, ymax=dens_weighted+se_w)) +
       # scale_x_continuous(expand = c(0, 0)) + 
       # scale_y_continuous(expand = c(0, 0)) +
       scale_color_manual(values = color2) +
       ylab("Mean Seastar Density (Includes all lumps except blood stars") +
       xlab("Year")+
       facet_wrap(~site ) +
       ggtitle("Site-Level Means") +
       theme_classic()+
       theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

seastar.pred.prey %>%
     ggplot(aes(x=year, y = dens_weighted, color = as.factor(pred_prey_group))) +
       geom_point() +
       geom_line() +
   # geom_errorbar(aes(ymin=dens_weighted-se_w, ymax=dens_weighted+se_w)) +
       scale_color_manual(values = color2) +
       xlab("Year") +
       ylab("Mean Seastar Density (Includes all lumps except blood stars") +
       facet_wrap(~site, scales= "free") +
       ggtitle("Site-Level Means, Scales Free") +
       theme_classic()+
       theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


```{r, eval = FALSE}
## Henricia Predator-Prey
Henricia has pretty different diets than many other stars, so I removed them from the above comparisions and did a seperate one below. THey mostly eat sponges and bacteria, so the only prey group I have here is sponge. 
#make seperate prey DF 
blood_star_prey <- inverts %>% 
  dplyr::filter(lump %in% c("sponge"),
                !site %in% c("Anderson Point", "Point of the Arches","Rock 305", "Chibadehl Rocks","Teahwhit Head")) %>%    
  dplyr::select(year,site,zone,lump,dens_weighted,se_w) %>%
  mutate(pred_prey_group = "prey")

blood.pred.prey <- full_join(seastars %>% filter(lump %in% c("blood_sea_star")), blood_star_prey)  

blood.pred.prey %>%
   group_by(year,site,zone,pred_prey_group) %>%
   summarise(dens_weighted=sum(dens_weighted)) %>%  
 #select(-se_w, -lump) %>%
   spread(pred_prey_group, dens_weighted) %>%
   ungroup() %>%
   ggplot() +
   geom_point(aes(x=seastar, y = prey,color = as.factor(site))) +
   scale_color_viridis(discrete = T) +
  ylab("Sponge Density") +
  xlab("Blood Star Density") +
  ggtitle("Henricia and sponges only ") +
    theme_classic()

# Times series
blood.pred.prey %>%
   group_by(year,site,zone,pred_prey_group) %>%
   summarise(dens_weighted=sum(dens_weighted)) %>%  
     ggplot(aes(x=year, y = dens_weighted, color = as.factor(pred_prey_group))) +
       geom_point() +
       geom_line() +
       scale_color_manual(values = color2) +
       ylab("Henricia Density (Includes all lumps except blood stars") +
       xlab("Year")+
       facet_grid(zone~site ) +
       ggtitle("Henricia Zone-Level Means") +
       theme_classic()+
       theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


blood.pred.prey %>%
   group_by(year,site,pred_prey_group) %>%
   summarise(dens_weighted=sum(dens_weighted)) %>%  
     ggplot(aes(x=year, y = dens_weighted, color = as.factor(pred_prey_group))) +
       geom_point() +
       geom_line() +
       scale_color_manual(values = color2) +
       xlab("Year") +
       ylab("Henricia Density") +
       facet_wrap(~site, scales= "free") +
       ggtitle("Henricia Site-Level Means, Scales Free") +
       theme_classic()+
       theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

 
```





```{r eval =FALSE }
 ## Size data - parce and plot 
#- Needs more work 
# - Plot Seastar size distribution by site, lumped all species because of availability of size data
# - Size data doesnt show that many differences... there may be a better way to parce it? 

#source("swatch_invert_overview_ORL.Rmd") 
#saveRDS(dat_all, here("Flagstone paper","Data","invert_lumps_with_size.RDS"))


#size data is not in current data sets so I coped some of Owens code from: #source("swatch_invert_overview_ORL.Rmd")  but adapted it to keep size.

lump <- read_csv(here('Data','CSV_2015_on','invert_groups.csv'))  %>%
  rename(classcode=CLASSCODE, tot_count=total_count, lump=group)

species_names <- read_csv(here('Data','CSV_2015_on','species_code_list_swath.csv'))

dat.2015 <- read.csv(here('Data','CSV_2015_on',"2015_OCNMSDataComplete_standardized_122116.csv"))
# species_names <- read_csv(here('Data','CSV_2015_on','species_code_list_swath.csv'))

# Make list of transects by site, and species group.
dat.2015 <- left_join( dat.2015, species_names %>% dplyr::select(species,group) %>% rename(PISCO.Classcode=species))

# subset the variables we need, filter for just the swath invert species, then expand the data, adding zeroes
dat.2015.clean <- dat.2015 %>% 
  filter(PISCO.datatype=="swath", data.type=="swath",group=="Invert") %>% 
  # select the needed columns
  dplyr::select(Site,Observer,Transect,Transect.area..inverts.,group,PISCO.Classcode,Count,Size.cm) %>% 
  # rename some variables
  rename(size=Size.cm, site=Site,observer=Observer,transect=Transect,observed_area=Transect.area..inverts.,classcode=PISCO.Classcode,count=Count) %>% 
  # fix some site names
  mutate(site=as.character(site),
         site=case_when(
                site=="Destruction Island SW"~"Destruction Island",
                site=="Anderson Pt."~"Anderson Point",
                site=="Chibahdel" ~ "Chibadehl Rocks",
                site=="Pt. of the Arches" ~ "Point of the Arches",
                site=="Teawhit Head"~"Teahwhit Head",
                TRUE ~ site)) %>% 
  mutate(observed_area=as.numeric(observed_area))

dat.2015.clean.complete <- dat.2015.clean %>% 
  # fill in zeroes for unobserved species (using all the unique species in the overall data as our universe of possible species)
  complete(nesting(site,transect),classcode,fill=list(count=0,observed_area=60)) %>% 
  mutate(year=2015,zone=5) %>%
  left_join(lump,by='classcode') %>%
  filter(!site %in% c("Anderson Point", "Point of the Arches","Rock 305", "Chibadehl Rocks","Teahwhit Head"),
         !is.na(size),
         lump %in% seastar_spp) %>%
  select(site, transect, size,year,zone,lump,classcode)
 
################################################################################################################
### now 2016-2019 data
dat <- read_csv(here('Data','2019','NWFSC_SWATH_ALLYEARS_data_2019.csv'),col_types = 'cdddccccddddddccddddccc') %>%  select(YEAR,SITE,TRANSECT,ZONE,`SWATH WIDTH`,`DEPTH (ft)`,`METERS sampled`,SEGMENT,CLASSCODE,COUNT,SIZE,notes) %>% 
  # rename some variables
  rename(year=YEAR, zone=ZONE,site=SITE,transect=TRANSECT,width=`SWATH WIDTH`,swath_length=`METERS sampled`,depth=`DEPTH (ft)`,segment=SEGMENT,classcode=CLASSCODE,count=COUNT,size=SIZE) %>%
  # fix one weird data entry issue
  mutate(segment=ifelse(segment=="10-20M","10-20m",segment)) %>%
  left_join(lump,by='classcode') %>%
  filter(lump %in% seastar_spp, !is.na(size)) %>%
  select(site, transect, size,year,zone,lump,classcode)

  seastar.size <- rbind(dat.2015.clean.complete,dat)

#violin plot
seastar.size %>%
  ggplot() +
  geom_violin(aes(site,size))  +
  theme_classic()
```

