---
title: "Summarise Relief Data OCNMS"
author: Genoa Sullaway
date: "July 2020"
output: 
  html_document:
      toc: true
      toc_float: true
---


\n This uses 2016-2019 data, 2015 Relief Data uses quadrats and is inconsistent with later methods, calculates mean and standard deviation (specific to beta distributed data) for a few different flavors of data summaries, summarized below. 

\n Data frames created here are saved in: OCNMS/Data/Summarized data

\n There are four **Percent Cover** data sets:

\n 1. dat.transect.summary = RELIEF_Transect_Level_Summary.csv: This has percent cover across TRANSECT/ZONE/SIDE/SITE/YEAR. Transect is lowest level of replication here. There are only 11 transects in whole dataset that have less than 30 upc counts (aka an incomplete transect)... In the cases where it is less than 30 I get total % cover from the total points sampled for that transect.

\n 2. dat.zone.summary = RELIEF_Zone_Level_Summary.csv: This has mean percent cover and variance across ZONE/SITE/YEAR.

\n 3. dat.site.summary = RELIEF_Site_Level_Summary.csv: This has mean percent cover and variance across SITES.


\n There are four **Simspons Diversity** data sets:

\n 1. dat.diversity = RELIEF_DIVERSITY_Transect_Level_Summary.csv: This has Simpsons Diversity across TRANSECT/ZONE/SIDE/SITE/YEAR. Transect is lowest level of replication here.

\n 2. dat.diversity.zone.summary = RELIEF_DIVERSITY_Zone_Level_Summary.csv: This has mean Simpsons Diversity and variance across ZONE/SITE/YEAR.

\n 3. dat.diversity.year.site.summary= RELIEF_DIVERSITY_Year_Site_Level_Summary.csv: This has mean Simpsons Diversity and variance across SITE/YEAR.

\n 4. dat.diversity.site.summary = RELIEF_DIVERSITY_Site_Level_Summary.csv: This has Simpsons Diversity and variance across SITES.


## Summary Thoughts
-  Tattoosh and Destruction appear to have greatest diversity in relief which makes sense with diver impressions. Cape Johnson has the lowest diversity in relief-- so if rockfish respond to complexity  non-biogenic habitat we may expect to see greater abundances or divesity in rockfish species at Tattoosh and Destruction. 

```{r setup,message=FALSE, warning=FALSE}
#to load data make sure directory is set to knit from project directory
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
 
library(here)
library(tidyverse)
#library(knitr)
library(viridis)
library(vegan)
 
substrate_codes <- read_csv(here("Data","CSV_2015_on","substrate_codes.csv"))
 
#2016 --> 2019
dat.relief <- read_csv(here("Data","CSV_2015_on","NWFSC_UPC_ALLYEARS_data_2019.csv"), col_types = cols(SIDE = col_character())) %>% 
              filter(CATEGORY == "RELIEF") %>%
              select(1:20) %>%
              rename(DEPTH.FT= "DEPTH (FT)")  %>%
              #mutate(QUADRAT = NA, PERCENT.COVER = NA) %>% #will circle back tothis and get % cover for 2016 to 2019 data 
              group_by(SITE, YEAR,SIDE,ZONE,TRANSECT, CLASSCODE) %>% #combine across segments 
              summarise(COUNT = sum(COUNT)) %>%
              spread(CLASSCODE, COUNT) %>%
              ungroup() %>%
              replace(is.na(.), 0) %>%
              dplyr::mutate(Sum = rowSums(.[,6:9])) %>%
              tidyr::gather(6:9, key = "CLASSCODE", value = "COUNT") %>%
              mutate(percent_cover = COUNT/Sum, CLASSCODE = factor(CLASSCODE, levels=c("0-10cm","10cm-1m","1m-2m",">2m")))

dat.zone.summary <- dat.relief %>%
  group_by(YEAR,SITE,ZONE,CLASSCODE) %>% 
  summarise(
      N=length(percent_cover),
      MEAN=mean(percent_cover),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N)) %>%
  ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay")))
 
dat.site.summary <- dat.relief %>%
  group_by(SITE,CLASSCODE) %>% 
  summarise(
      N=length(percent_cover),
      MEAN=mean(percent_cover),
      SD=sqrt(MEAN * (1 - MEAN) * N),
      SE=SD/sqrt(N)) %>%
  ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay")))

  write.csv(dat.relief, here("Data","Summarized data","RELIEF_Transect_Level_Summary.csv"))
  write.csv(dat.zone.summary, here("Data","Summarized data","RELIEF_Zone_Level_Summary.csv"))
  write.csv(dat.site.summary, here("Data","Summarized data","RELIEF_Site_Level_Summary.csv"))
  

```
## Percent Cover
### Create data frames
```{r}
ggplot(dat.zone.summary,aes(x=YEAR)) +
  geom_bar(aes(y=MEAN, fill= CLASSCODE), position="fill", stat="identity",width = 0.95
           )+
  scale_fill_viridis(discrete = T) + #,  guide = guide_legend(reverse = TRUE), name = "Stock") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "left",aspect.ratio = 3/1)  +
  theme_classic()+
  ylab("Mean Percent Cover") +
  xlab("Year") +
  facet_grid(ZONE~SITE) +
 # coord_flip() +
  scale_x_continuous(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0)) +
  theme(legend.position="left") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), strip.text.x = element_text(size = 7))

```

### Plot
```{r,echo=FALSE}
ggplot(dat.site.summary,aes(x=SITE)) +
  geom_bar(aes(y=MEAN, fill= CLASSCODE), position="fill", stat="identity",width = 0.95
           )+
  scale_fill_viridis(discrete = T) + #,  guide = guide_legend(reverse = TRUE), name = "Stock") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "left",aspect.ratio = 3/1)  +
  theme_classic()+
  ylab("Mean Percent Cover") +
  xlab("Year") +
  #facet_wrap(~ZONE, nrow =1) +
  scale_x_discrete(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0)) +
  theme(legend.position="left") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), strip.text.x = element_text(size = 7))

```


## Simpsons Diversity Index 
### Create data frames
```{r calculate Simpsons Diversity}

#use dat.relief to get Simpsons index for every transect and save data frames 
dat.diversity<- dat.relief %>% 
    group_by(YEAR,SITE,ZONE,SIDE,TRANSECT,CLASSCODE) %>%
    dplyr::summarise(
            PROPORTION=sum(percent_cover),
            PROPORTION_2=sum(PROPORTION)^2,
             .groups = "keep") %>%
      group_by(YEAR,SITE,ZONE,SIDE,TRANSECT) %>%
    dplyr::summarise(    
            DIVERSITY= sum(PROPORTION_2),
            RELIEF_SIMPSONS_DIVERSITY= 1 - DIVERSITY, # simpsons diversity
            .groups = "keep"
    ) %>%
    select(-DIVERSITY)

# calculate substrate diversity for each site-year-zone
  dat.diversity.zone.summary <- dat.diversity %>% 
    group_by(YEAR,SITE,ZONE) %>%
    dplyr::summarise(
      MEAN=mean(RELIEF_SIMPSONS_DIVERSITY),
      SD= sd(RELIEF_SIMPSONS_DIVERSITY),
      N=length(RELIEF_SIMPSONS_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )

# calculate substrate diversity for each site
  dat.diversity.year.site.summary <- dat.diversity %>% 
    group_by(YEAR,SITE) %>%
    dplyr::summarise(
      MEAN=mean(RELIEF_SIMPSONS_DIVERSITY),
      SD= sd(RELIEF_SIMPSONS_DIVERSITY),
      N=length(RELIEF_SIMPSONS_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  
  # calculate substrate diversity for each site
  dat.diversity.site.summary <- dat.diversity %>% 
    group_by(SITE) %>%
    dplyr::summarise(
      MEAN=mean(RELIEF_SIMPSONS_DIVERSITY),
      SD= sd(RELIEF_SIMPSONS_DIVERSITY),
      N=length(RELIEF_SIMPSONS_DIVERSITY),
      SE=SD/sqrt(N),
      .groups = "keep"
      )
  
 write.csv(dat.diversity, here("Data","Summarized data","RELIEF_DIVERSITY_Transect_Level_Summary.csv"))
  write.csv(dat.diversity.zone.summary, here("Data","Summarized data","RELIEF_DIVERSITY_Zone_Level_Summary.csv"))
  write.csv(dat.diversity.site.summary, here("Data","Summarized data","RELIEF_DIVERSITY_Site_Level_Summary.csv"))
  write.csv(dat.diversity.year.site.summary, here("Data","Summarized data","RELIEF_DIVERSITY_Year_Site_Level_Summary.csv"))

```

### Plot 
```{r,echo=FALSE}
dat.diversity.zone.summary %>%
  ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay"))) %>%
  ggplot(aes(x=YEAR, y =MEAN, group = as.factor(ZONE), color = as.factor(ZONE))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=MEAN -SE, ymax=MEAN+SE), width = 0.1)+
  facet_wrap(~SITE)+
  theme_classic() +
  ylab("Simpsons Diversity Index")


```

```{r,echo=FALSE}
dat.diversity.year.site.summary %>%
    ungroup() %>%
   mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay"))) %>%
  ggplot(aes(x=YEAR, y =MEAN))+
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=MEAN -SE, ymax=MEAN+SE), width = 0.1)+
  facet_wrap(~SITE) + 
  theme_classic() +
  ylab("Simpsons Diversity Index - RELIEF")
```

```{r,echo=FALSE}
dat.diversity.site.summary %>%
    ungroup() %>%
  mutate(SITE = factor(SITE, levels = c("Destruction Island","Cape Johnson", "Cape Alava","Tatoosh Island","Neah Bay"))) %>%
  ggplot(aes(x=SITE, y =MEAN))+
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=MEAN -SE, ymax=MEAN+SE), width = 0.1)+
  theme_classic() +
  ylab("Simpsons Diversity Index - RELIEF")
```






 