---
title: "Annual update - UPC"
author: Nick Toimieri
date: sys.date()
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
fontsize: 11pt
---
# NOTES

To update fish data to the most recent year, you will need: 

- update_swath_fish.R

in the 'update-data' chunk. The update is done automatically within this RMD, but you will need to check that the trigger is turned on in the if statement in the 
in the 'update-data' is "if(1==1)". 

You will also need to have:

- swath-process-pre-2017-data.R
- R-functions-ocnms.R  (lots of functions)

as helper files. These are in the repo.

# DATA YEAR 

Set the data_year year

```{r SETUP, include=FALSE}
rm(list = ls())
library(here)
repo_dir = here::here()
home_dir = paste0(repo_dir,"/_00_Annual-Update/")
setwd(home_dir) # set home dir to annual update folder
data_year = 2025
source("settings.r")
source("R-functions-ocnms.r")

################################################################################

knitr::opts_chunk$set(
  root.dir = home_dir,
  fig.path = fig_dir,
  cache.path = 'cache/graphics-', 
  echo=FALSE,
  error=FALSE,
  include = TRUE,
  dev='png',
  dpi=300,
  warning=FALSE,
  #out.width = '100%',
  fig.align='center'
)

```

# Bring in data

```{r upc-import}

upc = data.frame(read.csv( paste0(data_dir,'NWFSC_UPC_ALLYEARS_data_2025.csv'), header=T))

upc = upc %>% 
  mutate(depth_m = `DEPTH..FT.` / 3.281) %>%
  dplyr::select(-c(entry.order, Cntrl.a, notes, problems, `DEPTH..FT.`, `DEPTH..M.`) )
  
upc_transect = upc %>% group_by(YEAR, SITE, SIDE, ZONE, TRANSECT, CATEGORY, CLASSCODE) %>%
  summarize(count = sum(COUNT))

upc_cover = upc_transect %>% filter(CATEGORY == 'COVER')
upc_relief = upc_transect %>% filter(CATEGORY == 'RELIEF')
upc_subs = upc_transect %>% filter(CATEGORY == 'SUBSTRATE')

write.csv(upc_transect, paste0(data_dir,"Summarized_Data_UPC_by_Transect.csv"), row.names = F)


```


```{r}
# check counts by transect
x_relief = upc_relief %>%
 group_by(YEAR, SITE, SIDE, ZONE, TRANSECT) %>% 
  summarize(SUM = sum(count))

# quick plot

upc_relief$CLASSCODE[upc_relief$CLASSCODE==">2M"] = ">2m"
upc_relief$CLASSCODE[upc_relief$CLASSCODE=="<2M"] = "1m-2m"
upc_relief$CLASSCODE[upc_relief$CLASSCODE=="0-10m"] = "0-10cm"

upc_relief$CLASSCODE = factor(upc_relief$CLASSCODE, 
                                 levels = c("0-10cm", "10cm-1m","1m-2m", ">2m"))
levels(upc_relief$CLASSCODE)

relief = upc_relief %>%
  group_by(YEAR, SITE, CLASSCODE) %>%
  summarize(avg_count = mean(count))
relief$prop = relief$avg_count/30
relief$perc = 100*relief$avg_count/30

```





