---
title: "YOY analysis: Black and yellowtail rockfish complex"
author: "Nick Tolimieri & Ole Shelton"
date: " `r Sys.Date()` "
format: 
  docx: default
bibliography: references.bib
---

```{r setup}
#| echo: false
#| messange: false
#| warning: false

knitr::opts_chunk$set(echo = FALSE)

survey_year = 2024
SPECIES = 'SEME'
SPECIES = 'SEFL'
SPECIES = c('SEFL', 'SEME')


if(length(SPECIES)==1){
  if(SPECIES=='SEME'){
    SPP = 'Black rockfish' 
    spp='black rockfish'
    sciname = "Sebastes melanostictus"}

  if(SPECIES=='SEFL'){
    SPP = 'Yellowtail rockfish' 
    spp='yellowtail rockfish'
   sciname = "Sebastes flavidus"}
}  

if(length(SPECIES)==2){
  if(SPECIES[1] == 'SEFL' & SPECIES[2]=='SEME'){
  SPPx = c('Black rockfish','Yellowtail rockfish')
  SPPz = c("SEBYT", "SEME","SEFL")
  SPP =  'Black and yellowtail rockfishes' 
  spp =  'black and yellowtail rockfishes'
  sciname = "Sebastes melanostictus and S. flavidus"}
}

```

```{r}
#| echo: false
#| messange: false
#| warning: false
#
  library(tidyverse)
  library(xtable)
  library(terra)
  library(ggplot2)
  library(maps)
  library(mapdata)
  library(mapproj)

# set directories ####
home_dir <- '~/Github/OCNMS/'
work_dir <- '~/Github/OCNMS/_01_Assessment-indicies-BRF-YTRF/'
data_dir = paste0(home_dir,'Data/',survey_year,'/')

# This is a padded data.frame which includes all transects
# includes zero observations
# includes low visibility transects which will be dropped.

# All processing steps can be found in the Git repo: 
# "/GitHub/OCNMS/_Annual-UPdate/update-swath-fish.R") for all of the species. 
# We have filtered out the just the black rockfish data for this analysis. 

dat.fish <- readRDS( paste0(data_dir,"Fish_2015-", survey_year,".rds") )

# Filter to include only black rockfish or "small" size category (rockfish YOY)
dat.yoy <- dat.fish %>% filter(species %in% SPECIES) %>%
              bind_rows(.,
                dat.fish %>% filter(size_class %in% c("small")))
# canary and black/yellowtail complex -- look similar
dat.can <- dat.fish %>% filter(species %in% c("SEPI","SEPYT"))
          

# Drop transects where the visibility is less than the visibility threshold (2m):
VIS_LIM <- 2
dat.yoy <- dat.yoy %>% filter(vis_m >= VIS_LIM)
dat.can <- dat.can %>% filter(vis_m >= VIS_LIM)


```

# Data description

These are derived data products relevant to estimating recruitment of the black rockfish and yellowtail rockfish comlex stemming from NWFSC dive surveys in the Olympic Coast National Marine Sanctuary (OCNMS) conducted between 2015 and `r survey_year`. We estimate recruitment (young-of-year abundance) for the complex because it is difficult to distinguish small recruits for these species. Description of survey methods and aims are detailed in [@tolimieri2023changes] We provide a brief summary of methods relevant to fish surveys here.

Divers on SCUBA conducted in situ surveys to count fish at each site along benthic belt transects (30 m by 2 m) following slightly modified procedures described in @malone2022large. Transects were conducted within or directly adjacent to canopy kelp beds (consisting of giant *Macrocystis pyrifera* or bull *Nereocystis luetkeana* kelps). From 2016 on, we surveyed at five (5) sites ([Fig. @fig-site-map]), sampling at two (2) locations within each site separated by \>100 m, and 2 depths within each location (5 and 10 m) Our goal was to complete six (6) replicate transects at each year-site-depth combination. In 2015 surveyed at 10 sites and conducted four (4) transects per site at 5 m depth ([Fig. @fig-site-map]).

During each fish transect, we counted and estimated the size (total length to nearest cm) of all fishes \>5 cm total length; the exception was rockfishes *Sebastes* spp., for which we estimated sizes of all individuals. Rockfishes $\leq$ 10 cm were considered young of year. Divers also estimated horizontal visibility on each transect by determining the distance at which the lead diver could distinguish their buddy's extended fingers. Transects with visibility less than 2 m were excluded from analyses.

As noted above, it is difficult to visually distinguish many rockfish species when they are small. Therefore, on our surveys, we categorized juvenile rockfishes into five (5) groups established in the literature [@johansson2018seasonal; @markel2020contrasting]:

\(1\) Yellowtail and black (YTB) included both yellowtail (*S. flavidus*) and black (*S. melanops*) rockfishes.

\(2\) The copper/quillback/brown (CQB) group included copper (*S. caurinus*), quillback (*S. maliger*), and brown (*S. auriculatus*) rockfishes.

\(3\) Canary (*S. pinniger*) and

\(4\) Blue rockfish (*S. mystinus*) to species.

\(5\) Unidentified individuals were categorized as juvenile rockfishes.

The estimated recruitment trend for (1), the black rockfish and yellowtail rockfish complex, is presented here.

```{r map-prep}
#| echo: false
#| messange: false
#| warning: false

  #### READ IN WASHINGTON MARINE SHORELINES (FROM http://www.ecy.wa.gov/services/gis/data/data.htm
  
  #base.dir <- getwd()
#   gis.dir<- "/Users/ole.shelton/GitHub/OCNMS/Data/PG Shore shore_thematic_UTM10/"
#   setwd(gis.dir)
#   # shore.poly	<-	readShapePoly("shore_thematic_UTM10.shp")
#   
#   shore.poly	<-	readOGR(dsn=".","shore_thematic_UTM10")
#   shore		<-	data.frame(shore.poly)
# 
#   shore.dat	<-	 as(shore.poly,"data.frame")
#   shore.dat$NEW	<- 1:nrow(shore.dat)
#   shore.poly@data$id 	= rownames(shore.poly@data)
#   shore.points		<-	fortify(shore.poly,region="id")
#   shore.df			  <- merge(shore.points,shore.poly@data,by="id")
#   write.csv(shore.df,file="shore df.csv",row.names=F)
  shore.df        <- read.csv("../Data/PG Shore shore_thematic_UTM10/shore df.csv")
  
  # setwd(base.dir)
  PROJ <- "+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
  
  temp <- vect(cbind(shore.df$long,shore.df$lat),crs=PROJ)
  temp2 <- project(temp,"+proj=longlat +datum=WGS84")
  shore.df[,c("lon2","lat2")] <- geom(temp2)[,c("x","y")]
  
  loc <-  read.csv( paste0(work_dir,"Survey locations.csv"),fileEncoding="latin1")
  # update years
  loc$type[loc$type == '2015-22'] <- paste0('2015-',survey_year)
  loc <- loc %>% filter(type %in% c("2015 only",paste0('2015-',survey_year)))
  loc$nudge_x = 0.09
  loc$nudge_x[loc$Site =="Pt. of the Arches"] = 0.12
  loc$nudge_x[loc$Site =="Tatoosh Is."] = -0.08
  loc$nudge_x[loc$Site =="Chibahdehl"] = 0.08
  loc$nudge_x[loc$Site =="Neah Bay"] = 0.08
  loc$nudge_y = 0
  #loc$nudge_y[loc$Site =="Neah Bay"] = 0
  loc$nudge_y[loc$Site =="Chibahdehl"] = 0.01
  #loc$nudge_y[loc$Site =="Tatoosh Is."] = 0
  # loc.new				<- cbind(loc$lon, loc$latitude)
  # loc.new.sp		<- SpatialPoints(loc.new,CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
  # 
  # loc.new.sp.NAD83	<- spTransform(loc.new.sp, CRS(PROJ))
  # new.dat			      <- as(loc.new.sp.NAD83,"data.frame")
  # colnames(new.dat)	<- c("LonUTM","LatUTM")
  # loc <- data.frame(loc,new.dat)
  
  
  x.lim =c(-124.85,-124.3)
  y.lim =c(47.65,48.4)
  
  #wa <- map_data('worldHires','USA')
  #quartz(width=3,height=5)
  #wa <- map_data("state",region = c('washington'))
  site_map	<-	ggplot()+
				theme_bw() +
				theme(legend.key =  element_rect(fill = "white", colour = "white"),
					panel.border = element_rect(fill = NA, colour = "black",size=1.5)) +
		  #  geom_polygon(data=wa,aes(long,lat,group=group),fill=8,color="black")	+
				#coord_cartesian(xlim = x.lim,ylim=y.lim)     +
        geom_polygon(data=shore.df[shore.df$POLYTYPE == "Land" | shore.df$POLYTYPE == "Island" ,],
                     aes(lon2,lat2,group=group),fill=grey(0.7),color=NA)+
        geom_point(data= loc %>% filter(type %in% c(paste0('2015-',survey_year),"2015 only")),aes(x=lon,y=lat,color=type)) +
        scale_color_discrete("Survey")+
            geom_text(data= loc %>% filter(type %in% c(paste0('2015-',survey_year),"2015 only")),
                  aes(label=Site,x=lon,y=lat),size=2.5,
                  #check_overlap = TRUE,
                  nudge_x=loc$nudge_x,
                  nudge_y=loc$nudge_y
                  ) +
        xlab("Longitude") +
			  ylab("Latitude") +
        coord_fixed(ratio=1.3,xlim =x.lim,ylim=y.lim)
	
	#dev.off()
  		# 	bGrid +bAxis + bBack + bTics +

```


```{r fig-site-map}
#| echo: false
#| messange: false
#| warning: false
#| fig-cap: "Dive survey locations along the coast of Washington state."

site_map
```

As a result of our 2-m visibility requirements, some transects were omitted from the analyses. The following tables ([@tbl-2m-samples-YS]) show how the number fish transects with visibility \>2 m were distributed across depth, site, and years ([@tbl-2m-samples-YSD]), as well as across sites and year ([@tbl-2m-samples-YS]). 2015 includes only surveys conducted at 5-m depth; other years have data approximately evenly split between 5-m and 10-m depths.

```{r table-samples-prep}
#| echo: false
#| messange: false
#| warning: false

library(xtable)

# here are the sites ordered from S to N
nom <- c("Destruction Island","Teahwhit Head","Cape Johnson","Rock 305","Cape Alava","Point of the Arches","Anderson Point","Tatoosh Island","Chibadehl Rocks","Neah Bay")
N.trans <- dat.yoy %>% distinct(year,site,zone,transect,observer) %>% group_by(year,site,zone) %>%
              dplyr::summarise(N=length(transect))
N.trans.site.year <- N.trans %>% group_by (year,site) %>% summarise(N=sum(N))

N.trans$site <- factor(N.trans$site,levels=nom)
N.trans.site.year$site <- factor(N.trans.site.year$site,levels=nom)

N.trans <- N.trans %>% arrange(site)
#A <- left_join(A,nom.merge,by=c("site"="nom"))
N.trans.1 <- pivot_wider(N.trans, id_cols = c("year","zone"),
                                names_from=c("site"),
                                values_from=c("N")) 
N.trans.site.year <- N.trans.site.year %>% arrange(site,year)
N.trans.2 <- pivot_wider(N.trans.site.year, id_cols = c("year"),
                                names_from=c("site"),
                                values_from=c("N")) 
N.trans.site.year <- N.trans.site.year %>% arrange(year)
N.trans.3 <- pivot_wider(N.trans.site.year, id_cols =c("site") ,
                                names_from= c("year"),
                                values_from=c("N")) 

N.trans.2[is.na(N.trans.2)] <- 0
N.trans.3[is.na(N.trans.3)] <- 0
N.trans.2 <- N.trans.2 %>% arrange(year)

N.trans.3$site <- factor(N.trans.3$site,levels=nom)
N.trans.3 <- N.trans.3 %>% arrange(site)

N.trans.1[is.na(N.trans.1)] <- 0
N.trans.1 <- N.trans.1 %>%ungroup() %>% dplyr::select(-year,-zone) %>% mutate(Total= rowSums(.)) %>%
              pull(Total) %>%
              bind_cols(N.trans.1,.) %>%
              arrange(year,zone)
colnames(N.trans.1)[ncol(N.trans.1)] <- "Total"

N.trans.3 <- N.trans.3 %>% dplyr::select(-site) %>% colSums(.) %>% data.frame() %>% 
                t() %>% bind_cols(site="TOTAL",.) %>% bind_rows(N.trans.3,.)
```

```{r tbl-2m-samples-YSD}
#| echo: false
#| messange: false
#| warning: false
#| tbl-cap: "Number of transects conducted by year, site and depth zone.  Only transects that had at least 2m visibility are included"

library(knitr)
N.trans.1 = N.trans.1 %>% rename(Year = year, Zone = zone)

kable(N.trans.1)
```

\newpage

```{r tbl-2m-samples-YS}
#| echo: false
#| messange: false
#| warning: false
#| tbl-cap: "Number of transects conducted by year and site.  Only transects that had at least 2-m visibility are included"


N.trans.3 = N.trans.3 %>% rename(Site = site)

kable(N.trans.3)
```

## Recruitment: Young-of-year abundance trends

To calculate the average density of black/yellowtail rockfish complex in each year, we first calculate the mean density and standard error per site in each year. This approach means we are treating each transect as a i.i.d. sample of YOY density within each site and thus we ignore differences in abundance by depth zone.

From these site-year level means, we calculated a year-specific mean density by simulation. Specifically, for each year we independently drew a mean density for each site using a t-distribution with $\mu$ (the estimated site mean), $\sigma$ (the estimated site-specific standard error) and degrees of freedom, $\tau$. So for the $i^{th}$ realization, for site $s$ in year $y$ we have a predicted density, $X_{isy}$

```{=tex}
\begin{align}  & X_{isy} \sim T(\mu_{sy},\sigma_{sy},\tau_{sy}) \\ \end{align}
```
and then the predicted density for a single realization in a given year is the mean among sites observed. We repeat the simulation 100,000 times to provide an estimated mean density and uncertainty for a given year ([Fig. @fig-yoy-ts]).

Nearly all small rockfish fall into the 4 to 7 cm length range and all are considered to have recruited from the plankton during the calender year of the survey. Therefore, we view the density of \<10cm rockfish to be an indicator of recruitment for the black/yellowtail rockfish complex ([Fig. @fig-yoy-ts]).

```{r yoy-ts-prep}
#| echo: false
#| messange: false
#| warning: false

# GENERATE INDEXES BY SITE and AMONG SITES
#dat.yoy.summary <- dat.yoy %>% filter(species =="SEBYT")
dat.yoy.summary <- dat.yoy %>% filter(species %in% SPPz,
                                      size_class=='small')

#dat.yoy.summary %>% distinct(species,year,size_class)
  
# summarize using means and standard errors
dat.yoy.summary <- dat.yoy.summary %>%
                    # group_by(year,site,species) %>% 
                    group_by(year,site) %>% 
                    summarise(Mean=mean(Count),
                              VAR = var(Count),
                              SD=sd(Count),
                              N=length(Count),
                              SE=SD/sqrt(N))

# This is a simple calculation treating the means from each site as observed and iid samples of a grand Mean
dat.yoy.out.A <- dat.yoy.summary %>%
                  group_by(year)  %>%
                  summarise(grand.mean = mean(Mean),
                            N= length(Mean),
                            SE = sd(Mean)/sqrt(N))

      # We have a bunch of 0 observations, which preclude using weighting by the SE.
      # I looked extensively at two stage sampling expressions 
      # for design based estimators. 
      # They were complicated and didn't necessarily work for our design.  
      # Plus, they have troubles with the zero observations that we have.  
      # So I decided to simulate to determine the uncertainty.

#dim(dat.yoy.summary)
n.sim <- 100000

sim <- matrix(# simulate from a centered t-distribution.
        rt(nrow(dat.yoy.summary) * n.sim,dat.yoy.summary$N - 1) * dat.yoy.summary$SE +
                dat.yoy.summary$Mean, # Adjust to a different location and scale.
                nrow(dat.yoy.summary), n.sim) # Put it in a matrix.
sim[sim <0] <- 0
sim <- as.data.frame(sim)
colnames(sim) <- paste0("X",1:ncol(sim))

sim <- bind_cols(dat.yoy.summary %>% dplyr::select(year,site), sim) %>%
          pivot_longer(.,cols=3:ncol(.),values_to="val",names_to="real")

yoy.trend <- sim %>% group_by(year,real) %>%
          dplyr::summarise(mean.real = mean(val, na.rm=T)) %>% ungroup() %>%
          group_by(year) %>%
          dplyr::summarise(grand.mean = mean(mean.real, na.rm=T),
                           SE = sd(mean.real, na.rm=T),
                           q025 = quantile(mean.real,probs=0.025, na.rm=T),
                           q05 = quantile(mean.real,probs=0.05, na.rm=T),
                           q25 = quantile(mean.real,probs=0.25, na.rm=T),
                           q75 = quantile(mean.real,probs=0.75, na.rm=T),
                           q95 = quantile(mean.real,probs=0.95, na.rm=T),
                           q975 = quantile(mean.real,probs=0.975, na.rm=T))

# Make yoy rockfish time-series plots
yoy.ts <- ggplot() + 
  geom_point(data=dat.yoy.summary, aes(x=year,y=Mean),alpha=0.2) +
  geom_errorbar(data=dat.yoy.summary,
                aes(x=year,ymin=Mean-SE,ymax=Mean+SE),
                width=0,
                alpha=0.2) +
  geom_point(data=yoy.trend, aes(x=year,y=grand.mean),
             color="blue",size=3)  +
  geom_errorbar(data=yoy.trend,
                aes(x=year,ymin=q025,ymax=q975),
                width=0,
                color="blue",size=0.5) +
  geom_errorbar(data=yoy.trend,
                aes(x=year,ymin=q25,ymax=q75),
                width=0,
                color="blue",size=1.2) +
  scale_y_sqrt(expression("Young-of-Year Density (fish 120 m"^"-3"*")"),
                  #trans="sqrt",
                  limits=c(0,NA),
                  expand=c(0,NA),
                  breaks=c(0,1,5,10,25,50,100),
                  labels=c("0","1","5","10","25","50","100")) +
  xlab("Year") +
  #geom_hline(yintercept=0,color="red")+
  theme_bw()
  

# save out ####
dat.yoy.sites = dat.yoy.summary
write.csv(yoy.trend, paste(work_dir, "Estimated-YOY-trend-coast.csv"), row.names = F)
write.csv(dat.yoy.sites, paste(work_dir, "Estimated-YOY-trend-sites.csv"), row.names = F)


```

```{r fig-yoy-ts}
#| echo: false
#| messange: false
#| warning: false
#| fig-cap: "Time-series of estimated young-of-year rockfish (black-yellowtail complex) density on the Washington coast. Black points show means and standard errors for individual sites. Blue points show coastwide density estimates, interquartile range and 95\\% intervals for each year. Note y-axis is square root."

yoy.ts
```

\newpage

## References

```{r save-out}
#| echo: false
#| messange: false
#| warning: false
#| eval: false

# Write output from analysis to file for ease of use 
bk.rockfish.output <- list(
  yoy.trend = yoy.trend, # year-specific estimates of density for young of year
  bk.trend = bk.trend   # year-specific estimates of density for >10cm black rockfish.
)

save(bk.rockfish.output, file= paste0(work_dir,"OCNMS_YOY-SBYT-indices.Rdata"))
```
