---
title: "OCNMS recruitment analysis: Black and yellowtail rockfish complex"
author: "Ole Shelton & Nick Tolimieri"
date: " `r Sys.Date()` "
format:
  pdf:
    toc: false
    fig-pos: H
    fig-align: center
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
      - heightrounded
bibliography: references.bib
csl: "oecologia.csl"
---

```{r setup}
#| echo: false
#| messange: false
#| warning: false

knitr::opts_chunk$set(echo = TRUE)

survey_year = 2024
SPECIES = 'SEME'
SPECIES = 'SEFL'
SPECIES = c('SEFL', 'SEME')


if(length(SPECIES)==1){
  if(SPECIES=='SEME'){
    SPP = 'Black rockfish' 
    spp='black rockfish'
    sciname = "Sebastes melanostictus"}

  if(SPECIES=='SEFL'){
    SPP = 'Yellowtail rockfish' 
    spp='yellowtail rockfish'
   sciname = "Sebastes flavidus"}
}  

if(length(SPECIES)==2){
  if(SPECIES[1] == 'SEFL' & SPECIES[2]=='SEME'){
  SPPx = c('Black rockfish','Yellowtail rockfish')
  SPPz = c("SEBYT", "SEME","SEFL")
  SPP =  'Black and yellowtail rockfishes' 
  spp =  'black and yellowtail rockfishes'
  sciname = "Sebastes melanostictus and S. flavidus"}
}

# tinytex::reinstall_tinytex(repository = "illinois")
```

```{r}
#| echo: false
#| messange: false
#| warning: false

  library(tidyverse)
  library(xtable)
  library(terra)
  library(ggplot2)
  library(maps)
  library(mapdata)
  library(mapproj)
  library(tinytex)

# set directories ####
home_dir <- '~/Github/OCNMS/'
work_dir <- '~/Github/OCNMS/_01_Assessment-YOY-index/'
# nt personal computer
home_dir <- "C:/Users/tolim/Documents/GitHub/OCNMS/"
work_dir <- paste0(home_dir,'/_01_Assessment-YOY-index/')
data_dir = paste0(home_dir,'Data/',survey_year,'/')

# This is a padded data.frame which includes all transects
# includes zero observations
# includes low visibility transects, which will be dropped.

# All processing steps can be found in the Git repo: 
# "/GitHub/OCNMS/_Annual-UPdate/update-swath-fish.R") for all of the species. 
# We have filtered out the just the black rockfish data for this analysis. 

dat.fish <- readRDS( paste0(data_dir,"Fish_2015-", survey_year,".rds") )

# Filter to include only black rockfish or "small" size category (rockfish YOY)
dat.yoy <- dat.fish %>% filter(species %in% SPECIES) %>%
              bind_rows(.,
                dat.fish %>% filter(size_class %in% c("small")))
# canary and black/yellowtail complex -- look similar
dat.can <- dat.fish %>% filter(species %in% c("SEPI","SEPYT"))
          

# Drop transects where the visibility is less than the visibility threshold (2m):
VIS_LIM <- 2
dat.yoy <- dat.yoy %>% filter(vis_m >= VIS_LIM)
dat.can <- dat.can %>% filter(vis_m >= VIS_LIM)


```

These are derived data products relevant to estimating recruitment (young-of-year abundance) of the black rockfish (*Sebastes melanops*) and yellowtail rockfish (*S. flavidus*) complex (BYT); the data source is the NWFSC dive survey in Olympic Coast National Marine Sanctuary (OCNMS) conducted between 2015 and `r survey_year`. We estimate recruitment for the BYT domplex because it is difficult to distinguish small recruits for these species. Description of survey methods and aims are detailed in [@tolimieri2023changes]. We also estimate an abundance index for large (\>10 cm total length) black rockfishes and provide a size class analysis. We do not include large yellowtail rockfish because they are rarely seen on our dive surveys in this area.

# Data description

Divers on SCUBA conducted in situ surveys to count fish at each site along benthic belt transects (30 m by 2 m) following procedures modified from @malone2022large. Transects were conducted within or directly adjacent to canopy kelp beds (consisting of giant *Macrocystis pyrifera* or bull *Nereocystis luetkeana* kelps). In 2015 surveyed at 10 sites and conducted four (4) transects per site at 5 m depth ([Fig. @fig-site-map], [@shelton2018predictable]). From 2016 on, we surveyed at five (5) sites ([Fig. @fig-site-map]), sampling at two (2) locations within each site separated by \>100 m, and 2 depths within each location (5 and 10 m) Our goal was to complete six (6) replicate transects at each year-site-depth combination [@tolimieri2023changes].

During each fish transect, we counted and estimated the size (total length to nearest cm) of all fishes \>5 cm total length; the exception was rockfishes *Sebastes* spp., for which we estimated sizes of all individuals. Rockfishes $\leq$ 10 cm were considered young-of-year. Divers also estimated horizontal visibility on each transect by determining the distance at which the lead diver could distinguish their buddy's extended fingers. Transects with visibility less than 2 m were excluded from analyses.

As noted above, it is difficult to visually distinguish many rockfish species when they are small. Therefore, on our surveys, we categorized juvenile rockfishes into five (5) groups established in the literature [@johansson2018seasonal; @markel2020contrasting]:

\(1\) Yellowtail and black (YTB) included both yellowtail (*S. flavidus*) and black (*S. melanops*) rockfishes

\(2\) The copper/quillback/brown (CQB) group included copper (*S. caurinus*), quillback (*S. maliger*), and brown (*S. auriculatus*) rockfishes

\(3\) Canary (*S. pinniger*)

\(4\) Blue rockfish (*S. mystinus*)

\(5\) Unidentified individuals were categorized as juvenile rockfishes

The estimated recruitment trend for (1) the black rockfish and yellowtail rockfish complex (BYT complex), is presented here.

```{r map-prep}
#| echo: false
#| messange: false
#| warning: false

  #### READ IN WASHINGTON MARINE SHORELINES (FROM http://www.ecy.wa.gov/services/gis/data/data.htm
  
  #base.dir <- getwd()
#   gis.dir<- "/Users/ole.shelton/GitHub/OCNMS/Data/PG Shore shore_thematic_UTM10/"
#   setwd(gis.dir)
#   # shore.poly	<-	readShapePoly("shore_thematic_UTM10.shp")
#   
#   shore.poly	<-	readOGR(dsn=".","shore_thematic_UTM10")
#   shore		<-	data.frame(shore.poly)
# 
#   shore.dat	<-	 as(shore.poly,"data.frame")
#   shore.dat$NEW	<- 1:nrow(shore.dat)
#   shore.poly@data$id 	= rownames(shore.poly@data)
#   shore.points		<-	fortify(shore.poly,region="id")
#   shore.df			  <- merge(shore.points,shore.poly@data,by="id")
#   write.csv(shore.df,file="shore df.csv",row.names=F)
  shore.df        <- read.csv("../Data/PG Shore shore_thematic_UTM10/shore df.csv")
  
  # setwd(base.dir)
  PROJ <- "+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
  
  temp <- vect(cbind(shore.df$long,shore.df$lat),crs=PROJ)
  temp2 <- project(temp,"+proj=longlat +datum=WGS84")
  shore.df[,c("lon2","lat2")] <- geom(temp2)[,c("x","y")]
  
  loc <-  read.csv( paste0(work_dir,"Survey locations.csv"),fileEncoding="latin1")
  # update years
  loc$type[loc$type == '2015-22'] <- paste0('2015-',survey_year)
  loc <- loc %>% filter(type %in% c("2015 only",paste0('2015-',survey_year)))
  loc$nudge_x = 0.09
  loc$nudge_x[loc$Site =="Pt. of the Arches"] = 0.12
  loc$nudge_x[loc$Site =="Tatoosh Is."] = -0.08
  loc$nudge_x[loc$Site =="Chibahdehl"] = 0.08
  loc$nudge_x[loc$Site =="Neah Bay"] = 0.08
  loc$nudge_y = 0
  #loc$nudge_y[loc$Site =="Neah Bay"] = 0
  loc$nudge_y[loc$Site =="Chibahdehl"] = 0.01
  #loc$nudge_y[loc$Site =="Tatoosh Is."] = 0
  # loc.new				<- cbind(loc$lon, loc$latitude)
  # loc.new.sp		<- SpatialPoints(loc.new,CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
  # 
  # loc.new.sp.NAD83	<- spTransform(loc.new.sp, CRS(PROJ))
  # new.dat			      <- as(loc.new.sp.NAD83,"data.frame")
  # colnames(new.dat)	<- c("LonUTM","LatUTM")
  # loc <- data.frame(loc,new.dat)
  
  
  x.lim =c(-124.85,-124.3)
  y.lim =c(47.65,48.4)
  
  #wa <- map_data('worldHires','USA')
  #quartz(width=3,height=5)
  #wa <- map_data("state",region = c('washington'))
  site_map	<-	ggplot()+
				theme_bw() +
				theme(legend.key =  element_rect(fill = "white", colour = "white"),
					panel.border = element_rect(fill = NA, colour = "black",size=1.5)) +
		  #  geom_polygon(data=wa,aes(long,lat,group=group),fill=8,color="black")	+
				#coord_cartesian(xlim = x.lim,ylim=y.lim)     +
        geom_polygon(data=shore.df[shore.df$POLYTYPE == "Land" | shore.df$POLYTYPE == "Island" ,],
                     aes(lon2,lat2,group=group),fill=grey(0.7),color=NA)+
        geom_point(data= loc %>% filter(type %in% c(paste0('2015-',survey_year),"2015 only")),aes(x=lon,y=lat,color=type)) +
        scale_color_discrete("Survey")+
            geom_text(data= loc %>% filter(type %in% c(paste0('2015-',survey_year),"2015 only")),
                  aes(label=Site,x=lon,y=lat),size=2.5,
                  #check_overlap = TRUE,
                  nudge_x=loc$nudge_x,
                  nudge_y=loc$nudge_y
                  ) +
        xlab("Longitude") +
			  ylab("Latitude") +
        coord_fixed(ratio=1.3,xlim =x.lim,ylim=y.lim) +
    theme(axis.text.x = element_text(angle=90))
	
	#dev.off()
  		# 	bGrid +bAxis + bBack + bTics +

```

```{r fig-site-map}
#| echo: false
#| messange: false
#| warning: false
#| fig-width: 5
#| fig-height: 5
#| fig-cap: "Dive survey locations along the coast of Washington state."

site_map
```

Transects were omitted from analyses if the horizontal visibility was \<2 m. The following tables ([@tbl-samples-YS]) show how the number fish transects with visibility \>2 m were distributed across depth, site, and years ([@tbl-samples-YSD]), as well as across sites and year ([@tbl-samples-YS]). 2015 includes only surveys conducted at 5-m depth; other years have data approximately evenly split between 5-m and 10-m depths.

```{r table-samples-prep}
#| echo: false
#| messange: false
#| warning: false

library(xtable)

# here are the sites ordered from S to N
nom <- c("Destruction Island","Teahwhit Head","Cape Johnson","Rock 305","Cape Alava","Point of the Arches","Anderson Point","Tatoosh Island","Chibadehl Rocks","Neah Bay")
N.trans <- dat.yoy %>% distinct(year,site,zone,transect,observer) %>% group_by(year,site,zone) %>%
              dplyr::summarise(N=length(transect))
N.trans.site.year <- N.trans %>% group_by (year,site) %>% summarise(N=sum(N))

N.trans$site <- factor(N.trans$site,levels=nom)
N.trans.site.year$site <- factor(N.trans.site.year$site,levels=nom)

N.trans <- N.trans %>% arrange(site)
#A <- left_join(A,nom.merge,by=c("site"="nom"))
N.trans.1 <- pivot_wider(N.trans, id_cols = c("year","zone"),
                                names_from=c("site"),
                                values_from=c("N")) 
N.trans.site.year <- N.trans.site.year %>% arrange(site,year)
N.trans.2 <- pivot_wider(N.trans.site.year, id_cols = c("year"),
                                names_from=c("site"),
                                values_from=c("N")) 
N.trans.site.year <- N.trans.site.year %>% arrange(year)
N.trans.3 <- pivot_wider(N.trans.site.year, id_cols =c("site") ,
                                names_from= c("year"),
                                values_from=c("N")) 

N.trans.2[is.na(N.trans.2)] <- 0
N.trans.3[is.na(N.trans.3)] <- 0
N.trans.2 <- N.trans.2 %>% arrange(year)

N.trans.3$site <- factor(N.trans.3$site,levels=nom)
N.trans.3 <- N.trans.3 %>% arrange(site)

N.trans.1[is.na(N.trans.1)] <- 0
N.trans.1 <- N.trans.1 %>%ungroup() %>% dplyr::select(-year,-zone) %>% mutate(Total= rowSums(.)) %>%
              pull(Total) %>%
              bind_cols(N.trans.1,.) %>%
              arrange(year,zone)
colnames(N.trans.1)[ncol(N.trans.1)] <- "Total"

N.trans.3 <- N.trans.3 %>% dplyr::select(-site) %>% colSums(.) %>% data.frame() %>% 
                t() %>% bind_cols(site="TOTAL",.) %>% bind_rows(N.trans.3,.)
```

```{r tbl-samples-YSD}
#| echo: false
#| messange: false
#| warning: false
#| tbl-cap: "Number of transects conducted by year, site and depth zone.  Only transects that had at least 2m visibility are included"

library(knitr)
N.trans.1 = N.trans.1 %>% rename(Year = year, Zone = zone)

kable(N.trans.1)
```

\newpage

```{r tbl-samples-YS}
#| echo: false
#| messange: false
#| warning: false
#| tbl-cap: "Number of transects conducted by year and site.  Only transects that had at least 2-m visibility are included"


N.trans.3 = N.trans.3 %>% rename(Site = site)

kable(N.trans.3)
```

# Abundance trends

## Recruitment: BYT young-of-year abundance trends

To calculate the average density of BYT complex in each year, we first calculate the mean density and standard error per site in each year. This approach means we are treating each transect as a i.i.d. sample of YOY density within each site and thus we ignore differences in abundance by depth zone.

From these site-year level means, we calculated a year-specific mean density by simulation. Specifically, for each year we independently drew a mean density for each site using a t-distribution with $\mu$ (the estimated site mean), $\sigma$ (the estimated site-specific standard error) and degrees of freedom, $\tau$. So for the $i^{th}$ realization, for site $s$ in year $y$ we have a predicted density, $X_{isy}$

\begin{align}  & X_{isy} \sim T(\mu_{sy},\sigma_{sy},\tau_{sy}) \\ \end{align}

and then the predicted density for a single realization in a given year is the mean among sites observed. We repeat the simulation 100,000 times to provide an estimated mean density and uncertainty for a given year ([Fig. @fig-yoy-ts]).

Nearly all small rockfish fall into the 4 to 7 cm length range and all are considered to have recruited from the plankton during the calender year of the survey. Therefore, we view the density of \<10cm rockfish to be an indicator of recruitment for the black/yellowtail rockfish complex ([Fig. @fig-yoy-ts]).

```{r yoy-ts-prep}
#| echo: false
#| messange: false
#| warning: false

# GENERATE INDEXES BY SITE and AMONG SITES
#dat.yoy.summary <- dat.yoy %>% filter(species =="SEBYT")
dat.yoy.summary <- dat.yoy %>% filter(species %in% SPPz,
                                      size_class=='small')

#dat.yoy.summary %>% distinct(species,year,size_class)
  
# summarize using means and standard errors
dat.yoy.summary <- dat.yoy.summary %>%
                    # group_by(year,site,species) %>% 
                    group_by(year,site) %>% 
                    summarise(Mean=mean(Count),
                              VAR = var(Count),
                              SD=sd(Count),
                              N=length(Count),
                              SE=SD/sqrt(N))

# This is a simple calculation treating the means from each site as observed and iid samples of a grand Mean
dat.yoy.out.A <- dat.yoy.summary %>%
                  group_by(year)  %>%
                  summarise(grand.mean = mean(Mean),
                            N= length(Mean),
                            SE = sd(Mean)/sqrt(N))

      # We have a bunch of 0 observations, which preclude using weighting by the SE.
      # I looked extensively at two stage sampling expressions 
      # for design based estimators. 
      # They were complicated and didn't necessarily work for our design.  
      # Plus, they have troubles with the zero observations that we have.  
      # So I decided to simulate to determine the uncertainty.

#dim(dat.yoy.summary)
n.sim <- 100000

sim <- matrix(# simulate from a centered t-distribution.
        rt(nrow(dat.yoy.summary) * n.sim,dat.yoy.summary$N - 1) * dat.yoy.summary$SE +
                dat.yoy.summary$Mean, # Adjust to a different location and scale.
                nrow(dat.yoy.summary), n.sim) # Put it in a matrix.
sim[sim <0] <- 0
sim <- as.data.frame(sim)
colnames(sim) <- paste0("X",1:ncol(sim))

sim <- bind_cols(dat.yoy.summary %>% dplyr::select(year,site), sim) %>%
          pivot_longer(.,cols=3:ncol(.),values_to="val",names_to="real")

yoy.trend <- sim %>% group_by(year,real) %>%
          dplyr::summarise(mean.real = mean(val, na.rm=T)) %>% ungroup() %>%
          group_by(year) %>%
          dplyr::summarise(grand.mean = mean(mean.real, na.rm=T),
                           SE = sd(mean.real, na.rm=T),
                           q025 = quantile(mean.real,probs=0.025, na.rm=T),
                           q05 = quantile(mean.real,probs=0.05, na.rm=T),
                           q25 = quantile(mean.real,probs=0.25, na.rm=T),
                           q75 = quantile(mean.real,probs=0.75, na.rm=T),
                           q95 = quantile(mean.real,probs=0.95, na.rm=T),
                           q975 = quantile(mean.real,probs=0.975, na.rm=T))

# Make yoy rockfish time-series plots
yoy.ts <- ggplot() + 
  geom_point(data=dat.yoy.summary, aes(x=year,y=Mean),alpha=0.2) +
  geom_errorbar(data=dat.yoy.summary,
                aes(x=year,ymin=Mean-SE,ymax=Mean+SE),
                width=0,
                alpha=0.2) +
  geom_point(data=yoy.trend, aes(x=year,y=grand.mean),
             color="blue",size=3)  +
  geom_errorbar(data=yoy.trend,
                aes(x=year,ymin=q025,ymax=q975),
                width=0,
                color="blue",size=0.5) +
  geom_errorbar(data=yoy.trend,
                aes(x=year,ymin=q25,ymax=q75),
                width=0,
                color="blue",size=1.2) +
  scale_y_sqrt(expression("Young-of-Year Density (fish 120 m"^"-3"*")"),
                  #trans="sqrt",
                  limits=c(0,NA),
                  expand=c(0,NA),
                  breaks=c(0,1,5,10,25,50,100),
                  labels=c("0","1","5","10","25","50","100")) +
  xlab("Year") +
  #geom_hline(yintercept=0,color="red")+
  theme_bw()
  

# save out ####
dat.yoy.sites = dat.yoy.summary
write.csv(yoy.trend, paste0(work_dir, "Estimated-YOY-trend-coast-", Sys.Date(),".csv"), row.names = F)
write.csv(dat.yoy.sites, paste0(work_dir, "Estimated-YOY-trend-sites-", Sys.Date(),".csv"), row.names = F)


```

```{r fig-yoy-ts}
#| echo: false
#| messange: false
#| warning: false
#| fig-cap: "Time-series of estimated young-of-year rockfish (black-yellowtail complex) density on the Washington coast. Black points show means and standard errors for individual sites. Blue points show coastwide density estimates, interquartile range and 95\\% intervals for each year. Note y-axis is square root."

yoy.ts
```

\newpage

## Large (\>10 cm total length) black rockfish

We used the same approach as above to calculate the average density of large black rockfish in each year ([Fig. @fig-spp-ts]). Yellowtail rockfish are not common in our dive surveys and are not considered below.

```{r }
#| echo: false
#| messange: false
#| warning: false


# GENERATE INDEXES BY SITE and AMONG SITES
dat.bk = dat.fish %>% filter(species== 'SEME', size_class=="large")

#check to see that this only contains large rockfish (should be species "SEME")
#dat.yoy.summary %>% distinct(species,year,size_class)
  
# summarize using means and standard errors
dat.bk.summary <- dat.bk %>%
                    group_by(year,site,species) %>% 
                    summarise(Mean=mean(Count),
                              VAR = var(Count),
                              SD=sd(Count),
                              N=length(Count),
                              SE=SD/sqrt(N))

# This is a simple calculation treating the means from each site as observed and iid samples of a grand Mean
dat.bk.out.A <- dat.bk.summary %>%
                  group_by(year)  %>%
                  summarise(grand.mean = mean(Mean),
                            N= length(Mean),
                            SE = sd(Mean)/sqrt(N))

      # We have a bunch of 0 observations, which preclude using weighting by the SE.
      # I looked extensively at two stage sampling expressions 
      # for design based estimators. 
      # They were complicated and didn't necessarily work for our design.  
      # Plus, they have troubles with the zero observations that we have.  
      # So I decided to simulate to determine the uncertainty.

#dim(dat.yoy.summary)
n.sim <- 100000

sim <- matrix(# simulate from a centered t-distribution.
        rt(nrow(dat.bk.summary) * n.sim,dat.bk.summary$N - 1) * dat.bk.summary$SE +
                dat.bk.summary$Mean, # Adjust to a different location and scale.
                nrow(dat.bk.summary), n.sim) # Put it in a matrix.
sim[sim <0] <- 0
sim <- as.data.frame(sim)
colnames(sim) <- paste0("X",1:ncol(sim))

sim <- bind_cols(dat.bk.summary %>% dplyr::select(year,site), sim) %>%
          pivot_longer(.,cols=3:ncol(.),values_to="val",names_to="real")

bk.trend <- sim %>% group_by(year,real) %>%
          dplyr::summarise(mean.real = mean(val)) %>% ungroup() %>%
          group_by(year) %>%
          dplyr::summarise(grand.mean = mean(mean.real),
                           SE = sd(mean.real),
                           q025 = quantile(mean.real,probs=0.025),
                           q05 = quantile(mean.real,probs=0.05),
                           q25 = quantile(mean.real,probs=0.25),
                           q75 = quantile(mean.real,probs=0.75),
                           q95 = quantile(mean.real,probs=0.95),
                           q975 = quantile(mean.real,probs=0.975))

# Make black rockfish time-series plots
bk.rockfish.ts <- ggplot() + 
  geom_point(data=dat.bk.summary, aes(x=year,y=Mean),alpha=0.2) +
  geom_errorbar(data=dat.yoy.summary,
                aes(x=year,ymin=Mean-SE,ymax=Mean+SE),
                width=0,
                alpha=0.2) +
  geom_point(data=bk.trend, aes(x=year,y=grand.mean),
             color="red",size=3)  +
  geom_errorbar(data=bk.trend,
                aes(x=year,ymin=q025,ymax=q975),
                width=0,
                color="red",size=0.5) +
  geom_errorbar(data=bk.trend,
                aes(x=year,ymin=q25,ymax=q75),
                width=0,
                color="red",size=1.2) +
  scale_y_continuous(expression("Black rockfish density (fish 120 m"^"-3"*")")) +
  xlab("Year") +
   theme_bw()
  
```

```{r fig-spp-ts}
#| echo: false
#| messange: false
#| warning: false
#| fig-cap: "Time-series of estimated black rockfish density on the Washington coast. Black points show means and standard errors for individual sites. Red points show coastwide density estimates, interquartile range and 95\\% intervals for each year"

print(bk.rockfish.ts)
```

\newpage

## Size structure of black rockfish 2015-`r survey_year`

In addition to abundance data, we visually estimate size (total length) for all individuals observed during the surveys. Young-of-year show remarkably limited variation in size and so we exclude them from the analysis. The data here are for black rockfish; yellowtail are not presented because the adults are uncommon at our sites. Plots of size distribution grouped into 5-cm bins are shown in [Figures  @fig-spp-size-bin1] & [-@fig-size-bin2]. @fig-size-dist2 shows the size distribution 1-cm increments.

```{r }
#| echo: false
#| messange: false
#| warning: false

# This data comes from the processing script:
# source("_00_process-length-structure-BLYT.R")

load("Black rockfish_lengths_2015-2024.Rdata")

# all black rockfish (include YOYs in 5cm bins)
dat.spp.all.binned <- length.dat$dat.spp.all.binned  
 # all black rockfish (include YOYs)
dat.spp.all        <- length.dat$dat.spp.all  
# adult black rockfish (>10cm in 5cm bins)
dat.spp.ad.binned  <- length.dat$dat.spp.ad.binned 
# adult black rockfish 
dat.spp.ad         <- length.dat$dat.spp.ad 


################3 PLOTS
p.spp.size1 <- ggplot(dat.spp.ad) +
                    geom_col(aes(x=SIZE,y=COUNT)) +
                    facet_grid(SITE~YEAR) +
                    labs(x="Length(cm)",y="Count") +
                    theme_bw()
#p.spp.size1

p.spp.size2 <- ggplot(dat.spp.ad) +
  geom_col(aes(x=SIZE,y=COUNT,color=SITE,fill=SITE)) +
  facet_wrap(~YEAR,ncol=1) +
  labs(x="Length(cm)",y="Count") +
  theme_bw()
#p.spp.size2

p.spp.size3 <- ggplot(dat.spp.ad) +
  geom_col(aes(x=SIZE,y=COUNT)) +
  facet_wrap(~YEAR,ncol=1) +
  labs(x="Length(cm)",y="Count") +
  theme_bw()
#p.spp.size3

### BINNED (5cm bins)
dfx = dat.spp.ad.binned 
dfx$SITE = stringr::str_remove(dfx$SITE, " Island")
dfx$SITE = stringr::str_remove(dfx$SITE, "Cape ")
p.spp.size.bin1 <- ggplot(dfx) +
  geom_col(aes(x=bin,y=COUNT)) +
  facet_grid(SITE~YEAR) +
  labs(x="Length (5cm bins)",y="Count") +
  theme_bw()+
  theme( axis.text.x = element_text(size=7, angle=90),
         strip.text.y.right = element_text(size=8) )
#p.spp.size.bin1

p.spp.size.bin2 <- ggplot(dat.spp.ad.binned) +
  geom_col(aes(x=bin,y=COUNT,color=SITE,fill=SITE),width=4) +
  facet_wrap(~YEAR,ncol=1) +
  labs(x="Length (5 cm bins)",y="Count") +
  theme_bw() + theme(axis.title.y = element_text(size=7))
#p.spp.size.bin2

###
p.spp.size.bin3 <- ggplot(dat.spp.ad.binned) +
  geom_col(aes(x=bin,y=COUNT,color=SITE,fill=SITE),width=4) +
  facet_wrap(~YEAR,ncol=1) +
  labs(x="Length (5 cm bins)",y="Count") +
  theme_bw()
#p.spp.size.bin3


## 3 INCLUDE YOY sizes

p.spp.size.all.bin1 <- ggplot(dat.spp.all.binned) +
  geom_col(aes(x=bin,y=COUNT)) +
  facet_grid(SITE~YEAR) +
  labs(x="Length (5cm bins)",y="Count") +
  scale_y_continuous(trans="sqrt") +
  theme_bw() 
#p.spp.size.all.bin1


```

```{r fig-spp-size-bin1}
#| echo: false
#| messange: false
#| warning: false
#| fig-width: 6
#| fig-height: 5
#| fig-cap: "Black rockcish size distributions by 5-cm size bins for years and sites."


print(p.spp.size.bin1)

```

\newpage

```{r fig-size-bin2}
#| echo: false
#| messange: false
#| warning: false
#| fig-height: 8
#| fig-cap: "Black rockfish size distribution by 5-cm size bins plotted by site, summed across sites within each year"
 
print(p.spp.size.bin2)

```

\newpage


```{r fig-size-dist2}
#| echo: false
#| messange: false
#| warning: false
#| fig-height: 8
#| fig-cap: "Black rockfish size distribution by 1-cm size bins plotted by site, summed across sites within each year. "

print(p.spp.size2)

```

```{r save-out}
#| echo: false
#| messange: false
#| warning: false

# Write output from analysis to file for ease of use 
bk.rockfish.output <- list(
  yoy.trend = yoy.trend, # year-specific estimates of density for young-of-year
  dat.yoy.sites, # yoy trend by site
  bk.trend = bk.trend   # year-specific estimates of density for >10cm black rockfish.
)

save(bk.rockfish.output, 
     file = paste0(work_dir,"OCNMS_YOY-and-adult-indices-",Sys.Date(),".Rdata"))
```

\newpage

# References

<div id="refs"></div>
